### 2024-01-11 paper_draft/main.tex
adding paper

\documentclass[tikz,fleqn,12pt]{wlscirep}
\usepackage[english]{babel}
\usepackage[autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{bm}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage[nobiblatex]{xurl}
\usepackage{amsmath}
\DeclareMathAlphabet{\altmathcal}{OMS}{cmsy}{m}{n}
\usepackage{mathptmx}
\usepackage{float}
\usepackage{graphicx}
\usepackage{subcaption}
\usetikzlibrary{arrows.meta,positioning,fit}
\tikzset{%
   neuron missing/.style={
    draw=none, 
    scale=2,
    text height=0.333cm,
    execute at begin node=\color{black}$\vdots$
  },
}
\urlstyle{same}

\title{SIR-RL: Reinforcement Learning for Optimized Policy Control during Epidemiological Outbreaks in Emerging Market and Developing Economies}
\author[1,*]{Maeghal Jain}
\author[1]{Ziya Uddin}

\affil[1]{BML Munjal University}
\affil[*]{e-mail: maeghaljain@gmail.com}
\begin{abstract}
The outbreak of COVID-19 has highlighted the intricate interplay between public health and economic stability on a global scale. This paper aims to develop a reinforcement learning framework to balance health and economic outcomes during infectious disease outbreaks. The framework utilizes the SIR model without vital dynamics and incorporates globally comparable government responses. The study acknowledges the limitations of deterministic models and proposes the use of deep reinforcement learning to reduce input dimensionality and normalize input. While the model offers transparency in terms of its dependency on the reward policy defined, it recognizes the need for a comprehensive consideration of decision factors beyond pure reinforcement learning results.
\end{abstract}
\begin{document}
\flushbottom
\maketitle
\thispagestyle{empty}
\subsection*{Introduction}
In the past, global spread of infectious diseases was largely due to colonization, slavery, and war, leading to widespread illness and death from diseases like tuberculosis, polio, smallpox, and diphtheria. Medical advancements, better access to health care, and improved sanitation have worked towards improving the situation of mortality and morbidity linked to infectious diseases in the past twenty years. However, in low and lower-middle income countries the burden of infectious diseases still persists. The rapid pace of urbanization in low and middle-income countries, along with the rise in populations living in crowded, poor-quality homes, has led to new conditions that favor the emergence of infectious diseases~\cite{Baker2022, inequitableworld}.

Recently, the COVID-19 pandemic caused a havoc worlwide. Till date there have been 772 million cases and more than 6 million deaths~\cite{WHOCovid}. The pandemic triggered the sharpest economic recession in modern history with a 3\% decline, much worse than during the 2008-09 financial crisis~\cite{IMFCovid}. As nations grappled with the immediate health crisis, the economic fallout disproportionately affected vulnerable populations and exacerbated existing inequalities. Lockdowns and restrictions imposed to curb the spread of the virus led to widespread unemployment, business closures, and disruptions in global supply chains~\cite{Nicola2020}. The challenges faced by low and lower-middle income countries were particularly acute, highlighting the intricate interplay between public health and economic stability on a global scale~\cite{Gagnon2023}.

The need for a nuanced understanding of how interventions impact both health outcomes and economic indicators became increasingly evident, prompting a comprehensive examination by epidemiologists to assist policy makers~\cite{Anderson2020}. The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022, Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, stringency measures/non-pharmaceutical interventions (NPIs)~\cite{Jalloh2022, Caldwell2021, Ferguson2020}, and financial considerations~\cite{DeFoo2023}. Despite the numerous studies conducted, very few explore how common interventions meet multiple policy objectives or how a precise articulation of the main policy goals directs the selection of the most effective interventions in terms of health and economic results~\cite{Hollingsworth2011, Song2022, Pangallo2023, Ash2022, Ohi2020, PADMANABHAN2021102676}. The economic impact of the COVID-19 pandemic varied between rich and poor countries. Although COVID-19 deaths had a slightly larger negative effect on the Gross Domestic Product (GDP) in advanced economies, this difference was not statistically significant. However, lockdown restrictions were found to have a more damaging impact on economic activity in emerging and developing economies~\cite{Gagnon2023, Redlin2022,Liang2021}. It's also suggested that an increase in COVID-19 cases was associated with the introduction of harsher NPIs and lockdown measures could be relaxed once vaccination rates increase~\cite{Redlin2022, Patel2021}.

Many economists have studied the effect of COVID-19 on the economy of nations~\cite{Gagnon2023, GagnonKorea, Deb2020,Eichenbaum2021}. In advanced economies like Korea, where the stringency index was below the median the recession was milder than other advanced economies like the United Kingdom where the stringency was much higher~\cite{GagnonKorea}, they achieved it mostly with very aggressive testing, contact tracing, and enforced quarantines and isolations~\cite{Lim2023, KoreaMinister}. In India, social distancing and containment measures have been effective in reducing the number of COVID-19 cases but have come with economic costs. Social distancing had the most adverse effect on the economy in areas with high urbanization~\cite{Deb2020}.

In this paper we want to optimize the government policies regarding stringency. Therefore, alongside epidemiological data, we use the measures of globally comparable government responses~\cite{Hale2021}. We use the simple SIR model without vital dynamics~\cite{Hethcote1989, Hethcote2008, ALLEN2017128} as it is assumed that the time scale is short enough so that can be neglected~\cite{Cooper2020}. By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we can effectively address the real-world conditions and propose a solution that is both effective and extendable. However, the study has limitations; the deterministic SIR model fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}.

In order to capture competing costs within the environment and achieve a balance between health and economic outcomes, we intend to employ reinforcement learning~\cite{Nguyen2022, Bastani2021, Song2022, Ohi2020, PADMANABHAN2021102676}. Not only does the formulation of the model deal better with competing costs, but it also offers a more transparency behind the reasoning of the decisions being made in such circumstances. When we conceptualize our problem as a reinforcement learning task, an agent is tasked with making decisions in an environment with the aim of optimizing cumulative rewards. This approach relies on trial and error, in contrast to dynamic programming, which assumes complete knowledge of the environment~\cite{Lavet2018}. We make the use of deep reinforcement learning, which is an advancement to reinforcement learning as it allows to reduce the input dimensionality and normalize the input~\cite{Lavet2018, Arulkumaran2017, Henderson_Islam_Bachman_Pineau_Precup_Meger_2018, Mnih2015}. While our model is more transparent in terms of its dependency on the reward policy defined, it has its limitations. A universal optimal policy may not suit diverse socioeconomic contexts due to variations in healthcare resources and economic vulnerabilities across countries, regions, or cities and a comprehensive consideration of decision factors, extending beyond pure reinforcement learning results is needed~\cite{Song2022, Dunn2017, Demir2006}. Also, the current model doesn't account for the herd immunity reached through vaccination which it can be extended to account for~\cite{Marinov2022}.

\subsection*{Methodology}
In SIR-RL, we use a compartmental model to model infectious disease. Here we divide people of the population, based on whether they are yet to come into contact with an infected person (Susceptible), are infectious themselves (Infectious), or have recovered from the infection (Recovered). These compartments create the SIR model which can be represented as follows:

\begin{figure}[H]
\centering    
\begin{tikzpicture}[node distance=1cm, auto,
    >=Latex, 
    every node/.append style={align=center},
    int/.style={draw, minimum size=1cm}]
   \node [int] (S)             {$S$};
   \node [int, right=of S] (I) {$I$};
   \node [int, right=of I] (R) {$R$};
   \coordinate[right=of I] (out);
   \path[->] (S) edge node {$\lambda$} (I)
             (I) edge node {$\gamma$} (out);
\end{tikzpicture}
\end{figure}
\begin{equation}
  \frac{d S}{d t}=-\lambda S
  \label{eq:S_without_lockdown}
\end{equation}
\begin{equation}
  \frac{d I}{d t}=\lambda S-\gamma I
  \label{eq:I_without_lockdown}
\end{equation}
\begin{equation}
  \frac{d R}{d t}=\gamma I
  \label{eq:R_without_lockdown}
\end{equation}

Here, $\lambda$ is the force of infection, it is the rate at which susceptible individuals acquire an infectious disease~\cite{hens_aerts_faes_shkedy_lejeune_vandamme_beutels_2010}. It depends on other factors:

\begin{equation}
  \lambda = pc\frac{I}{N}
  \label{eq:lambda_force_of_infection}
\end{equation}

Here, 
$c$ is the average number of contacts a susceptible person makes per day.
$p$ is the probability of the susceptible person becomes infectious after coming into contact with an infectious person.
$\frac{I}{N}$ is the proportion of the contacts that are infectious.

And, $\beta$ the effective transmission rate is defined as:

\begin{equation}
  \beta = p c
  \label{eq:beta_effective_transmission_rate}
\end{equation}

During an epidemic, the fundamental drivers of an epidemic growth is the rate of infection $\beta$ i.e. the average number of infections per infected case and the infectious period $1/\gamma$ i.e. the average period for which the infected case is infected for. Epidemics can only happen if the case is infectious enough for long enough and this defined by $R_0 = \beta / \gamma$. Here, $R_0$ is The average number of secondary infections caused by each infected case, in an otherwise fully susceptible population.

At the peak of an epidemic $R_0$ declines as there are no more susceptible people left in the pool, therefore, $R_{eff}$ comes into play. $R_{eff}$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic - therefore, it takes into account the existing immunity of the system.

\begin{equation}
  R_{eff} = R_0 \frac{S}{N}
  \label{eq:R_effective}
\end{equation}

$S$ is the number of susceptible people, $N$ is the total population. At the start of an epidemic when everyone is susceptible, $R_{eff} = R_0$ as, $S = N$ (i.e. the whole population is susceptible). $\beta$ and $\gamma$ are also used to define probability of and infectious individual infecting another individual $\beta / (\beta + \gamma)$ and the probability of recovery, $\gamma / (\beta + \gamma)$.

Most government policies look at the value of $R_{eff}$ to come us with an effective strategy to combat the disease as the fate of the evolution of the disease depends upon it. When $R_{eff}$ is less than one, the infected population $I$ will steadily decline to zero. Conversely, if $R_{eff}$ is greater than one, the infected population will increase. In other words, when $\frac{dI(t)}{dt} < 0 \Rightarrow R_{eff} < 1$ and $\frac{dI(t)}{dt} > 0 \Rightarrow R_{eff} > 1},$ the effective reproductive rate $R_{eff}$ serves as a critical threshold that determines whether an infectious disease will rapidly extinguish or escalate into an epidemic~\cite{Cooper2020}.

To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from January, 2020, to October, 2022, we simply define the cost to calibrate the model using the huber loss~\cite{huberloss}:

\begin{equation}
  L_{\delta}(y, f(x)) = 
  \begin{cases}
    \frac{1}{2}{(y - f(x))^2} & \text{for } |y - f(x)| \leq \delta, \\
    \delta \cdot (|y - f(x)| - \frac{1}{2}\delta) & \text{otherwise.}
  \label{eq:huberloss}
  \end{cases}
\end{equation}
\begin{equation}
  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
  \label{eq:cost_SIR_model}
\end{equation}

Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered. Minimizing this cost function using the Nelder-Mead method\cite{Gao2010}, we get:

\begin{equation}
  \beta_{optimal} = 0.07565820733562756
  \label{eq:beta_optimal_without_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.046594691406826524
  \label{eq:gamma_optimal_without_lockdown}
\end{equation}
\begin{equation}
  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.6237516560640421
  \label{eq:r0_without_lockdown}
\end{equation}
\begin{equation}
  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 69949738.9977647
  \label{eq:cost_without_lockdown}
\end{equation}
See figure \ref{fig:sir_without_lockdown_IND} to see how the model compares with the real data.

Now that, we have a model -- we want to say $\beta$ is a time-varying parameter controlled by the stringency index, $s$~\cite{SIRLockdown,SIRLockdown2,analytical_covid_lockdown_model}. The stringency index is a composite measure based on nine response indicators including school closures, workplace closures, and travel bans, rescaled to a value from 0 to 100 (100 = strictest)~\cite{owidcoronavirus}. This index simply records the strictness of government policies and does not measure or imply the appropriateness or effectiveness of a country's response i.e. a higher score does not necessarily mean that a country's response is "better" than others lower on the index.

To define the new time-varying beta that is dependent on the current stringency index, we say,

\begin{equation}
  \frac{dS}{dt} = -\beta(t) \frac{S I}{N}
  \label{eq:S_with_lockdown}
\end{equation}
\begin{equation}
  \frac{dI}{dt} = \beta(t) \frac{S I}{N} - \gamma I
  \label{eq:I_with_lockdown}
\end{equation}
\begin{equation}
  \frac{dR}{dt} = \gamma I
  \label{eq:R_with_lockdown}
\end{equation}
\begin{equation}
  \beta(t) = \beta_{optimal} + (1 - s(t)/100)
  \label{eq:beta_time_varying_and_stringency_dependent}
\end{equation}

Where, $s(t)$ is the stringency index at time $t$ and is scaled down by a factor of 100 to normalize it and bring it in the range $[0, 1]$. Optimizing this using the Nelder-Mead method we get:

\begin{equation}
  \beta_{optimal} = 0.302941346409129
  \label{eq:beta_optimal_with_lockdown}
\end{equation}
\begin{equation}
  \gamma_{optimal} = 0.07020572542840187
  \label{eq:gamma_optimal_with_lockdown}
\end{equation}
\begin{equation}
  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.315051864510377
  \label{eq:r0_with_lockdown}
\end{equation}
\begin{equation}
  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 68015990.34546319
  \label{eq:cost_with_lockdown}
\end{equation}

See figure \ref{fig:sir_with_lockdown_IND} to see how the model compares with the real data. The model's fit is better, as the $cost = 68015990.34546319$ is $1933748.65230152$ less than the model without lockdown.

Now, that we have set up the following relation between $\beta$ and $s(t)$, we investigate how stringency index affects the normalized Gross domestic product (GDP). To do this we fit a polynomial equation of the third degree to the points $(x, y)$ where $x$ is the $s(t)$ stringency at time t, and $y$ the normalized GDP and minimize the squared error. For India after fitting a 3 degree polynomial we get the following equation:

\begin{equation}
    \begin{split}
      GDP = -5.96640236e-05 s^{3} + 6.65064332e-03 s^{2} -2.23109924e-01 s^{1} \\
    + 1.01357226e+02
    \end{split}
    \label{eq:gdp_modelled_with_stringency_IND}
\end{equation}

Given that the government is an agent that takes decisions in a deterministic environment defined above, we use reinforcement learning to model the competing costs of the environment. This environment is formally as a Markov decision process, and can be described as follows:
\begin{itemize}
    % \item \url{https://jmlr.org/papers/volume22/20-1364/20-1364.pdf} RL Paper that references \url{https://araffin.github.io/slides/rl-tuto-jnrr19/#/13} RL ppt (for our RL notes).
    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP ($(GDP_{predicted} - GDP_{min}) / (GDP_{max} - GDP_{min})$), $R_{eff}$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible, infected and recovered. The starting states are simply these values at the starting date and no previous actions.
    \item Actions $\altmathcal{A}$: The stringency index variable was analyzed with a sample size of 1034. The mean value was approximately 59.62, with a standard deviation of 22.82. The minimum value was 0, while the maximum value reached 100. And the differences between two consecutive stringencies had a mean of 0.030474, and standard deviation of 2.387331, with the minimum being -14.360000 and maximum 55.560000. Based on this we define the discrete action space. There are 7 actions for the agent, it can keep the stringency index same, reduce/increase by 2.5, reduce/increase by 5, and reduce/increase by 10 given that the stringency index doesn't exceed 100 or go below 0.
    \item Transition dynamics $\altmathcal{T}\left(\mathbf{s}_{t+1} \mid \mathbf{s}_t, \mathbf{a}_t\right)$ that map a state-action pair at time $t$ onto a distribution of states at time $t+1$. This state transition is defined by the SIR model with lockdown and the model of how stringency index affects the GDP.
    \item Immediate reward $\altmathcal{R}\left(\mathbf{s}_t, \mathbf{a}_t, \mathbf{s}_{t+1}\right)$. Here we define a reward strategy however it should be noted that this strategy can be easily changed to prioritize different needs.
    \item Discount Factor $\gamma \in[0,1]$, where lower values place more emphasis on immediate rewards. Here, we choose the default discount factor of 0.99.
\end{itemize}
In general, the policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$. If the MDP is episodic, i.e., the state is reset after each episode of length $T$, then the sequence of states, actions and rewards in an episode constitutes a trajectory or rollout of the policy. Every rollout of a policy accumulates rewards from the environment, resulting in the return $R=\sum_{t=0}^{T-1} \gamma^t r_{t+1}$. The goal of RL is to find an optimal policy, $\pi^*$, which achieves the maximum expected return from all states.

The stringency index emerges as a critical factor influencing both the Gross Domestic Product (GDP) and the rate of infection spread. The decision to escalate or de-escalate the stringency index is a strategic one, with significant implications. Increasing the stringency decreases the spread of the infection. Conversely, it must be noted that herd immunity can only be achieved when the epidemic reaches its peak i.e. when the effective reproductive number is equal to one ($R_{eff} = 1$). This can only happen by lowering the stringency index which would allow the natural dynamics of the epidemic to transpire such that the population of susceptible individuals has depleted enough such that it is insufficient to propagate the disease further. Therefore, stringency is used to control the number of infected people and slow down the rate at which the epidemic reaches its peak, so that hospitals could house the number of infected people.

With this basis we define the reward function~\cite{Ohi2020}. In Deep Reinforcement Learning (DRL), positive rewards promote and negative rewards demote actions. The agent tries to generate such a policy/knowledge to avoid the discouraging situation by following the policy. By designing a proper reward function, it is possible to generate such an agent that may follow the human desired situation.

The agent observes the percentage of the population that is susceptible, infected, recovered, and the GDP which extrapolated from the stringency index and all the past moves played, i.e., all the past stringency indexes decided by the agent. The following values are fed into a simple network. Stable Baselines3 supports multiple inputs by using Dict Gym space. This can be done using MultiInputPolicy, which by default uses the CombinedExtractor features extractor to turn multiple inputs into a single vector, handled by the network arch network.

% we utilized a custom combined extractor, a unique type of Long Short-Term Memory (LSTM) network, for processing our data. This network was designed to handle different types of input data encapsulated in an observation space, which is a common concept in reinforcement learning representing the information an agent observes from the environment. The CustomCombinedExtractor class, which inherits from the BaseFeaturesExtractor class, was initialized with an observation space that is a dictionary of Gym spaces. The Gym spaces module provides a set of high-level, easy-to-use data structures for defining observation and action spaces, which are essential components in reinforcement learning. The extractor contains a dictionary of individual extractors for each key in the observation space. If the subspace shape is equal to the total number of days plus one, an LSTM extractor is created with an input size of 1 and a hidden size of 16. This LSTM extractor is designed to handle sequential data, which is particularly useful for time-series data such as ours. For the 'other_stats' key, a Flatten extractor is used to flatten the input data. In the forward method, the observations are processed through their respective extractors. For LSTM extractors, the final hidden state is used as the output. All the processed observations are then concatenated into a single tensor. The observation space for the deep learning type is defined as a dictionary of four keys: 'stringency', 'normalized_gdp', 'r_eff', and 'other_stats'. Each key is associated with a Box space, which is a multi-dimensional box that allows for a range of numerical values. The shape of the Box space defines the dimensionality of the data, and the low and high parameters define the range of possible values. This LSTM network and the associated observation space provide a robust framework for processing and learning from our complex, multi-dimensional data.

\newcommand{\DrawNeuronalNetwork}[2][]{
\xdef\Xmax{0}
\foreach \Layer/\X/\Col/\Miss/\Lab/\Count/\Content [count=\Y] in {#2}
{\pgfmathsetmacro{\Xmax}{max(\X,\Xmax)}
 \xdef\Xmax{\Xmax}
 \xdef\Ymax{\Y}
}
\foreach \Layer/\X/\Col/\Miss/\Lab/\Count/\Content [count=\Y] in {#2}
{\node[anchor=south] at ({2*\Y},{\Xmax/2+0.1}) {\Layer};
 \foreach \m in {1,...,\X}
 {
  \ifnum\m=\Miss
   \node [neuron missing] (neuron-\Y-\m) at ({2*\Y},{\X/2-\m}) {};
  \else
   \node [circle,fill=\Col!50,minimum size=1cm] (neuron-\Y-\m) at 
  ({2*\Y},{\X/2-\m}) {\Content};
 \ifnum\Y=1
  \else
   \pgfmathtruncatemacro{\LastY}{\Y-1}
   \foreach \Z in {1,...,\LastX}
   {
    \ifnum\Z=\LastMiss
    \else
     \draw[->] (neuron-\LastY-\Z) -- (neuron-\Y-\m);
    \fi
    }
  \fi
 \fi
 \ifnum\Y=1
  \ifnum\m=\X
   \draw [overlay] (neuron-\Y-\m) -- (state);
  \else
   \ifnum\m=\Miss
   \else
    \draw [overlay] (neuron-\Y-\m) -- (state);
   \fi
  \fi
 \else
 \fi     
 }
 \xdef\LastMiss{\Miss}
 \xdef\LastX{\X}
}
}
\begin{tikzpicture}[x=1.5cm, y=1.5cm,
>=stealth,font=\sffamily,nodes={align=center}]
 \begin{scope}[local bounding box=T]
  \path  node[draw,minimum width=6em,minimum height=4em] (state) {State};
  \begin{scope}[local bounding box=NN]
   \DrawNeuronalNetwork{Input Layer/5/green/4///,
     Hidden Layer/5/blue/4//11/,
     Output Layer/4/red/3//11/}
  \end{scope}
  \path (NN.south) node[below]{Estimated parameter\\ $\theta$};
  \path(NN.east) -- node[above]{Policy\\ $\Pi(\theta,a)$}++ (4em,0);
 \end{scope} 
 \node[fit=(T),label={[anchor=north west]north west:Agent},inner sep=1em,draw]
  (TF){};
 \node[below=3em of TF,draw,inner sep=1em] (Env) {Environment};
 \draw[<-] (TF.200) -- ++ (-1em,0) |- (Env.160) node[pos=0.45,right]{$r_t$};
 \draw[<-] (TF.180) -- ++ (-2em,0) |- (Env.180) node[pos=0.45,left]{$s_t$};
 \draw[->] (NN.east) -- ++ (7em,0)node[right]{$a_t$} |- (Env);
\end{tikzpicture}

% Use this for the RL formulae: https://arxiv.org/pdf/1708.05866.pdf and https://www.nature.com/articles/s41598-022-23668-x
% for reward function use this: https://www.nature.com/articles/s41598-020-79147-8

\subsection*{Results}
% https://www.nature.com/articles/s41598-020-79147-8 (see this to write results)
Using the simple SIR model (\ref{eq:S_without_lockdown}, ~\ref{eq:cost_without_lockdown}) to model the disease dynamics we get:
\begin{figure}[H]
  \centering
  \caption{SIR Model for India}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/SIR_model_IND.eps}
    \caption{SIR model for India}
  \end{subfigure}
  \label{fig:SIR_model_IND}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[scale=0.25]{images/SIR_model_infections_IND.eps}
    \caption{Infections modelled with SIR model}
  \end{subfigure}
  \label{fig:SIR_model_infections_IND}
\end{figure}

Combining the lockdown dynamics in the SIR model we get this following:
\begin{figure}[H]
  \centering
  \caption{SIR Model with lockdown for India}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/SIR_model_with_lockdown_IND.eps}
    \caption{SIR model with lockdown}
  \end{subfigure}
  \label{fig:SIR_model_with_lockdown_IND}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[scale=0.25]{images/SIR_model_with_lockdown_infections_IND.eps}
    \caption{Infections modelled with SIR model with Lockdown}
  \end{subfigure}
  \label{fig:SIR_model_with_lockdown_infections_IND}
\end{figure}

Combining vaccination dynamics with SIR with lockdown model we get the following:
\begin{figure}[H]
  \centering
  \caption{SIRV Model with lockdown for India}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/SIRV_model_with_lockdown_IND.eps}
    \caption{SIRV model with lockdown}
  \end{subfigure}
  \label{fig:SIRV_model_with_lockdown_IND}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[scale=0.25]{images/SIRV_model_with_lockdown_infections_IND.eps}
    \caption{Infections modelled with SIRV model with Lockdown}
  \end{subfigure}
  \label{fig:SIRV_model_with_lockdown_infections_IND}
\end{figure}

Stringency affects the economy of developing countries. Below are scatter plots with pearson correlation, for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May, 2020 to October, 2022.

\begin{figure}[H]
  \centering
  \caption{Stringency and GDP for Developing Economies}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_IND.eps}
    \caption{Stringency and Normalized GDP for India}
  \end{subfigure}
  \label{fig:stringency_vs_gdp_IND}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_IND.eps}
    \caption{Normalized GDP modelled with Stringency for India}
  \end{subfigure}
  \label{fig:gdp_modelled_with_stringency_IND}

  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_MEX.eps}
    \caption{Stringency and Normalized GDP for Mexico}
  \end{subfigure}
  \label{fig:stringency_vs_gdp_MEX}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_MEX.eps}
    \caption{Normalized GDP modelled with Stringency for Mexico}
  \end{subfigure}
  \label{fig:gdp_modelled_with_stringency_MEX}

  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_BRA.eps}
    \caption{Stringency and Normalized GDP for Brazil}
  \end{subfigure}
  \label{fig:stringency_vs_gdp_BRA}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_BRA.eps}
    \caption{Normalized GDP modelled with Stringency for Brazil}
  \end{subfigure}
  \label{fig:gdp_modelled_with_stringency_BRA}
\end{figure}

We observe that stringency has a negative impact on the normalized Gross domestic product (GDP) of each country. Therefore, in some countries policies made during an epidemic have competing costs. This is not the case for Advanced Economies like (USA, Japan, Canada).

\begin{figure}[H]
  \centering
  \caption{Stringency and GDP for Advanced Economies}
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_USA.eps}
    \caption{Stringency and Normalized GDP for United States}
  \end{subfigure}
  \label{fig:stringency_vs_gdp_USA}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_USA.eps}
    \caption{Normalized GDP modelled with Stringency for United States}
  \end{subfigure}
  \label{fig:gdp_modelled_with_stringency_USA}

  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_JPN.eps}
    \caption{Stringency and Normalized GDP for Japan}
  \end{subfigure}
  \label{fig:stringency_vs_gdp_JPN}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_JPN.eps}
    \caption{Normalized GDP modelled with Stringency for Japan}
  \end{subfigure}
  \label{fig:gdp_modelled_with_stringency_JPN}

  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_CAN.eps}
    \caption{Stringency and Normalized GDP for Canada}
  \end{subfigure}
  \label{fig:stringency_vs_gdp_CAN}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_CAN.eps}
    \caption{Normalized GDP modelled with Stringency for Canada}
  \end{subfigure}
  \label{fig:gdp_modelled_with_stringency_CAN}
\end{figure}

\subsection*{Discussion}
The paper seeks to inspire epidemiologists by highlighting the advancements achieved through the application of reinforcement learning in policy making during the pandemic. We introduce a virtual environment that closely simulates a pandemic scenario and thoroughly explore innovative strategies for disease mitigation using reinforcement learning. Our proposed approach demonstrates compelling efficacy in achieving optimal decision-making, effectively balancing the formidable challenges posed by the pandemic and economic considerations. We are confident that this research contribution will forge a connection between epidemic studies and reinforcement learning, offering valuable insights to help humanity better defend against the ongoing pandemic crisis.

\subsection*{Experiment Settings}
\subsection{Dataset}
The population-level epidemiological data can be obtained from the 'Our World In Data COVID-19' dataset: \url{https://ourworldindata.org/coronavirus} or more specifically: \url{https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-data.csv}~\cite{owidcoronavirus}. Quaterly GDP data can be obtained from the 'Organisation for Economic Co-operation and Development': \url{https://www.oecd-ilibrary.org/economics/data/main-economic-indicators/main-economic-indicators-complete-database_data-00052-en}~\cite{economic_indicators_data}.

\subsection{Code}
We used stable-baseline3~\cite{JMLR:v22:20-1364}, Pytorch~\cite{paszke2019pytorch}, Scipy~\cite{Virtanen_2020}, Pandas, Matplotlib, Python~\cite{ScientificPython}.
Code: \url{https://github.com/psymbio/sir_rl}
% \bibliographystyle{plain}
\bibliography{sample}
\end{document}

### 2024-01-11 sir_lockdown.m

@@ -1,906 +1,0 @@
-clear all; clc; close all; set(0,'DefaultLegendAutoUpdate','off');
-
-% Set latex as default interpreter
-set(groot, 'defaulttextinterpreter','latex');  
-set(groot, 'defaultAxesTickLabelInterpreter','latex');  
-set(groot, 'defaultLegendInterpreter','latex');
-
-% Epidemiological Parameters
-Rnot     = 2.5;            % Total number of transmissions of an infectious (I) person (Source: EC), Atkeson is 3
-Tinf     = 7;              % Duration of infectious period (Source: EC), Atkeson is 18
-beta     = Rnot/Tinf;      % Transmission rate of an infectious (I) person 
-gamma    = 1/Tinf;         % Exit rate from I state into recovery or death (R)
-
-implicit = 0;
-
-T    = 400;    % Length of simulation
-dt   = 0.1;    % Time step 
-Nt   = T/dt+1;
-time = linspace(0,T,Nt);
-
-mu = zeros(3,Nt); 
-S = zeros(Nt,1); I = zeros(Nt,1); R = zeros(Nt,1);
-N = zeros(Nt,1); D = zeros(Nt,1);
- 
-% Transition matrix
-A_t = cell(Nt,1);
-
-% Initial condition
-I0 = 0.35*10^(-4); % March 1st 2020 so that March 31st there are appx. 150,000 cases (50% more than measured)
-S0 = 1-I0;
-R0 = 0;
-
-
-%% SIR MODEL EVENTUAL CASES FOR COMPARISON: solution to log(S) = R_0*(S - 1) or log(1-R) = -R_0*R
-fun = @(x) log(x) - Rnot*(x-1);    % function of x alone
-S_st = fzero(fun,[0.0001,0.99999]);
-R_st = 1-S_st;
-
-S_herd = 1./Rnot;
-R_herd = (1-1./Rnot);
-
-mu0 = [S0,I0,R0]';
-mu(:,1) = mu0;
-
-% DO NOTHING
-
-for n=1:Nt
-    S(n) = mu(1,n);
-    I(n) = mu(2,n);
-    R(n) = mu(3,n);
-    Re(n) = Rnot.*S(n);
-    A_t{n} = [
-        % into S     into I   into R
-        -beta*I(n), beta*I(n),  0;    % from S
[…]

### 2024-01-11 sir_lockdown_2.m

@@ -1,220 +1,0 @@
-clear all; clc; close all; set(0,'DefaultLegendAutoUpdate','off');
-
-% Set latex as default interpreter
-set(groot, 'defaulttextinterpreter','latex');  
-set(groot, 'defaultAxesTickLabelInterpreter','latex');  
-set(groot, 'defaultLegendInterpreter','latex');
-
-% Epidemiological Parameters
-Rnot     = 2.5;            % Total number of transmissions of an infectious (I) person (Source: EC), Atkeson is 3
-Tinf     = 7;              % Duration of infectious period (Source: EC), Atkeson is 18
-beta     = Rnot/Tinf;      % Transmission rate of an infectious (I) person 
-gamma    = 1/Tinf;         % Exit rate from I state into recovery or death (R)
-
-implicit = 0;
-
-T    = 400;    % Length of simulation
-dt   = 0.1;    % Time step 
-Nt   = T/dt+1;
-time = linspace(0,T,Nt);
-
-mu = zeros(3,Nt); 
-S = zeros(Nt,1); I = zeros(Nt,1); R = zeros(Nt,1);
-N = zeros(Nt,1); D = zeros(Nt,1);
- 
-% Transition matrix
-A_t = cell(Nt,1);
-
-% Initial condition
-I0 = 0.35*10^(-4); % March 1st 2020 so that March 31st there are appx. 150,000 cases (50% more than measured)
-S0 = 1-I0;
-R0 = 0;
-
-
-%% SIR MODEL EVENTUAL CASES FOR COMPARISON: solution to log(S) = R_0*(S - 1) or log(1-R) = -R_0*R
-fun = @(x) log(x) - Rnot*(x-1);    % function of x alone
-S_st = fzero(fun,[0.0001,0.99999]);
-R_st = 1-S_st;
-
-S_herd = 1./Rnot;
-R_herd = (1-1./Rnot);
-
-mu0 = [S0,I0,R0]';
-mu(:,1) = mu0;
-
-% DO NOTHING
-
-for n=1:Nt
-    S(n) = mu(1,n);
-    I(n) = mu(2,n);
-    R(n) = mu(3,n);
-    Re(n) = Rnot.*S(n);
-    A_t{n} = [
-        % into S     into I   into R
-        -beta*I(n), beta*I(n),  0;    % from S
[…]

### 2024-01-11 paper_draft/main.tex
including trash, time varying nu model, changing the rl

@@
 
-\title{SIR-RL: Reinforcement Learning for Optimized Policy Control during Epidemiological Outbreaks in Emerging Market and Developing Economies}
+\title{SIRV-RL: Reinforcement Learning for Optimized Policy Control during Epidemiological Outbreaks in Emerging Market and Developing Economies}
 \author[1,*]{Maeghal Jain}
@@
 
-\subsection*{Methodology}
+\subsection*{Mathematical Formulation}
 In SIR-RL, we use a compartmental model to model infectious disease.
@@
   \centering
-  \caption{SIR Model for India}
-  \begin{subfigure}[t]{0.48\textwidth}
+  \caption{SIR Model with lockdown for India}
+  \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[width=\linewidth]{images/SIR_model_IND.eps}
-    \caption{SIR model for India}
+    \includegraphics[scale=0.5]{images/SIR_model_IND.eps}
+    \caption{SIR model with lockdown}
+    \label{fig:SIR_mode_IND}
   \end{subfigure}
-  \label{fig:SIR_model_IND}
-  \hfill
-  \begin{subfigure}[t]{0.48\textwidth}
+  \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.25]{images/SIR_model_infections_IND.eps}
+    \includegraphics[scale=0.50]{images/SIR_model_infections_IND.eps}
     \caption{Infections modelled with SIR model}
+    \label{fig:SIR_model_infections_IND}
   \end{subfigure}
-  \label{fig:SIR_model_infections_IND}
 \end{figure}
@@
   \caption{SIR Model with lockdown for India}
-  \begin{subfigure}[t]{0.48\textwidth}
+
+  \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[width=\linewidth]{images/SIR_model_with_lockdown_IND.eps}
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_IND.eps}
     \caption{SIR model with lockdown}
+    \label{fig:SIR_model_with_lockdown_IND}
   \end{subfigure}
-  \label{fig:SIR_model_with_lockdown_IND}
-  \hfill
-  \begin{subfigure}[t]{0.48\textwidth}
+
+  \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.25]{images/SIR_model_with_lockdown_infections_IND.eps}
+    \includegraphics[scale=0.50]{images/stringency_varying_with_time_IND.eps}
+    \caption{Stringency varying with Time}
+    \label{fig:stringency_varying_with_time_IND}
+  \end{subfigure}
+\end{figure}
+\begin{figure}[H]\ContinuedFloat
+  \begin{subfigure}[t]{\textwidth}
+    \centering
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_infections_IND.eps}
     \caption{Infections modelled with SIR model with Lockdown}
+    \label{fig:SIR_model_with_lockdown_infections_IND}
   \end{subfigure}
-  \label{fig:SIR_model_with_lockdown_infections_IND}
 \end{figure}
@@
   \caption{SIRV Model with lockdown for India}
-  \begin{subfigure}[t]{0.48\textwidth}
+
+  \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[width=\linewidth]{images/SIRV_model_with_lockdown_IND.eps}
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_IND.eps}
     \caption{SIRV model with lockdown}
+    \label{fig:SIRV_model_with_lockdown_IND}
   \end{subfigure}
-  \label{fig:SIRV_model_with_lockdown_IND}
-  \hfill
-  \begin{subfigure}[t]{0.48\textwidth}
+
+  \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.25]{images/SIRV_model_with_lockdown_infections_IND.eps}
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_IND.eps}
     \caption{Infections modelled with SIRV model with Lockdown}
+    \label{fig:SIRV_model_with_lockdown_infections_IND}
   \end{subfigure}
-  \label{fig:SIRV_model_with_lockdown_infections_IND}
+
+\end{figure}
+
+For time-varying $\nu$ we get the folllowing:
+\begin{figure}[H]
+  \centering
+  \caption{SIRV Model with lockdown and time-varying $\nu$ for India}
+
+  \begin{subfigure}[t]{\textwidth}
+    \centering
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_time_varying_nu_IND.eps}
+    \caption{SIRV model with lockdown and time-varying $\nu$}
+    \label{fig:SIRV_model_with_lockdown_time_varying_nu_IND}
+  \end{subfigure}
+
+  \begin{subfigure}[t]{\textwidth}
+    \centering
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_time_varying_nu_IND.eps}
+    \caption{Infections modelled with SIRV model with Lockdown and time-varying $\nu$}
+    \label{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND}
+  \end{subfigure}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \caption{Loss for Different Models}
+  \includegraphics[width=\linewidth]{images/comparing_costs_IND.eps}
+  \label{fig:comparing_costs_IND}
 \end{figure}

### 2024-01-11 04_reinforcement_learning.py

@@ -1,4 +1,4 @@
-from rl_utils_new_complex_reward_sirvlockdown import *
+from rl_utils_new_complex_reward_sirvlockdown_nu_varying import *
 import time
 
 # base testing code

### 2024-01-11 05_reinforcement_run.py

@@ -1,4 +1,4 @@
-from rl_utils_new_complex_reward_sirvlockdown import *
+from rl_utils_new_complex_reward_sirvlockdown_nu_varying import *
     
 env = SIREnvironment()
 env.reset()

### 2024-01-11 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -1,0 +1,362 @@
+import pandas as pd
+import numpy as np
+from stable_baselines3 import PPO
+from stable_baselines3.common.env_checker import check_env
+from stable_baselines3.common.torch_layers import BaseFeaturesExtractor
+from stable_baselines3.common.preprocessing import get_flattened_obs_dim
+import torch as th
+import torch.nn as nn
+import os
+import gymnasium as gym
+from gymnasium import spaces
+from scipy.integrate import odeint
+from collections import deque
+import json
+import matplotlib.pyplot as plt
+
+from constants import LOCATION_CHOOSEN, OUTPUT_DIR, DATA_CACHE_DIR, STRINGENCY_BASED_GDP, OPTIMAL_VALUES_FILE, MODELS_DIR, RL_LEARNING_TYPE
+OUTPUT_RL = os.path.join(OUTPUT_DIR, "rl")
+
+with open(OPTIMAL_VALUES_FILE, 'r') as f:
+    optimal_values_read = f.read()
+    optimal_values = json.loads(optimal_values_read)
+optimal_beta = optimal_values['optimal_beta']
+optimal_gamma = optimal_values['optimal_gamma']
+
+stringency_data_points = np.arange(0, 100, 0.5)
+fit_line_loaded = np.poly1d(np.load(STRINGENCY_BASED_GDP))
+predicted_gdp = fit_line_loaded(stringency_data_points)
+MIN_GDP = min(predicted_gdp)
+MAX_GDP = max(predicted_gdp)
+
+data_path = os.path.join(DATA_CACHE_DIR, LOCATION_CHOOSEN + "_merged_data.csv")
+df = pd.read_csv(data_path)
+df['date'] = pd.to_datetime(df['date'])
+TOTAL_DAYS = (max(df['date']) - min(df['date'])).days
+print(f'Total Days: {TOTAL_DAYS}')
+
+START_STRINGENCY = df.loc[min(df.index), ['stringency_index']].item()
+print(f'Start stringency: {START_STRINGENCY}')
+N = df.loc[min(df.index), ['N']].item()
+y0 = df.loc[min(df.index), ['S']].item(), df.loc[min(df.index), ['I']].item(), df.loc[min(df.index), ['R']].item()
+nu_varying = list(df['nu_varying_with_time'])
+N_DISCRETE_ACTIONS = 7
+
+def deriv(y, t, N, beta, gamma, nu_varying, lockdown):
+    S, I, R = y
[…]

### 2024-01-11 trash/05_reinforcement_run.py

@@ -1,0 +1,248 @@
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from collections import deque
+import gymnasium as gym
+import numpy as np
+from gymnasium import spaces
+from scipy.integrate import odeint
+from stable_baselines3 import PPO
+import os
+import random
+
+from constants import LOCATION_CHOOSEN, OUTPUT_DIR, DATA_CACHE_DIR, MODELS_DIR
+OUTPUT_RL = os.path.join(OUTPUT_DIR, "rl")
+
+random.seed(0)
+np.random.seed(0)
+
+data_path = os.path.join(DATA_CACHE_DIR, LOCATION_CHOOSEN + "_with_GDP.csv")
+df = pd.read_csv(data_path)
+df['date'] = pd.to_datetime(df['date'])
+print(df)
+
+stringency = df["stringency_index"]
+gdp = df[' Gross Domestic Product (GDP)  Normalised']
+fit_coefficients = np.polyfit(stringency, gdp, 4)
+fit_line = np.poly1d(fit_coefficients)
+
+stringency_data_points = np.arange(0, 100, 0.5)
+predicted_gdp = fit_line(stringency_data_points)
+
+MIN_GDP = min(predicted_gdp)
+MAX_GDP = max(predicted_gdp)
+print(f'Min GDP: {MIN_GDP}, Max GDP: {MAX_GDP}')
+beta_optimal = 0.03925422815833437
+gamma_optimal = 0.022659519392619114
+s_weight_optimal = -1.7905485677020872e-07
+
+data_path = os.path.join(DATA_CACHE_DIR, LOCATION_CHOOSEN + ".csv")
+df = pd.read_csv(data_path)
+df['date'] = pd.to_datetime(df['date'])
+df = df[(df['date'].dt.year == 2021) & (df['date'].dt.month >= 5) | (df['date'].dt.year == 2022)]
+
+df['N'] = df['population']
+df['S'] = df['population'] - (df['total_cases'] + df['people_fully_vaccinated'])
+df['I'] = df['total_cases']
+df['R'] = df['people_fully_vaccinated']
+TOTAL_DAYS = (max(df['date']) - min(df['date'])).days
+print(f'Total Days: {TOTAL_DAYS}')
+print(len(df['stringency_index']))
+
+def compute_cost(data, predictions):
+    return np.abs(data - predictions).mean()
+
+def deriv(y, t, N, beta, gamma):
[…]

### 2024-01-11 trash/rl_utils.py

@@ -1,0 +1,319 @@
+import pandas as pd
+import numpy as np
+from stable_baselines3 import PPO
+from stable_baselines3.common.env_checker import check_env
+from stable_baselines3.common.torch_layers import BaseFeaturesExtractor
+from stable_baselines3.common.preprocessing import get_flattened_obs_dim
+import torch as th
+import torch.nn as nn
+import os
+import gymnasium as gym
+from gymnasium import spaces
+from scipy.integrate import odeint
+from collections import deque
+import json
+import matplotlib.pyplot as plt
+
+from constants import LOCATION_CHOOSEN, OUTPUT_DIR, DATA_CACHE_DIR, STRINGENCY_BASED_GDP, OPTIMAL_VALUES_FILE, MODELS_DIR, RL_LEARNING_TYPE
+OUTPUT_RL = os.path.join(OUTPUT_DIR, "rl")
+
+with open(OPTIMAL_VALUES_FILE, 'r') as f:
+    optimal_values_read = f.read()
+    optimal_values = json.loads(optimal_values_read)
+optimal_beta = optimal_values['optimal_beta']
+optimal_gamma = optimal_values['optimal_gamma']
+optimal_stringency_weight = optimal_values['optimal_stringency_weight']
+
+stringency_data_points = np.arange(0, 100, 0.5)
+fit_line_loaded = np.poly1d(np.load(STRINGENCY_BASED_GDP))
+predicted_gdp = fit_line_loaded(stringency_data_points)
+MIN_GDP = min(predicted_gdp)
+MAX_GDP = max(predicted_gdp)
+
+data_path = os.path.join(DATA_CACHE_DIR, LOCATION_CHOOSEN + "_merged_data.csv")
+df = pd.read_csv(data_path)
+df['date'] = pd.to_datetime(df['date'])
+TOTAL_DAYS = (max(df['date']) - min(df['date'])).days
+print(f'Total Days: {TOTAL_DAYS}')
+
+START_STRINGENCY = df.loc[min(df.index), ['stringency_index']].item()
+N = df.loc[min(df.index), ['N']].item()
+y0 = df.loc[min(df.index), ['S']].item(), df.loc[min(df.index), ['I']].item(), df.loc[min(df.index), ['R']].item()
+N_DISCRETE_ACTIONS = 7
+
+def compute_cost(data, predictions):
+    return np.abs(data - predictions).mean()
+
[…]

### 2024-01-11 trash/rl_utils_old.py

@@ -1,0 +1,268 @@
+import pandas as pd
+import numpy as np
+from stable_baselines3 import PPO
+from stable_baselines3.common.env_checker import check_env
+from stable_baselines3.common.torch_layers import BaseFeaturesExtractor
+import torch as th
+import torch.nn as nn
+import os
+import gymnasium as gym
+import numpy as np
+from gymnasium import spaces
+from scipy.integrate import odeint
+from collections import deque
+import json
+import matplotlib.pyplot as plt
+
+from constants import LOCATION_CHOOSEN, OUTPUT_DIR, DATA_CACHE_DIR, STRINGENCY_BASED_GDP, OPTIMAL_VALUES_FILE, MODELS_DIR
+OUTPUT_RL = os.path.join(OUTPUT_DIR, "rl")
+
+with open(OPTIMAL_VALUES_FILE, 'r') as f:
+    optimal_values_read = f.read()
+    optimal_values = json.loads(optimal_values_read)
+optimal_beta = optimal_values['optimal_beta']
+optimal_gamma = optimal_values['optimal_gamma']
+optimal_stringency_weight = optimal_values['optimal_stringency_weight']
+
+stringency_data_points = np.arange(0, 100, 0.5)
+fit_line_loaded = np.poly1d(np.load(STRINGENCY_BASED_GDP))
+predicted_gdp = fit_line_loaded(stringency_data_points)
+MIN_GDP = min(predicted_gdp)
+MAX_GDP = max(predicted_gdp)
+
+data_path = os.path.join(DATA_CACHE_DIR, LOCATION_CHOOSEN + ".csv")
+df = pd.read_csv(data_path)
+df['date'] = pd.to_datetime(df['date'])
+df = df[(df['date'].dt.year == 2021) & (df['date'].dt.month >= 5) | (df['date'].dt.year == 2022)]
+
+df['N'] = df['population']
+df['S'] = df['population'] - (df['total_cases'] + df['people_fully_vaccinated'])
+df['I'] = df['total_cases']
+df['R'] = df['people_fully_vaccinated']
+TOTAL_DAYS = (max(df['date']) - min(df['date'])).days
+print(f'Total Days: {TOTAL_DAYS}')
+
+START_STRINGENCY = df.loc[min(df.index), ['stringency_index']].item()
+N = df.loc[min(df.index), ['N']].item()
[…]

### 2024-01-11 paper_draft/main.tex
paper edits

@@
 By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we can effectively address the real-world conditions and propose a solution that is both effective and extendable.
+The current model accounts for the recovery reached through vaccination~\cite{Marinov2022, MaurciodeCarvalho2023, Thater, Turkyilmazoglu2022, YALADANDA2022101052}.
 However, the study has limitations; the deterministic SIR model fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}.
@@
 A universal optimal policy may not suit diverse socioeconomic contexts due to variations in healthcare resources and economic vulnerabilities across countries, regions, or cities and a comprehensive consideration of decision factors, extending beyond pure reinforcement learning results is needed~\cite{Song2022, Dunn2017, Demir2006}.
-Also, the current model doesn't account for the herd immunity reached through vaccination which it can be extended to account for~\cite{Marinov2022}.
 
@@
 In SIR-RL, we use a compartmental model to model infectious disease.
-Here we divide people of the population, based on whether they are yet to come into contact with an infected person (Susceptible), are infectious themselves (Infectious), or have recovered from the infection (Recovered).
+We iteratively develop this model to fit the actual data better.
+In an SIR model people of the population are divided based on whether they are yet to come into contact with an infected person (Susceptible), are infectious themselves (Infectious), or have recovered from the infection (Recovered).
 These compartments create the SIR model which can be represented as follows:
@@
 
-To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from January, 2020, to October, 2022, we simply define the cost to calibrate the model using the huber loss~\cite{huberloss}:
+To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define the cost to calibrate the model using the huber loss~\cite{huberloss}:
 
@@
 \begin{equation}
-  \beta_{optimal} = 0.07565820733562756
+  \beta_{optimal} = 0.04208479828695971
   \label{eq:beta_optimal_without_lockdown}
@@
 \begin{equation}
-  \gamma_{optimal} = 0.046594691406826524
+  \gamma_{optimal} = 0.02388356686032017
   \label{eq:gamma_optimal_without_lockdown}
@@
 \begin{equation}
-  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.6237516560640421
+  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.7620817917644795
   \label{eq:r0_without_lockdown}
@@
 \begin{equation}
-  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 69949738.9977647
+  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 85051490.53250012
   \label{eq:cost_without_lockdown}
@@
 \begin{equation}
-  \frac{dS}{dt} = -\beta(t) \frac{S I}{N}
+  \frac{dS}{dt} = -\beta (1 - (s(t)/100)) \frac{S I}{N}
   \label{eq:S_with_lockdown}
@@
 \begin{equation}
-  \frac{dI}{dt} = \beta(t) \frac{S I}{N} - \gamma I
+  \frac{dI}{dt} = -\beta (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
   \label{eq:I_with_lockdown}
@@
 \end{equation}
-\begin{equation}
-  \beta(t) = \beta_{optimal} + (1 - s(t)/100)
-  \label{eq:beta_time_varying_and_stringency_dependent}
-\end{equation}
 
@@
 \begin{equation}
-  \beta_{optimal} = 0.302941346409129
+  \beta_{optimal} = 0.4013340889432941
   \label{eq:beta_optimal_with_lockdown}
@@
 \begin{equation}
-  \gamma_{optimal} = 0.07020572542840187
+  \gamma_{optimal} = 0.09017476605499258
   \label{eq:gamma_optimal_with_lockdown}
@@
 \begin{equation}
-  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.315051864510377
+  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.450625230328213
   \label{eq:r0_with_lockdown}
@@
 \begin{equation}
-  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 68015990.34546319
+  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 98438821.45587364
   \label{eq:cost_with_lockdown}
@@
 
-See figure \ref{fig:sir_with_lockdown_IND} to see how the model compares with the real data.
-The model's fit is better, as the $cost = 68015990.34546319$ is $1933748.65230152$ less than the model without lockdown.
+Finally, an additional flow from the susceptible population to the can be shown by adding a vaccination rate $\nu$ in the model.
+
+\begin{equation}
+  \frac{dS}{dt} = -\beta (1 - (s(t)/100)) \frac{S I}{N} - \nu S
+  \label{eq:S_with_lockdown_and_nu}
+\end{equation}
+\begin{equation}
+  \frac{dI}{dt} = \beta (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
+  \label{eq:I_with_lockdown_and_nu}
+\end{equation}
+\begin{equation}
+  \frac{dR}{dt} = \gamma I + \nu S
+  \label{eq:R_with_lockdown_and_nu}
+\end{equation}
+
+Optimizing these equations with the Nelder-Mead method we get:
+
+\begin{equation}
+  \beta_{optimal} = 0.40897034072952304
+  \label{eq:beta_optimal_with_lockdown_and_nu}
+\end{equation}
+\begin{equation}
+  \gamma_{optimal} = 0.09196829370123338
+  \label{eq:gamma_optimal_with_lockdown_and_nu}
+\end{equation}
+\begin{equation}
+  \nu_{optimal} = 2.9044029843851936e-05
+  \label{eq:nu_optimal_with_lockdown_and_nu}
+\end{equation}
+\begin{equation}
+  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.446862329077204
+  \label{eq:r0_with_lockdown_and_nu}
+\end{equation}
+\begin{equation}
+  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 94636860.38436058
+  \label{eq:cost_with_lockdown_and_nu}
+\end{equation}
+
+However, as observed by the value of $\nu_{optimal}$ which is almost negligible and the overestimation of infected individuals in \ref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time, i.e., the flow of people from susceptibility to direct recovery changes with a time dependent vaccination rate.
+Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \ref{eq:beta_optimal_with_lockdown_and_nu} and \ref{eq:gamma_optimal_with_lockdown_and_nu}, we optimize the value of $\nu$ given a sub-interval of $[{start}, {start} + {window}]$ where the ${window} = 20$.
+Here, ${window}$ is a hyperparameter that can be changed so as to reduce the cost function.
+
+\begin{equation}
+  \frac{dS}{dt} = -\beta_{optimal}  (1 - (s(t)/100))  \frac{S I}{N} - \nu S
+  \label{eq:S_with_lockdown_and_nu_calc_nu}
+\end{equation}
+\begin{equation}
+  \frac{dI}{dt} = \beta_{optimal}  (1 - (s(t)/100)) \frac{S I}{N} - \gamma_{optimal} I
+  \label{eq:I_with_lockdown_and_nu_calc_nu}
+\end{equation}
+\begin{equation}
+  \frac{dR}{dt} = \gamma_{optimal} I + \nu S
+  \label{eq:R_with_lockdown_and_nu_calc_nu}
+\end{equation}
+
+This model gives us how $\nu$ varies with time.
+
+\begin{figure}[H]
+  \centering
+  \caption{$\nu$ Varying with Time}
+  \includegraphics[width=\linewidth]{images/nu_varying_with_time_IND.eps}
+  \label{fig:nu_varying_with_time_IND}
+\end{figure}
+
+Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations below:
+
+\begin{figure}[H]
+\begin{tikzpicture}[node distance=1cm, auto,
+  >=Latex, 
+  every node/.append style={align=center},
+  int/.style={draw, minimum size=1cm}]
+ \node [int] (S)             {$S$};
+ \node [int, right=of S] (I) {$I$};
+ \node [int, right=of I] (R) {$R$};
+ \coordinate[right=of I] (out);
+ \path[->] (S) edge node {$\lambda$} (I)
+           (I) edge node {$\gamma$} (out)
+           (S) edge[bend left=45] node[above] {$\nu(t)$} (R); % Arrow from S to R with text \nu(t)
+\end{tikzpicture}
+\end{figure}
+
+\begin{equation}
+  \frac{dS}{dt} = -\beta  (1 - (s(t)/100))  \frac{S I}{N} - \nu(t) S
+  \label{eq:S_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  \frac{dI}{dt} = \beta  (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
+  \label{eq:I_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  \frac{dR}{dt} = \gamma I + \nu(t) S
+  \label{eq:R_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  \beta_{optimal} = 0.5129414911377119
+  \label{eq:beta_optimal_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  \gamma_{optimal} = 0.12716245167626106
+  \label{eq:gamma_optimal_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.033749620081199
+  \label{eq:r0_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 29742589.698910963
+  \label{eq:cost_with_lockdown_and_time_varying_nu}
+\end{equation}
+
+See (\ref{fig:SIR_model_IND}, ~\ref{fig:comparing_costs_IND}) to see how the different models compare against each other.
 
@@
     \caption{SIR model with lockdown}
-    \label{fig:SIR_mode_IND}
+    \label{fig:SIR_model_IND}
   \end{subfigure}

### 2024-01-11 04_reinforcement_learning.py

@@ -28,7 +28,7 @@ if not os.path.exists(logdir):
 	os.makedirs(logdir)
 
 # https://stackoverflow.com/questions/56700948/understanding-the-total-timesteps-parameter-in-stable-baselines-models
-TIMESTEPS = 4570
+TIMESTEPS = 2742
 
 if RL_LEARNING_TYPE == "normal":
 	env = SIREnvironment()

### 2024-01-11 05_reinforcement_run.py

@@ -19,7 +19,8 @@ env.reset()
 
 # models/1704873042/5165.zip
 # models/1704874281/374740.zip
-model_path = os.path.join(MODELS_DIR, "1704874281", "374740.zip")
+# models/1704994098/2742.zip
+model_path = os.path.join(MODELS_DIR, "1704994098", "2742.zip")
 
 model = PPO.load(model_path, env=env)
 

### 2024-01-12 google_colab_upload/constants.py
google_colab_setup for testing, changing the reward function

@@ -1,0 +1,15 @@
+# Countries to try out: ['JPN' 'AUS' 'DEU' 'ZAF' 'MEX' 
+# 'TUR' 'FRA' 'ITA' 'BRA' 'GBR' 'CAN' 'KOR' 'USA' 'ESP'
+# 'IDN' 'IND' 'CHN' 'A5M' 'LUX']
+# because GDP data is only present for them
+LOCATION_CHOOSEN = "IND"
+LOCATION_CHOOSEN_2 = "India"
+INPUT_DIR = "input"
+OUTPUT_DIR = "output"
+DATA_CACHE_DIR = "data_cache"
+MODELS_DIR = "models"
+
+STRINGENCY_BASED_GDP = "stringency_based_gdp.npy"
+OPTIMAL_VALUES_FILE = "optimal_values.json"
+
+RL_LEARNING_TYPE = "deep" # can be deep or normal

### 2024-01-12 google_colab_upload/rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -51,8 +51,8 @@ def deriv(y, t, N, beta, gamma, nu_varying, lockdown):
     return dSdt, dIdt, dRdt
 
 def calculate_reward_weighted(gdp_min_max_normalized_list, r_eff_list):
-    GDP_WEIGHT_1 = 100 # change this value and see how it affects the reward
-    GDP_WEIGHT_2 = 200 # change this value and see how it affects the reward
+    GDP_WEIGHT_1 = 10 # change this value and see how it affects the reward
+    GDP_WEIGHT_2 = 20 # change this value and see how it affects the reward
     reward = []
     for i in range(len(gdp_min_max_normalized_list)):
         if r_eff_list[i] > 1.9:
@@ -127,12 +127,12 @@ class SIREnvironment(gym.Env):
         # REMEMBER: to change this definition of the reward in the render as well!!!
         # self.reward = self.normalized_GDP - (2 * self.r_eff)
 
-        reward_inertia = abs(diff)*-1*5
-        reward_r_eff = 10 if self.r_eff <= 1.9 else -10
-        reward_I_percentage = -5000 if self.I_proportion >= 0.003 else 20
+        reward_inertia = abs(diff)*-1*2
+        reward_r_eff = 1 if self.r_eff <= 1.9 else -1
+        reward_I_percentage = -5000 if self.I_proportion >= 0.002 else 0
 
-        gdp_reward_weight_1 = 100
-        gdp_reward_weight_2 = 200
+        gdp_reward_weight_1 = 10
+        gdp_reward_weight_2 = 20
         if self.r_eff > 1.9:
             reward_weighted = -20 * self.r_eff
         elif self.r_eff <= 1.9 and self.r_eff >= 1.5:
@@ -227,21 +227,20 @@ class SIREnvironment(gym.Env):
         legend.get_texts()[1].set_text(f'R_eff (modelled); R_eff=1 at {first_time_r_eff_modelled_1}')
         legend.get_texts()[2].set_text(f'R_eff (rl); R_eff=1 at {first_time_r_eff_1}')
 
-        hospital_capacity = 0.003
-        hospital_capacity_punishment = -5000
-        hospital_capacity_reward = 20
-        I_reward_actual = [hospital_capacity_punishment if I_percentage >= hospital_capacity else hospital_capacity_reward for I_percentage in self.df["I"] / self.df["N"]]
[…]

### 2024-01-12 paper_draft/main.tex
google_colab_setup for testing, changing the reward function
comparing rewards

@@
 
-At the peak of an epidemic $R_0$ declines as there are no more susceptible people left in the pool, therefore, $R_{eff}$ comes into play.
+At the peak of an epidemic $R_0$ declines as there are no more susceptible people left in the pool, therefore, $R_{eff}$ (effective reproductive number) comes into play.
 $R_{eff}$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic - therefore, it takes into account the existing immunity of the system.
@@
 \end{equation}
-See figure \ref{fig:sir_without_lockdown_IND} to see how the model compares with the real data.
+See figure \ref{fig:SIR_model_IND} to see how the model compares with the real data.
 
@@
 \begin{equation}
-  \frac{dI}{dt} = -\beta (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
+  \frac{dI}{dt} = \beta (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
   \label{eq:I_with_lockdown}
@@
 
-However, as observed by the value of $\nu_{optimal}$ which is almost negligible and the overestimation of infected individuals in \ref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time, i.e., the flow of people from susceptibility to direct recovery changes with a time dependent vaccination rate.
-Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \ref{eq:beta_optimal_with_lockdown_and_nu} and \ref{eq:gamma_optimal_with_lockdown_and_nu}, we optimize the value of $\nu$ given a sub-interval of $[{start}, {start} + {window}]$ where the ${window} = 20$.
+However, as observed by the value of $\nu_{optimal}$ from \ref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible (close to zero) and the overestimation of infected individuals in \ref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time.
+This suggests a time-varying vaccination rate, wherein the transition from susceptibility to direct recovery fluctuates with time.
+Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \ref{eq:beta_optimal_with_lockdown_and_nu} and \ref{eq:gamma_optimal_with_lockdown_and_nu}, we optimize the value of $\nu$ given a sub-interval of $[{start}, {start} + {window}]$ where the ${window} = 20$ \cite{Marinov2022}.
 Here, ${window}$ is a hyperparameter that can be changed so as to reduce the cost function.
@@
 The starting states are simply these values at the starting date and no previous actions.
+    \item TODO: FIX BELOW ITEM AFTER MODEL RUNS
     \item Actions $\altmathcal{A}$: The stringency index variable was analyzed with a sample size of 1034.
@@
 
+The reward function is parameterized to account for key factors influencing decision-making.
+To incentivize reduction of $R{eff}$ (effective reproductive number) a negative reward is imposed of $-20 * R_{eff}$, but as the $R_{eff}$ is between $[1.9, 1,5]$ we start to positively reward it with a multiple of the GDP ($100 * \textrm{min\_max\_normalized\_GDP}$) and if the $R_{eff}$ is below $1.5$ then $200 * \textrm{min\_max\_normalized\_GDP}$.
+Furthermore, there's a simple positive reward of $10$ if $R_{eff}$ is below the threshold value ($1.9$) and negative reward of $-10$ otherwise.
+To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| * -1 * 5$) [TODO: this factor should be reduced back to 2 from 5].
+If the proportion of the infected population were to rise above 0.003 the model is punished ($-5000$) and otherwise rewarded ($+20$).
+It should be realized there can be an infinite number of ways to design the reward function to be more human and upgrade the way a decision is taken given the situation \cite{AWSDeepRacer}.
+Therefore, this research act as a  framework for promoting the development of more efficient reward strategies for the same.
+
 The agent observes the percentage of the population that is susceptible, infected, recovered, and the GDP which extrapolated from the stringency index and all the past moves played, i.e., all the past stringency indexes decided by the agent.
@@
 This can be done using MultiInputPolicy, which by default uses the CombinedExtractor features extractor to turn multiple inputs into a single vector, handled by the network arch network.
+
+For data that varies with time we use a simple LSTM architecture.
 
@@
 \end{figure}
+Here, it can be observed that the SIR model accurately fits the susceptible population and recovered population but overestimates the infected population by a significant margin.
+Although overestimating the infected population may not always be problematic, in our specific case, it can create complications.
+This is because our research involves rewarding the agent when the proportion of infected individuals falls below a predetermined threshold.
+Consequently, an overestimation of the infected population could lead to incorrect decision-making and undesirable outcomes.
 
-Combining the lockdown dynamics in the SIR model we get this following:
+Combining the lockdown dynamics in the SIR model (\ref{eq:S_with_lockdown}, ~\ref{eq:cost_with_lockdown}) we get the following:
 \begin{figure}[H]
@@
 \end{figure}
+Here, it can be observed there's an overestimation of infected individuals, but, the two stages of the epidemic are being accounted for.
+This is what suggests that there might be depletion of infected individuals through vaccination.
 
-Combining vaccination dynamics with SIR with lockdown model we get the following:
+Combining vaccination dynamics with SIR with lockdown model (\ref{eq:S_with_lockdown_and_nu}, ~\ref{eq:cost_with_lockdown_and_nu}) we get the following:
 \begin{figure}[H]
@@
 \end{figure}
+Here, because the value of $\nu$ (\ref{eq:nu_optimal_with_lockdown_and_nu}) is negligible, it makes no change. However, time-varying $\nu$ shall be able to better account for the these dynamics.
 
-For time-varying $\nu$ we get the folllowing:
+For time-varying $\nu$ and SIRV model with lockdown (\ref{eq:S_with_lockdown_and_time_varying_nu}, ~\ref{eq:cost_with_lockdown_and_time_varying_nu}), we get the folllowing:
 \begin{figure}[H]
@@
 
-Stringency affects the economy of developing countries.
+With a time-varying $\nu$ (vaccination rate) and the effect of lockdown, our model is able to account for the infected individuals and reduce the cost in comparison to all the previously formalized models for the data.
+This shows how interventions and changes in the way people behave in response of a epidemic \cite{Caldwell2021} play a major role in the way the epidemic unfolds.
+
+However, Non-pharmaceutical Interventions (NPIs) come with costs for developing nations.
 Below are scatter plots with pearson correlation, for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May, 2020 to October, 2022.
@@
 
-We observe that stringency has a negative impact on the normalized Gross domestic product (GDP) of each country.
+It can be observed that stringency has a negative impact on the normalized Gross domestic product (GDP).
 Therefore, in some countries policies made during an epidemic have competing costs.

### 2024-01-12 05_reinforcement_run.py

@@ -20,7 +20,9 @@ env.reset()
 # models/1704873042/5165.zip
 # models/1704874281/374740.zip
 # models/1704994098/2742.zip
-model_path = os.path.join(MODELS_DIR, "1704994098", "2742.zip")
+# models/1705049761/79518.zip
+# models/1705085840/16452.zip
+model_path = os.path.join(MODELS_DIR, "1705085840", "16452.zip")
 
 model = PPO.load(model_path, env=env)
 

### 2024-01-12 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -80,28 +80,28 @@ class SIREnvironment(gym.Env):
     def step(self, action):
         self.prev_actions.append(action)
         
-        diff = 0
+        diff_action = 0
         if action == 0:
             self.stringency_index = max(0, self.stringency_index - 10)
-            # diff = -10
+            diff_action = -10
         elif action == 1:
             self.stringency_index = max(0, self.stringency_index - 5)
-            # diff = -5
+            diff_action = -5
         elif action == 2:
             self.stringency_index = max(0, self.stringency_index - 2.5)
-            # diff = -2.5
+            diff_action = -2.5
         elif action == 3:
             self.stringency_index = max(0, self.stringency_index + 0)
-            # diff = 0
+            diff_action = 0
         elif action == 4:
             self.stringency_index = min(100, self.stringency_index + 2.5)
-            # diff = 2.5
+            diff_action = 2.5
         elif action == 5:
             self.stringency_index = min(100, self.stringency_index + 5)
-            # diff = 5
+            diff_action = 5
         elif action == 6:
             self.stringency_index = min(100, self.stringency_index + 10)
-            # diff = 10
+            diff_action = 10
         
         # each action is on ith day
         t_ith_day = np.linspace(0, self.ith_day, self.ith_day + 1)
@@ -127,7 +127,7 @@ class SIREnvironment(gym.Env):
         # REMEMBER: to change this definition of the reward in the render as well!!!
         # self.reward = self.normalized_GDP - (2 * self.r_eff)
 
-        reward_inertia = abs(diff)*-1*5
+        reward_inertia = abs(diff_action)*-1*5
         reward_r_eff = 10 if self.r_eff <= 1.9 else -10
         reward_I_percentage = -5000 if self.I_proportion >= 0.003 else 20
 
@@ -171,9 +171,19 @@ class SIREnvironment(gym.Env):
         return observation, self.reward, self.terminated, self.truncated, info
     
     def render(self, score=0.0, learning=False):
-
+        # N = df.loc[min(df.index), ["N"]].item()
+        # y0 = df.loc[min(df.index), ["S"]].item(), df.loc[min(df.index), ["I"]].item(), df.loc[min(df.index), ["R"]].item()
+        # days_difference = (max(df["date"]) - min(df["date"])).days
[…]

### 2024-01-13 05_reinforcement_run.py
fixing plotting

@@ -40,9 +40,9 @@ for episode in range(1, episodes+1):
         info_save.append(info)
         score += reward
     print(f'Episode: {episode}, Score: {score}')
-    env.render(score=score)
     with open(os.path.join(OUTPUT_DIR, "actions_taken", f"{score:.2f}.txt"), "w") as f:
         for action in actions_taken:
             f.write(str(action) +"\n")
     info_save_df = pd.DataFrame.from_dict(info_save)
     info_save_df.to_csv(os.path.join(OUTPUT_DIR, "info_save", f"{score:.2f}.csv"), index=False)
+    env.render(score=score, stringency=np.array(info_save_df['stringency_index']))

### 2024-01-13 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -170,121 +170,256 @@ class SIREnvironment(gym.Env):
             }
         return observation, self.reward, self.terminated, self.truncated, info
     
-    def render(self, score=0.0, learning=False):
-        # N = df.loc[min(df.index), ["N"]].item()
-        # y0 = df.loc[min(df.index), ["S"]].item(), df.loc[min(df.index), ["I"]].item(), df.loc[min(df.index), ["R"]].item()
-        # days_difference = (max(df["date"]) - min(df["date"])).days
-        # t = np.linspace(0, days_difference, days_difference + 1)
-        # nu_varying = list(df['nu_varying_with_time'])
+    def render(self, score=0.0, stringency=np.array([]), learning=False):
+        N = self.df.loc[min(df.index), ["N"]].item()
+        y0 = self.df.loc[min(df.index), ["S"]].item(), self.df.loc[min(df.index), ["I"]].item(), self.df.loc[min(df.index), ["R"]].item()
+        days_difference = (max(self.df["date"]) - min(self.df["date"])).days
+        t = np.linspace(0, days_difference, days_difference + 1)
+        nu_varying = list(self.df['nu_varying_with_time'])
+
+        moves_lockdown = stringency / 100
+        moves_ret = odeint(deriv, y0, t, args=(N, optimal_beta, optimal_gamma, nu_varying, moves_lockdown))
+        moves_S, moves_I, moves_R = moves_ret.T
         
-        # stringency_index_random_choice = []
-        # store_S = np.zeros(days_difference + 1)
-        # store_I = np.zeros(days_difference + 1)
-        # store_R = np.zeros(days_difference + 1)
-        # moves_lockdown = moves / 100
-        self.t = np.linspace(0, TOTAL_DAYS, TOTAL_DAYS + 1)
-        fig, axes = plt.subplots(nrows=4, ncols=2, figsize=(10, 12))
+        self.df["S_moves"] = moves_S
+        self.df["I_moves"] = moves_I
+        self.df["R_moves"] = moves_R
+        
         modelling_type = "with_lockdown_with_vaccination_time_varying_nu"
-        axes[0, 0].plot(self.t, self.df['S']/self.df['N'], 'b', alpha=0.5, lw=2, label='Susceptible (actual)')
[…]

### 2024-01-13 rough.py

@@ -1,0 +1,117 @@
+modelling_type = "with_lockdown_with_vaccination_time_varying_nu"
+
+output_path_img = os.path.join(OUTPUT_RL, str(score))
+try:
+    os.makedirs(output_path_img)
+except:
+    print("path exists")
+
+plt.figure(figsize=(12, 8))
+plt.plot(self.df['date'], self.df['S']/self.df['N'], 'b', alpha=0.5, lw=2, label='Susceptible (actual)')
+plt.plot(self.df['date'], self.df['I']/self.df['N'], 'r', alpha=0.5, lw=2, label='Infected (actual)')
+plt.plot(self.df['date'], self.df['R']/self.df['N'], 'g', alpha=0.5, lw=2, label='Recovered (actual)')
+plt.plot(self.df['date'], self.df['S_modelled_' + modelling_type]/self.N, 'b:', alpha=0.5, lw=2, label='Susceptible (modelled)')
+plt.plot(self.df['date'], self.df['I_modelled_' + modelling_type]/self.N, 'r:', alpha=0.5, lw=2, label='Infected (modelled)')
+plt.plot(self.df['date'], self.df['R_modelled_' + modelling_type]/self.N, 'g:', alpha=0.5, lw=2, label='Recovered (modelled)')
+plt.plot(self.df['date'], self.df['S_moves']/self.N, 'b--', alpha=0.5, lw=2, label='Susceptible (rl)')
+plt.plot(self.df['date'], self.df['I_moves']/self.N, 'r--', alpha=0.5, lw=2, label='Infected (rl)')
+plt.plot(self.df['date'], self.df['R_moves']/self.N, 'g--', alpha=0.5, lw=2, label='Recovered (rl)')
+plt.xlabel('Date')
+plt.ylabel('Percentage of Population')
+plt.title('SIR Epidemic Trajectory')
+plt.tick_params(length=0)
+plt.grid(True)
+plt.legend()
+plt.savefig(os.path.join(output_path_img, "rl_sir.png"))
+plt.savefig(os.path.join(output_path_img, "rl_sir.eps"))
+
+plt.figure(figsize=(12, 8))
+plt.plot(self.df['date'], self.df['I']/self.df['N'], 'r', alpha=0.5, lw=2, label='Infected (actual)')
+plt.plot(self.df['date'], self.df['I_modelled_' + modelling_type]/self.N, 'r:', alpha=0.5, lw=2, label='Infected (modelled)')
[…]

### 2024-01-15 05_reinforcement_run.py
beats modelled

@@ -22,8 +22,10 @@ env.reset()
 # models/1704994098/2742.zip
 # models/1705049761/79518.zip
 # models/1705085840/16452.zip
-model_path = os.path.join(MODELS_DIR, "1705085840", "16452.zip")
-
+# models/1705154668/548400.zi
+# models/1705305871/578562.zip
+# model_path = os.path.join(MODELS_DIR, "1705305871", "578562.zip")
+model_path = os.path.join(MODELS_DIR, "1705318802", "586788.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 10

### 2024-01-15 06_train_on_previous_model.py

@@ -1,0 +1,20 @@
+from rl_utils_new_complex_reward_sirvlockdown_nu_varying import *
+import time
+
+TIMESTEPS = 2742
+models_dir = f"models/{int(time.time())}/"
+if not os.path.exists(models_dir):
+	os.makedirs(models_dir)
+env = SIREnvironment()
+# model_path = os.path.join(MODELS_DIR, "1705154668", "548400.zip")
+# models/1705318802/586788.zip
+model_path = os.path.join(MODELS_DIR, "1705318802", "586788.zip")
+iters = 586788 / TIMESTEPS
+model = PPO.load(model_path)
+env.reset()
+model.set_env(env)
+while True:
+    iters += 1
+    print("ITER", iters)
+    model.learn(total_timesteps=TIMESTEPS, reset_num_timesteps=False, tb_log_name=f"PPO")
+    model.save(f"{models_dir}/{TIMESTEPS*iters}")

### 2024-01-15 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -192,198 +192,7 @@ class SIREnvironment(gym.Env):
         self.df["gdp_normalized_moves_min_max_normalized"] = ((fit_line_loaded(stringency) - MIN_GDP) / (MAX_GDP - MIN_GDP))
 
         self.t = np.linspace(0, TOTAL_DAYS, TOTAL_DAYS + 1)
-        # fig, axes = plt.subplots(nrows=4, ncols=2, figsize=(10, 12))
-        # modelling_type = "with_lockdown_with_vaccination_time_varying_nu"
-        # axes[0, 0].plot(self.df['date'], self.df['S']/self.df['N'], 'b', alpha=0.5, lw=2, label='Susceptible (actual)')
-        # axes[0, 0].plot(self.df['date'], self.df['I']/self.df['N'], 'r', alpha=0.5, lw=2, label='Infected (actual)')
-        # axes[0, 0].plot(self.df['date'], self.df['R']/self.df['N'], 'g', alpha=0.5, lw=2, label='Recovered (actual)')
-        # axes[0, 0].plot(self.df['date'], self.df['S_modelled_' + modelling_type]/self.N, 'b:', alpha=0.5, lw=2, label='Susceptible (modelled)')
-        # axes[0, 0].plot(self.df['date'], self.df['I_modelled_' + modelling_type]/self.N, 'r:', alpha=0.5, lw=2, label='Infected (modelled)')
-        # axes[0, 0].plot(self.df['date'], self.df['R_modelled_' + modelling_type]/self.N, 'g:', alpha=0.5, lw=2, label='Recovered (modelled)')
-        # axes[0, 0].plot(self.df['date'], self.df['S_moves']/self.N, 'b--', alpha=0.5, lw=2, label='Susceptible (rl)')
-        # axes[0, 0].plot(self.df['date'], self.df['I_moves']/self.N, 'r--', alpha=0.5, lw=2, label='Infected (rl)')
-        # axes[0, 0].plot(self.df['date'], self.df['R_moves']/self.N, 'g--', alpha=0.5, lw=2, label='Recovered (rl)')
-        # axes[0, 0].set_xlabel('Date')
-        # axes[0, 0].set_ylabel('Percentage of Population')
-        # axes[0, 0].set_title('SIR Epidemic Trajectory')
-        # axes[0, 0].tick_params(length=0)
-        # axes[0, 0].grid(True)
[…]

### 2024-01-15 rough.py

@@ -1,117 +1,86 @@
-modelling_type = "with_lockdown_with_vaccination_time_varying_nu"
+        plt.figure(figsize=(12, 8))
+        plt.plot(self.df['date'], self.df['S']/self.df['N'], color="#006EAE", alpha=0.5, lw=2, label='Susceptible (actual)')
+        plt.plot(self.df['date'], self.df['I']/self.df['N'], color="#C5373D", alpha=0.5, lw=2, label='Infected (actual)')
+        plt.plot(self.df['date'], self.df['R']/self.df['N'], color="#429130", alpha=0.5, lw=2, label='Recovered (actual)')
+        plt.plot(self.df['date'], self.df['S_modelled_' + modelling_type]/self.N, color="#006EAE", linestyle=':', alpha=0.5, lw=2, label='Susceptible (modelled)')
+        plt.plot(self.df['date'], self.df['I_modelled_' + modelling_type]/self.N, color="#C5373D", linestyle=':', alpha=0.5, lw=2, label='Infected (modelled)')
+        plt.plot(self.df['date'], self.df['R_modelled_' + modelling_type]/self.N, color="#429130", linestyle=':', alpha=0.5, lw=2, label='Recovered (modelled)')
+        plt.plot(self.df['date'], self.df['S_moves']/self.N, color="#006EAE", linestyle='--', alpha=0.5, lw=2, label='Susceptible (rl)')
+        plt.plot(self.df['date'], self.df['I_moves']/self.N, color="#C5373D", linestyle='--', alpha=0.5, lw=2, label='Infected (rl)')
+        plt.plot(self.df['date'], self.df['R_moves']/self.N, color="#429130", linestyle='--', alpha=0.5, lw=2, label='Recovered (rl)')
+        plt.xlabel('Date')
+        plt.ylabel('Percentage of Population')
+        plt.title('SIR Epidemic Trajectory')
+        plt.tick_params(length=0)
+        plt.grid(True)
+        plt.legend()
+        plt.savefig(os.path.join(output_path_img, "rl_sir.png"))
+        plt.savefig(os.path.join(output_path_img, "rl_sir.eps"))
 
-output_path_img = os.path.join(OUTPUT_RL, str(score))
-try:
-    os.makedirs(output_path_img)
-except:
-    print("path exists")
+        plt.figure(figsize=(12, 8))
[…]

### 2024-02-06 paper_draft/main.tex
beats modelled
changes in the paper

@@
+% to compile: https://tex.stackexchange.com/questions/140845/how-can-i-ignore-latex-error-while-compiling
+% xelatex -interaction=nonstopmode main.tex
 \documentclass[tikz,fleqn,12pt]{wlscirep}
@@
 \usepackage[autostyle, english = american]{csquotes}
-\MakeOuterQuote{"}
+\usepackage[nameinlink]{cleveref}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amsmath}
-\DeclareMathAlphabet{\altmathcal}{OMS}{cmsy}{m}{n}
+\usepackage{mathrsfs}
 \usepackage{mathptmx}
@@
 \usepackage{subcaption}
+\usepackage{nameref}
+\DeclareMathAlphabet{\altmathcal}{OMS}{cmsy}{m}{n}
+\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
 \usetikzlibrary{arrows.meta,positioning,fit}
+% https://texblog.org/2013/05/06/cleveref-a-clever-way-to-reference-in-latex/
+% https://tex.stackexchange.com/questions/433971/redefining-ref-such-that-it-is-eqref-when-you-use-refeq-and-then-it-is
+% https://tex.stackexchange.com/questions/18986/is-there-a-standard-way-to-refer-to-a-range-of-equations
+% \crefname{equation}{}{}
+% \creflabelformat{equation}{\textup{(#2#1#3)}}
+\newcommand{\crefrangeconjunction}{--}
+\MakeOuterQuote{"}
+
 \tikzset{%
@@
 The outbreak of COVID-19 has highlighted the intricate interplay between public health and economic stability on a global scale.
-This paper aims to develop a reinforcement learning framework to balance health and economic outcomes during infectious disease outbreaks.
-The framework utilizes the SIR model without vital dynamics and incorporates globally comparable government responses.
-The study acknowledges the limitations of deterministic models and proposes the use of deep reinforcement learning to reduce input dimensionality and normalize input.
-While the model offers transparency in terms of its dependency on the reward policy defined, it recognizes the need for a comprehensive consideration of decision factors beyond pure reinforcement learning results.
+This study proposes a novel reinforcement learning framework designed to optimize health and economic outcomes during pandemics.
+The framework leverages the SIR model, integrating both lockdown measures (via a stringency index) and vaccination strategies to simulate disease dynamics.
+The stringency index, indicative of the severity of lockdown measures, influences both the spread of the disease and the economic health of a country.
+Developing nations, which bear a disproportionate economic burden under stringent lockdowns, are the primary focus of our study.
+By implementing reinforcement learning, we aim to optimize governmental responses and strike a balance between the competing costs associated with public health and economic stability.
+This approach also enhances transparency in governmental decision-making by establishing a well-defined reward function for the reinforcement learning agent.
+In essence, this study introduces an innovative and ethical strategy to navigate the intricate challenge of balancing public health and economic stability amidst infectious disease outbreaks.
 \end{abstract}
@@
 \thispagestyle{empty}
-\subsection*{Introduction}
+\section{Introduction}
 In the past, global spread of infectious diseases was largely due to colonization, slavery, and war, leading to widespread illness and death from diseases like tuberculosis, polio, smallpox, and diphtheria.
@@
 However, in low and lower-middle income countries the burden of infectious diseases still persists.
-The rapid pace of urbanization in low and middle-income countries, along with the rise in populations living in crowded, poor-quality homes, has led to new conditions that favor the emergence of infectious diseases~\cite{Baker2022, inequitableworld}.
+The rapid pace of urbanization in low and middle-income countries, along with the rise in populations living in crowded, poor-quality homes, has led to new conditions that favor the emergence of infectious diseases~\cite{Baker2022,inequitableworld}.
 
-Recently, the COVID-19 pandemic caused a havoc worlwide.
+Recently, the COVID-19 pandemic caused a havoc worldwide.
 Till date there have been 772 million cases and more than 6 million deaths~\cite{WHOCovid}.
@@
 The need for a nuanced understanding of how interventions impact both health outcomes and economic indicators became increasingly evident, prompting a comprehensive examination by epidemiologists to assist policy makers~\cite{Anderson2020}.
-The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022, Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, stringency measures/non-pharmaceutical interventions (NPIs)~\cite{Jalloh2022, Caldwell2021, Ferguson2020}, and financial considerations~\cite{DeFoo2023}.
-Despite the numerous studies conducted, very few explore how common interventions meet multiple policy objectives or how a precise articulation of the main policy goals directs the selection of the most effective interventions in terms of health and economic results~\cite{Hollingsworth2011, Song2022, Pangallo2023, Ash2022, Ohi2020, PADMANABHAN2021102676}.
+The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022,Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, stringency measures/non-pharmaceutical interventions (NPIs)~\cite{Jalloh2022,Caldwell2021,Ferguson2020}, and financial considerations~\cite{DeFoo2023}.
+Despite the numerous studies conducted, very few explore how common interventions meet multiple policy objectives or how a precise articulation of the main policy goals directs the selection of the most effective interventions in terms of health and economic results~\cite{Hollingsworth2011, Song2022,Pangallo2023,Ash2022,Ohi2020, PADMANABHAN2021102676,NBERw26981,analytical_covid_lockdown_model}.
 The economic impact of the COVID-19 pandemic varied between rich and poor countries.
 Although COVID-19 deaths had a slightly larger negative effect on the Gross Domestic Product (GDP) in advanced economies, this difference was not statistically significant.
-However, lockdown restrictions were found to have a more damaging impact on economic activity in emerging and developing economies~\cite{Gagnon2023, Redlin2022,Liang2021}.
-It's also suggested that an increase in COVID-19 cases was associated with the introduction of harsher NPIs and lockdown measures could be relaxed once vaccination rates increase~\cite{Redlin2022, Patel2021}.
+However, lockdown restrictions were found to have a more damaging impact on economic activity in emerging and developing economies~\cite{Gagnon2023,Redlin2022,Liang2021}.
+It's also suggested that an increase in COVID-19 cases was associated with the introduction of harsher NPIs and lockdown measures could be relaxed once vaccination rates increase~\cite{Redlin2022,Patel2021}.
 
-Many economists have studied the effect of COVID-19 on the economy of nations~\cite{Gagnon2023, GagnonKorea, Deb2020,Eichenbaum2021}.
-In advanced economies like Korea, where the stringency index was below the median the recession was milder than other advanced economies like the United Kingdom where the stringency was much higher~\cite{GagnonKorea}, they achieved it mostly with very aggressive testing, contact tracing, and enforced quarantines and isolations~\cite{Lim2023, KoreaMinister}.
+Many economists have studied the effect of COVID-19 on the economy of nations~\cite{Gagnon2023,GagnonKorea,Deb2020,Eichenbaum2021}.
+In advanced economies like Korea, where the stringency index was below the median the recession was milder than other advanced economies like the United Kingdom where the stringency was much higher~\cite{GagnonKorea}, they achieved it mostly with very aggressive testing, contact tracing, and enforced quarantines and isolations~\cite{Lim2023,KoreaMinister}.
 In India, social distancing and containment measures have been effective in reducing the number of COVID-19 cases but have come with economic costs.
@@
 
-In this paper we want to optimize the government policies regarding stringency.
+In this paper we optimize the government policies regarding stringency.
 Therefore, alongside epidemiological data, we use the measures of globally comparable government responses~\cite{Hale2021}.
-We use the simple SIR model without vital dynamics~\cite{Hethcote1989, Hethcote2008, ALLEN2017128} as it is assumed that the time scale is short enough so that can be neglected~\cite{Cooper2020}.
+We use the simple SIR model without vital dynamics~\cite{Hethcote1989, Hethcote2008, ALLEN2017128} as it is assumed that the timescale is small enough that it can be neglected~\cite{Cooper2020}.
 By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we can effectively address the real-world conditions and propose a solution that is both effective and extendable.
-The current model accounts for the recovery reached through vaccination~\cite{Marinov2022, MaurciodeCarvalho2023, Thater, Turkyilmazoglu2022, YALADANDA2022101052}.
-However, the study has limitations; the deterministic SIR model fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}.
+The current model (SIRV with lockdown) accounts for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{SIRLockdown,NBERw26981,NBERw26867}.
+Traditional SIR model neglects the time-varying property of parameters like $\beta$ and $\gamma$, but in this paper since, lockdown is a parameter of the proposed model, and it changes with time therefore, we propose a time-dependent SIR model\cite{TimeDependentSIR}.
+However, the study has limitations; First, the deterministic SIR model (predecessor to our proposed model) fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}.
+Second, the underreporting of cases during the period selected by our study.
+Lastly, the reinforcement learning agent should be resistant to how the vaccination rate changes -- or to when it is first introduced because then it scopes the environment for succumbing to wishful thinking which can be  potentially dangerous.
 
-In order to capture competing costs within the environment and achieve a balance between health and economic outcomes, we intend to employ reinforcement learning~\cite{Nguyen2022, Bastani2021, Song2022, Ohi2020, PADMANABHAN2021102676}.
-Not only does the formulation of the model deal better with competing costs, but it also offers a more transparency behind the reasoning of the decisions being made in such circumstances.
-When we conceptualize our problem as a reinforcement learning task, an agent is tasked with making decisions in an environment with the aim of optimizing cumulative rewards.
-This approach relies on trial and error, in contrast to dynamic programming, which assumes complete knowledge of the environment~\cite{Lavet2018}.
-We make the use of deep reinforcement learning, which is an advancement to reinforcement learning as it allows to reduce the input dimensionality and normalize the input~\cite{Lavet2018, Arulkumaran2017, Henderson_Islam_Bachman_Pineau_Precup_Meger_2018, Mnih2015}.
-While our model is more transparent in terms of its dependency on the reward policy defined, it has its limitations.
-A universal optimal policy may not suit diverse socioeconomic contexts due to variations in healthcare resources and economic vulnerabilities across countries, regions, or cities and a comprehensive consideration of decision factors, extending beyond pure reinforcement learning results is needed~\cite{Song2022, Dunn2017, Demir2006}.
+After modelling the disease with lockdown (via stringency index) and vaccination, we try to understand the effects of lockdown on the GDP.
+Therefore, decisions made by the government regarding the level of lockdown to be enforced plays a role on both the public health outcomes and economic stability during a pandemic.
+On one hand, stringent lockdown measures can effectively slow the spread of the disease, thereby improving public health outcomes.
+However, these measures often come at the cost of significant economic disruption, leading to job losses, business closures, and reduced economic growth.
+On the other hand, relaxing lockdown measures may help to mitigate the economic impact of the pandemic, but could result in increased disease transmission and worsened public health outcomes.
+In order to capture competing costs within the environment and achieve a balance between health and economic outcomes, we intend to employ reinforcement learning~\cite{Nguyen2022,Bastani2021,Song2022,Ohi2020,PADMANABHAN2021102676}.
+Not only does the formulation of the model deal better with competing costs, but it also offers more transparency behind the reasoning of the decisions being made in such circumstances.
+When we conceptualize our problem as a reinforcement learning task, an agent is tasked with making decisions in an environment with the aim of optimizing cumulative rewards (i.e., the total amount of reward it receives over the long run).
+Simply put, given discrete time steps $t = 0, 1, 2, 3, \dots$, at each time step the agent receives a representation of the environment's state, $s_t \in \altmathcal{S}$, and selects an action $a_t \in \altmathcal{A}(s_t)$, where $\altmathcal{A}(s_t)$ is the set of actions available in state $s_t$, and one step later receives a reward $r_{t+1} \in \altmathcal{R}$ and the state is updated~\cite{sutton2018reinforcement}.
+The way we define the way these rewards that are given to the agent, makes this decision process more transparent, however, it has its limitations.
+A universal optimal policy may not suit diverse socioeconomic contexts due to variations in healthcare resources and economic vulnerabilities across countries, regions, or cities and a comprehensive consideration of decision factors, extending beyond pure reinforcement learning results is needed~\cite{Song2022,Dunn2017,Demir2006}.
 
-\subsection*{Mathematical Formulation}
-In SIR-RL, we use a compartmental model to model infectious disease.
+Additionally, since most modern reinforcement learning achievements are due to a combination of deep learning~\cite{Mnih2015}, in the following framework we make the use of this.
+Deep reinforcement learning is an advancement to reinforcement learning which helps normalize the input and reduce its dimensionality~\cite{Lavet2018,Arulkumaran2017,Henderson_Islam_Bachman_Pineau_Precup_Meger_2018,Mnih2015}.
+We use a Long Short Term Memory recurrent neural network for time-series data~\cite{NIPS2001_a38b1617,LSTM} and a simple fully connected network for the non-time-series data.
+
+In summary, by using reinforcement learning augmented with deep learning techniques for the SIRV with lockdown environment, we can better understand the effects of lockdown measures on both public health outcomes and economic stability during a pandemic.
+However, it is crucial to consider the limitations of this approach and take into account a comprehensive set of decision factors in order to make informed policy decisions that are tailored to specific socioeconomic contexts.
+
+\section{Mathematical Formulation and Numerical Computation}
+In SIRV-RL, we use a compartmental model to model infectious disease environment.
 We iteratively develop this model to fit the actual data better.
@@
 
+\subsection{Simple SIR Model}
 \begin{figure}[H]
@@
 
-Here, 
-$c$ is the average number of contacts a susceptible person makes per day.
+Here, $c$ is the average number of contacts a susceptible person makes per day.
 $p$ is the probability of the susceptible person becomes infectious after coming into contact with an infectious person.
@@
 
-At the peak of an epidemic $R_0$ declines as there are no more susceptible people left in the pool, therefore, $R_{eff}$ (effective reproductive number) comes into play.
-$R_{eff}$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic - therefore, it takes into account the existing immunity of the system.
+At the peak of an epidemic $R_0$ declines as there are no more susceptible people left in the pool, therefore, $R_e$ (effective reproductive number) comes into play.
+$R_e$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic, therefore, it takes into account the existing immunity of the system.
 
 \begin{equation}
-  R_{eff} = R_0 \frac{S}{N}
+  R_e = R_0 \frac{S}{N}
   \label{eq:R_effective}
@@
 $S$ is the number of susceptible people, $N$ is the total population.
-At the start of an epidemic when everyone is susceptible, $R_{eff} = R_0$ as, $S = N$ (i.e.
-the whole population is susceptible).
+At the start of an epidemic when everyone is susceptible, $R_e = R_0$ as, $S = N$ (i.e., the whole population is susceptible).
 $\beta$ and $\gamma$ are also used to define probability of and infectious individual infecting another individual $\beta / (\beta + \gamma)$ and the probability of recovery, $\gamma / (\beta + \gamma)$.
 
-Most government policies look at the value of $R_{eff}$ to come us with an effective strategy to combat the disease as the fate of the evolution of the disease depends upon it.
-When $R_{eff}$ is less than one, the infected population $I$ will steadily decline to zero.
-Conversely, if $R_{eff}$ is greater than one, the infected population will increase.
-In other words, when $\frac{dI(t)}{dt} < 0 \Rightarrow R_{eff} < 1$ and $\frac{dI(t)}{dt} > 0 \Rightarrow R_{eff} > 1},$ the effective reproductive rate $R_{eff}$ serves as a critical threshold that determines whether an infectious disease will rapidly extinguish or escalate into an epidemic~\cite{Cooper2020}.
+Most government policies look at the value of $R_e$ to come us with an effective strategy to combat the disease as the fate of the evolution of the disease depends upon it.
+When $R_e$ is less than one, the infected population $I$ will steadily decline to zero.
+Conversely, if $R_e$ is greater than one, the infected population will increase.
+In other words, when $\frac{dI(t)}{dt} < 0 \Rightarrow R_e < 1$ and $\frac{dI(t)}{dt} > 0 \Rightarrow R_e > 1$, therefore, the effective reproductive rate $R_e$ serves as a critical threshold that determines whether an infectious disease will rapidly extinguish or escalate into an epidemic~\cite{Cooper2020}.
 
-To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define the cost to calibrate the model using the huber loss~\cite{huberloss}:
+To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define two cost functions \cref{eq:cost_SIR,eq:cost_I} to calibrate the model with the use of huber loss~\cite{huberloss}.
+We generally calibrate the model using \cref{eq:cost_SIR}.
+However, sometimes we must strike a balance between modeling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}).
 
@@
     \frac{1}{2}{(y - f(x))^2} & \text{for } |y - f(x)| \leq \delta, \\
-    \delta \cdot (|y - f(x)| - \frac{1}{2}\delta) & \text{otherwise.}
+    \delta \cdot (|y - f(x)| - \frac{1}{2}\delta) & \text{otherwise.} \\
+  \end{cases}
   \label{eq:huberloss}
-  \end{cases}
+\end{equation}
+
+In the above equation, $y$ is the actual data and $f(x)$ is the prediction.
+
+\begin{equation}
+  \begin{split}
+  \textrm{loss\_SIR} = \textrm{cost\_function\_SIR}(S, \hat{S}, I, \hat{I}, R, \hat{R}) \\
+  = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
+  \end{split}
+  \label{eq:cost_SIR}
 \end{equation}
 \begin{equation}
-  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
-  \label{eq:cost_SIR_model}
+  \textrm{loss\_I} = \textrm{cost\_function\_I}(I, \hat{I}) = L_{\delta = 1}(I, \hat{I})
+  \label{eq:cost_I}
 \end{equation}
@@
 Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered.
-Minimizing this cost function using the Nelder-Mead method\cite{Gao2010}, we get:
+Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown} , we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data.
+The following parameters and loss is obtained:
 
 \begin{equation}
-  \beta_{optimal} = 0.04208479828695971
+  \beta_{optimal} = 0.042
   \label{eq:beta_optimal_without_lockdown}
@@
 \begin{equation}
-  \gamma_{optimal} = 0.02388356686032017
+  \gamma_{optimal} = 0.024
   \label{eq:gamma_optimal_without_lockdown}
@@
 \begin{equation}
-  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.7620817917644795
+  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.762
   \label{eq:r0_without_lockdown}
@@
 \begin{equation}
-  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 85051490.53250012
-  \label{eq:cost_without_lockdown}
+  \textrm{loss\_SIR} = 85051490.533
+  \label{eq:cost_SIR_without_lockdown}
 \end{equation}
-See figure \ref{fig:SIR_model_IND} to see how the model compares with the real data.
+\begin{equation}
+  \textrm{loss\_I} = 45187665.281
+  \label{eq:cost_I_without_lockdown}
+\end{equation}
 
-Now that, we have a model -- we want to say $\beta$ is a time-varying parameter controlled by the stringency index, $s$~\cite{SIRLockdown,SIRLockdown2,analytical_covid_lockdown_model}.
+See \cref{fig:SIR_model_IND} to see how the model compares with the actual data.
+
+\subsection{SIR Model with Lockdown}
+Now that, a simple SIR model has been established -- we need to model the effects of the stringency index (measure for the strictness of lockdown) on $\beta$ (the effective transmission rate).
+To do this, we say the flow of susceptibles not only depend on $\beta$ but also $s(t)$ the stringency index at time~\cite{NBERw26867,AlvarezLockdownSIR,SIRLockdown,SIRLockdown2,analytical_covid_lockdown_model}.
 The stringency index is a composite measure based on nine response indicators including school closures, workplace closures, and travel bans, rescaled to a value from 0 to 100 (100 = strictest)~\cite{owidcoronavirus}.
@@
 
-To define the new time-varying beta that is dependent on the current stringency index, we say,
+To define the new time-varying beta that is dependent on the current stringency index, the following equations have been formulated:
 
 \begin{equation}
-  \frac{dS}{dt} = -\beta (1 - (s(t)/100)) \frac{S I}{N}
+  \frac{dS}{dt} = -\beta (1 - s(t)/100) \frac{S I}{N}
   \label{eq:S_with_lockdown}
@@
 \begin{equation}
-  \frac{dI}{dt} = \beta (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
+  \frac{dI}{dt} = \beta (1 - s(t)/100) \frac{S I}{N} - \gamma I
   \label{eq:I_with_lockdown}
@@
 
-Where, $s(t)$ is the stringency index at time $t$ and is scaled down by a factor of 100 to normalize it and bring it in the range $[0, 1]$. Optimizing this using the Nelder-Mead method we get:
+Where, $s(t)$ is the stringency index at time $t$ and is scaled down by a factor of 100 to normalize it and bring it in the range $s(t)/100 \in [0, 1]$.
+Multiplying the rate of flow from $S$ to $I$ compartment with $1 - s(t)/100$ allows us to account for the effect that stringency has on the disease progression.
+A higher stringency index can theoretically, stop the flow from the susceptible population to the infected population entirely.
+Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method, we get the following parameters and loss:
 
 \begin{equation}
-  \beta_{optimal} = 0.4013340889432941
+  \beta_{optimal} = 0.401
   \label{eq:beta_optimal_with_lockdown}
@@
 \begin{equation}
-  \gamma_{optimal} = 0.09017476605499258
+  \gamma_{optimal} = 0.090
   \label{eq:gamma_optimal_with_lockdown}
@@
 \begin{equation}
-  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.450625230328213
+  \begin{split}
+    \overbar{R_0} = 1.693 \quad \text{(Mean)} \\
+    \widetilde{R_0} = 1.624 \quad \text{(Median)} \\
+    \text{Mode}(R_0) = 0.804 \quad \text{(Mode)} \\
+    \sigma_{R_0} = 0.786 \quad \text{(Standard Deviation)} \\
+    R_0 \in [0.16467, 3.0497] \quad \text{(Range)}
+  \end{split}
   \label{eq:r0_with_lockdown}
@@
 \begin{equation}
-  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 98438821.45587364
-  \label{eq:cost_with_lockdown}
+  \textrm{loss\_SIR} = 98438821.456
+  \label{eq:cost_SIR_with_lockdown}
+\end{equation}
+\begin{equation}
+  \textrm{loss\_I} = 11345389.686
+  \label{eq:cost_I_with_lockdown}
 \end{equation}
 
-Finally, an additional flow from the susceptible population to the can be shown by adding a vaccination rate $\nu$ in the model.
+\subsection{SIRV Model with Lockdown}
+Lastly, an additional flow from the susceptible population to the recovered population can be shown by adding a vaccination rate $\nu$ in the model.
 
 \begin{equation}
-  \frac{dS}{dt} = -\beta (1 - (s(t)/100)) \frac{S I}{N} - \nu S
+  \frac{dS}{dt} = -\beta (1 - s(t)/100) \frac{S I}{N} - \nu S
   \label{eq:S_with_lockdown_and_nu}
@@
 \begin{equation}
-  \frac{dI}{dt} = \beta (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
+  \frac{dI}{dt} = \beta (1 - s(t)/100) \frac{S I}{N} - \gamma I
   \label{eq:I_with_lockdown_and_nu}
@@
 
-Optimizing these equations with the Nelder-Mead method we get:
+Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method:
 
 \begin{equation}
-  \beta_{optimal} = 0.40897034072952304
+  \beta_{optimal} = 0.409
   \label{eq:beta_optimal_with_lockdown_and_nu}
@@
 \begin{equation}
-  \gamma_{optimal} = 0.09196829370123338
+  \gamma_{optimal} = 0.092
   \label{eq:gamma_optimal_with_lockdown_and_nu}
@@
 \begin{equation}
-  \nu_{optimal} = 2.9044029843851936e-05
+  \nu_{optimal} = 2.904 \times 10^{-5}
   \label{eq:nu_optimal_with_lockdown_and_nu}
@@
 \begin{equation}
-  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.446862329077204
+  \begin{split}
+    \overbar{R_0} = 1.691 \\
+    \widetilde{R_0} = 1.623 \\
+    \text{Mode}(R_0) = 0.803 \\
+    \sigma_{R_0} = 0.785 \\
+    R_0 \in [0.165, 3.047]
+  \begin{split}
   \label{eq:r0_with_lockdown_and_nu}
@@
 \begin{equation}
-  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 94636860.38436058
-  \label{eq:cost_with_lockdown_and_nu}
+  \textrm{loss\_SIR} = 94636860.384
+  \label{eq:cost_SIR_with_lockdown_and_nu}
+\end{equation}
+\begin{equation}
+  \textrm{loss\_SIR} = 10840360.995
+  \label{eq:cost_I_with_lockdown_and_nu}
 \end{equation}
 
-However, as observed by the value of $\nu_{optimal}$ from \ref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible (close to zero) and the overestimation of infected individuals in \ref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time.
-This suggests a time-varying vaccination rate, wherein the transition from susceptibility to direct recovery fluctuates with time.
-Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \ref{eq:beta_optimal_with_lockdown_and_nu} and \ref{eq:gamma_optimal_with_lockdown_and_nu}, we optimize the value of $\nu$ given a sub-interval of $[{start}, {start} + {window}]$ where the ${window} = 20$ \cite{Marinov2022}.
-Here, ${window}$ is a hyperparameter that can be changed so as to reduce the cost function.
+\subsection{Optimizing Window Length for Time-varying $\nu$}\label{optimizing_window_length}
+However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time.
+This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}.
+Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} i.e., the window length that result in the least loss from \cref{eq:cost_SIR}.
+Using different window lengths to estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$, we observe the following.
+It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
 
 \begin{equation}
-  \frac{dS}{dt} = -\beta_{optimal}  (1 - (s(t)/100))  \frac{S I}{N} - \nu S
+  \frac{dS}{dt} = -\beta_{optimal}  (1 - s(t)/100)  \frac{S I}{N} - \nu S
   \label{eq:S_with_lockdown_and_nu_calc_nu}
@@
 \begin{equation}
-  \frac{dI}{dt} = \beta_{optimal}  (1 - (s(t)/100)) \frac{S I}{N} - \gamma_{optimal} I
+  \frac{dI}{dt} = \beta_{optimal}  (1 - s(t)/100) \frac{S I}{N} - \gamma_{optimal} I
   \label{eq:I_with_lockdown_and_nu_calc_nu}
@@
 
-This model gives us how $\nu$ varies with time.
+The main difference here is instead of supplying these equations with the entire time series, only data lying within the sub-interval of $[{start}, {start} + {window\_length}]$ is given, then in the next step data of the next sub-inteval is provided to optimize for, i.e., $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$.
+This model gives shows how $\nu$ varies with time.
+Since, the window length acts like a hyperparameter, we can try different window lengths in order to evaluate which would give us the best results.
+Therefore, the below graphs show us the loss using \cref{eq:cost_SIR,eq:cost_I}.
+
+\begin{figure}
+  \caption{Loss for Different Window Lengths}
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/window_length_loss_SIR_IND.pdf}
+    \caption{Loss for Different Window Lengths for Susceptible, Infected and Recovered Population}
+  \end{subfigure}
+  \label{fig:window_length_loss_SIR_IND}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/window_length_loss_I_IND.pdf}
+    \caption{Loss for Different Window Lengths for Infected Population}
+  \end{subfigure}
+\label{fig:window_length_loss_I_IND}
+\end{figure}
+
+From the following graph, although the window length of $10$ is best for all three groups of the population but this window length leads to a bad approximation for the infected population therefore, we choose a window length of $15$.
 
@@
   \caption{$\nu$ Varying with Time}
-  \includegraphics[width=\linewidth]{images/nu_varying_with_time_IND.eps}
-  \label{fig:nu_varying_with_time_IND}
+  \includegraphics[width=\linewidth]{images/interpolated_nu_varying_with_time_IND.pdf}
+  \label{fig:interpolated_nu_varying_with_time_IND}
 \end{figure}
 
-Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations below:
+This shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India \cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
+Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIRV Model with time varying $\nu$.
 
+\subsection{SIRV Model with Lockdown and time-varying $nu$}
+Lastly, we present our model using the following set of equations:
 \begin{figure}[H]
@@
 \begin{equation}
-  \frac{dS}{dt} = -\beta  (1 - (s(t)/100))  \frac{S I}{N} - \nu(t) S
+  \frac{dS}{dt} = -\beta  (1 - s(t)/100)  \frac{S I}{N} - \nu(t) S
   \label{eq:S_with_lockdown_and_time_varying_nu}
@@
 \begin{equation}
-  \frac{dI}{dt} = \beta  (1 - (s(t)/100)) \frac{S I}{N} - \gamma I
+  \frac{dI}{dt} = \beta  (1 - s(t)/100) \frac{S I}{N} - \gamma I
   \label{eq:I_with_lockdown_and_time_varying_nu}
@@
 \begin{equation}
-  \beta_{optimal} = 0.5129414911377119
+  \beta_{optimal} = 0.463
   \label{eq:beta_optimal_with_lockdown_and_time_varying_nu}
@@
 \begin{equation}
-  \gamma_{optimal} = 0.12716245167626106
+  \gamma_{optimal} = 0.114
   \label{eq:gamma_optimal_with_lockdown_and_time_varying_nu}
@@
 \begin{equation}
-  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 4.033749620081199
+  \begin{split}
+    \overbar{R_0} = 1.546 \\
+    \widetilde{R_0} = 1.483 \\
+    \text{Mode}(R_0) = 0.734 \\
+    \sigma_{R_0} = 0.718 \\
+    R_0 \in [0.150, 2.785]
+  \end{split}
   \label{eq:r0_with_lockdown_and_time_varying_nu}
@@
 \begin{equation}
-  cost = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R}) = 29742589.698910963
-  \label{eq:cost_with_lockdown_and_time_varying_nu}
+  \textrm{loss\_SIR} = 29116762.926
+  \label{eq:cost_SIR_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  \textrm{loss\_I} = 658537.443
+  \label{eq:cost_I_with_lockdown_and_time_varying_nu}
 \end{equation}
 
-See (\ref{fig:SIR_model_IND}, ~\ref{fig:comparing_costs_IND}) to see how the different models compare against each other.
+See \Crefrange{fig:SIR_model_IND,fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \cref{fig:comparing_costs_IND} to see how the different models compare against each other.
 
-Now, that we have set up the following relation between $\beta$ and $s(t)$, we investigate how stringency index affects the normalized Gross domestic product (GDP).
-To do this we fit a polynomial equation of the third degree to the points $(x, y)$ where $x$ is the $s(t)$ stringency at time t, and $y$ the normalized GDP and minimize the squared error.
-For India after fitting a 3 degree polynomial we get the following equation:
+Now, that a relation between $\beta$ and $s(t)$ is set up, it must be investigated how stringency index affects the normalized Gross domestic product (GDP)~\cite{OECDNormalizedGDP,OECDNormalizedGDP2}.
+To do this a polynomial equation of the third degree is fitted to the data points $(x, y)$ where $x$ is the $s(t)$ stringency at time $t$, and $y$ the normalized GDP and minimize the squared error.
+For India after fitting a 3 degree polynomial, the following equation is obtained:
 
@@
     \begin{split}
-      GDP = -5.96640236e-05 s^{3} + 6.65064332e-03 s^{2} -2.23109924e-01 s^{1} \\
-    + 1.01357226e+02
+      GDP = -5.96640236 \times 10^{-5} s^{3} + 6.65064332 \times 10^{-3} s^{2} - 2.23109924 \times 10^{-1} s^{1} \\
+    + 1.01357226 \times 10^{2}
     \end{split}
@@
 This environment is formally as a Markov decision process, and can be described as follows:
+
 \begin{itemize}
     % \item \url{https://jmlr.org/papers/volume22/20-1364/20-1364.pdf} RL Paper that references \url{https://araffin.github.io/slides/rl-tuto-jnrr19/#/13} RL ppt (for our RL notes).
-    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP ($(GDP_{predicted} - GDP_{min}) / (GDP_{max} - GDP_{min})$), $R_{eff}$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible, infected and recovered.
+    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP ($(GDP_{predicted} - GDP_{min}) / (GDP_{max} - GDP_{min})$), $R_e$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible, infected and recovered.
 The starting states are simply these values at the starting date and no previous actions.
-    \item TODO: FIX BELOW ITEM AFTER MODEL RUNS
-    \item Actions $\altmathcal{A}$: The stringency index variable was analyzed with a sample size of 1034.
-The mean value was approximately 59.62, with a standard deviation of 22.82.
-The minimum value was 0, while the maximum value reached 100.
-And the differences between two consecutive stringencies had a mean of 0.030474, and standard deviation of 2.387331, with the minimum being -14.360000 and maximum 55.560000.
+    \item Actions $\altmathcal{A}$: The stringency index variable was analyzed with a sample size of 915.
+The mean value was approximately $61.965049$, with a standard deviation of $17.669831$.
+The minimum value was $31.480000$, while the maximum value reached $96.300000$.
+And the differences between two consecutive stringencies had a mean of $-0.070919$, and standard deviation of $1.427145$, with the minimum being $-14.360000$ and maximum $16.670000$.
 Based on this we define the discrete action space.
@@
 \end{itemize}
+
 In general, the policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$.
@@
 The goal of RL is to find an optimal policy, $\pi^*$, which achieves the maximum expected return from all states.
+To achieve this, reinforcement learning start with an initial arbitrary policy, i.e., a $Q$-table with no entries.
+$Q$-table is a mapping from states $s_k \in \mathscr{S}$ to a predefined set of actions to increase or decrease the stringency at time t, which are the actions $a_k \in \mathscr{A}$.
+Each entry of the $Q$-table $\left(Q_k\left(s_k, a_k\right)\right)$ associates an action in the finite sequence $\left(\mathscr{A}_j\right)_{j \in \mathbb{J}^{+}}$ to a state of the finite sequence $\left(\mathscr{S}_i\right)_{i \in \mathbb{I}^{+}}$ \cite{sutton2018reinforcement}.
+In this case of epidemic control by NPI based strategies this policy represents the series of stringencies to be imposed upon the population to shift the initial status of the environment to a targeted status which is equivalent to the desired set of system states.
+This is how the $Q$-table updates saying, if in state $s_k$ the most ideal action is $a_k$.
+After having more and more experience with the environment and understanding which actions lead to a higher reward $r$ an optimal policy is derived be maximizing the expected value of discounted reward $J\left(r_k\right)=\mathbb{E}\left[\sum_{k=1}^{\infty} \gamma^{(k-1)} r_k\right]$, where, discount factor $\gamma \in [0, 1]$ and time steps $k = 1, 2, \dots$.
 
@@
 Conversely, it must be noted that herd immunity can only be achieved when the epidemic reaches its peak i.e.
-when the effective reproductive number is equal to one ($R_{eff} = 1$).
+when the effective reproductive number is equal to one ($R_e = 1$).
 This can only happen by lowering the stringency index which would allow the natural dynamics of the epidemic to transpire such that the population of susceptible individuals has depleted enough such that it is insufficient to propagate the disease further.
@@
 
+% TODO add: http://www.incompleteideas.net/book/ebook/node29.html
+
 The reward function is parameterized to account for key factors influencing decision-making.
-To incentivize reduction of $R{eff}$ (effective reproductive number) a negative reward is imposed of $-20 * R_{eff}$, but as the $R_{eff}$ is between $[1.9, 1,5]$ we start to positively reward it with a multiple of the GDP ($100 * \textrm{min\_max\_normalized\_GDP}$) and if the $R_{eff}$ is below $1.5$ then $200 * \textrm{min\_max\_normalized\_GDP}$.
-Furthermore, there's a simple positive reward of $10$ if $R_{eff}$ is below the threshold value ($1.9$) and negative reward of $-10$ otherwise.
-To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| * -1 * 5$) [TODO: this factor should be reduced back to 2 from 5].
+To incentivize reduction of $R{eff}$ (effective reproductive number) a negative reward is imposed of $-20 * R_e$, but as the $R_e$ is between $[1.9, 1,5]$ we start to positively reward it with a multiple of the GDP ($100 * \textrm{min\_max\_normalized\_GDP}$) and if the $R_e$ is below $1.5$ then $200 * \textrm{min\_max\_normalized\_GDP}$.
+Furthermore, there's a simple positive reward of $10$ if $R_e$ is below the threshold value ($1.9$) and negative reward of $-10$ otherwise.
+To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| * -1 * 5$).
 If the proportion of the infected population were to rise above 0.003 the model is punished ($-5000$) and otherwise rewarded ($+20$).
+As a reward for herd immunity, another reward is defined using the proportion of the infected population.
+If it lies between $[0.0005, 0.003]$ a reward of 2000 is given - else $0$.
+
 It should be realized there can be an infinite number of ways to design the reward function to be more human and upgrade the way a decision is taken given the situation \cite{AWSDeepRacer}.
@@
 This can be done using MultiInputPolicy, which by default uses the CombinedExtractor features extractor to turn multiple inputs into a single vector, handled by the network arch network.
+For data that varies with time (stringency, normalized GDP, $R_e$) we use a simple LSTM architecture \cite{LSTM}.
+For other data like, the current proportion of the population that is susceptible, infected, and recovered we use a simple fully connected layer.
+The output from both these networks are concatenated and used by the reinforcement learning agent to train on.
 
-For data that varies with time we use a simple LSTM architecture.
-
-% we utilized a custom combined extractor, a unique type of Long Short-Term Memory (LSTM) network, for processing our data.
-This network was designed to handle different types of input data encapsulated in an observation space, which is a common concept in reinforcement learning representing the information an agent observes from the environment.
-The CustomCombinedExtractor class, which inherits from the BaseFeaturesExtractor class, was initialized with an observation space that is a dictionary of Gym spaces.
-The Gym spaces module provides a set of high-level, easy-to-use data structures for defining observation and action spaces, which are essential components in reinforcement learning.
-The extractor contains a dictionary of individual extractors for each key in the observation space.
-If the subspace shape is equal to the total number of days plus one, an LSTM extractor is created with an input size of 1 and a hidden size of 16.
-This LSTM extractor is designed to handle sequential data, which is particularly useful for time-series data such as ours.
-For the 'other_stats' key, a Flatten extractor is used to flatten the input data.
-In the forward method, the observations are processed through their respective extractors.
-For LSTM extractors, the final hidden state is used as the output.
-All the processed observations are then concatenated into a single tensor.
-The observation space for the deep learning type is defined as a dictionary of four keys: 'stringency', 'normalized_gdp', 'r_eff', and 'other_stats'.
-Each key is associated with a Box space, which is a multi-dimensional box that allows for a range of numerical values.
-The shape of the Box space defines the dimensionality of the data, and the low and high parameters define the range of possible values.
-This LSTM network and the associated observation space provide a robust framework for processing and learning from our complex, multi-dimensional data.
+We train the model for $2742$ time steps for $200$ iterations and the best results are presented below.
 
@@
 
-\subsection*{Results}
+\section{Results}
 % https://www.nature.com/articles/s41598-020-79147-8 (see this to write results)
-Using the simple SIR model (\ref{eq:S_without_lockdown}, ~\ref{eq:cost_without_lockdown}) to model the disease dynamics we get:
+Using the simple SIR model from \Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown},  to model the disease dynamics we get:
 \begin{figure}[H]
@@
     \centering
-    \includegraphics[scale=0.5]{images/SIR_model_IND.eps}
+    \includegraphics[scale=0.5]{images/SIR_model_IND.pdf}
     \caption{SIR model with lockdown}
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_infections_IND.eps}
+    \includegraphics[scale=0.50]{images/SIR_model_infections_IND.pdf}
     \caption{Infections modelled with SIR model}
@@
 
-Combining the lockdown dynamics in the SIR model (\ref{eq:S_with_lockdown}, ~\ref{eq:cost_with_lockdown}) we get the following:
+Combining the lockdown dynamics in the SIR model using \Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}, we get the following:
 \begin{figure}[H]
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_IND.eps}
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_IND.pdf}
     \caption{SIR model with lockdown}
@@
     \centering
-    \includegraphics[scale=0.50]{images/stringency_varying_with_time_IND.eps}
+    \includegraphics[scale=0.50]{images/stringency_varying_with_time_IND.pdf}
     \caption{Stringency varying with Time}
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_infections_IND.eps}
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_infections_IND.pdf}
     \caption{Infections modelled with SIR model with Lockdown}
@@
 
-Combining vaccination dynamics with SIR with lockdown model (\ref{eq:S_with_lockdown_and_nu}, ~\ref{eq:cost_with_lockdown_and_nu}) we get the following:
+Combining vaccination dynamics with SIR with lockdown model using \Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}, we get the following:
 \begin{figure}[H]
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_IND.eps}
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_IND.pdf}
     \caption{SIRV model with lockdown}
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_IND.eps}
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_IND.pdf}
     \caption{Infections modelled with SIRV model with Lockdown}
@@
 \end{figure}
-Here, because the value of $\nu$ (\ref{eq:nu_optimal_with_lockdown_and_nu}) is negligible, it makes no change. However, time-varying $\nu$ shall be able to better account for the these dynamics.
+Here, because the value of $\nu$ \cref{eq:nu_optimal_with_lockdown_and_nu} is negligible, it makes no change. However, time-varying $\nu$ shall be able to better account for the these dynamics.
 
-For time-varying $\nu$ and SIRV model with lockdown (\ref{eq:S_with_lockdown_and_time_varying_nu}, ~\ref{eq:cost_with_lockdown_and_time_varying_nu}), we get the folllowing:
+For time-varying $\nu$ and SIRV model with lockdown from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following:
 \begin{figure}[H]
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_time_varying_nu_IND.eps}
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_time_varying_nu_IND.pdf}
     \caption{SIRV model with lockdown and time-varying $\nu$}
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_time_varying_nu_IND.eps}
+    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_time_varying_nu_IND.pdf}
     \caption{Infections modelled with SIRV model with Lockdown and time-varying $\nu$}
@@
   \caption{Loss for Different Models}
-  \includegraphics[width=\linewidth]{images/comparing_costs_IND.eps}
+  \includegraphics[width=\linewidth]{images/comparing_costs_IND.pdf}
   \label{fig:comparing_costs_IND}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_IND.eps}
+    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_IND.pdf}
     \caption{Stringency and Normalized GDP for India}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_IND.eps}
+    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_IND.pdf}
     \caption{Normalized GDP modelled with Stringency for India}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_MEX.eps}
+    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_MEX.pdf}
     \caption{Stringency and Normalized GDP for Mexico}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_MEX.eps}
+    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_MEX.pdf}
     \caption{Normalized GDP modelled with Stringency for Mexico}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_BRA.eps}
+    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_BRA.pdf}
     \caption{Stringency and Normalized GDP for Brazil}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_BRA.eps}
+    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_BRA.pdf}
     \caption{Normalized GDP modelled with Stringency for Brazil}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_USA.eps}
+    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_USA.pdf}
     \caption{Stringency and Normalized GDP for United States}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_USA.eps}
+    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_USA.pdf}
     \caption{Normalized GDP modelled with Stringency for United States}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_JPN.eps}
+    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_JPN.pdf}
     \caption{Stringency and Normalized GDP for Japan}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_JPN.eps}
+    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_JPN.pdf}
     \caption{Normalized GDP modelled with Stringency for Japan}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_CAN.eps}
+    \includegraphics[width=\linewidth]{images/stringency_vs_gdp_CAN.pdf}
     \caption{Stringency and Normalized GDP for Canada}
@@
     \centering
-    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_CAN.eps}
+    \includegraphics[width=\linewidth]{images/gdp_modelled_with_stringency_CAN.pdf}
     \caption{Normalized GDP modelled with Stringency for Canada}
@@
 
-\subsection*{Discussion}
+From this we conclude, that maybe there are other factors that explain the dip in the normalized GDP than just an increase in the stringency.
+
+After median filtering to smooth the output (to reinforce the negative reward from changing the stringencies) from the trained reinforcement learning agent, here are some of the results obtained:
+
+\begin{figure}[H]
+  \centering
+  \caption{Strategy from Reinforcement Learning Agent 1}
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/914857/rl_sir.eps}
+    \caption{SIR Dynamics}
+  \end{subfigure}
+  \label{fig:914857_rl_sir}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/914857/rl_i.eps}
+    \caption{Infected Population changing over Time}
+  \end{subfigure}
+  \label{fig:914857_rl_i}
+
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/914857/rl_gdp.eps}
+    \caption{Normalized GDP changing over Time}
+  \end{subfigure}
+  \label{fig:914857_rl_gdp}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/914857/rl_stringency.pdf}
+    \caption{Stringency changing over Time}
+  \end{subfigure}
+  \label{fig:914857_rl_stringency}
+
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/914857/rl_r_eff.eps}
+    \caption{$R_e$ changing over Time}
+  \end{subfigure}
+  \label{fig:914857_rl_r_eff}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/914857/rl_reward.eps}
+    \caption{Reward changing over Time}
+  \end{subfigure}
+  \label{fig:914857_rl_reward}
+\end{figure}
+
+In the presented result, we can see the reinforcement learning agent outperform the modelled outcome.
+A strategic decision is made by the agent to maintain the stringency index below 80 throughout the course of the study.
+This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to $1.2$, significantly ahead of the projected timeline (refer to figure \cref{fig:914857_rl_r_eff}).
+A subsequent decrease in stringency in the later stages leads to an increase in the normalized Gross Domestic Product (GDP), indicating an economic upturn.
+While this strategy poses a higher risk in terms of infection rates during the initial phase of the epidemic (prior to vaccine rollout), it proves to be more beneficial for the nation's economy in the long run.
+It is suggested that a more balanced policy, falling between the actual and the proposed strategy in this study, could have been a more effective approach for the government to adopt.
+This would have potentially led to a more optimal balance between controlling the spread of the disease and minimizing economic impact.
+
+\begin{figure}[H]
+  \centering
+  \caption{Strategy from Reinforcement Learning Agent 2}
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/244620/rl_sir.eps}
+    \caption{SIR Dynamics}
+  \end{subfigure}
+  \label{fig:244620_rl_sir}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/244620/rl_i.eps}
+    \caption{Infected Population changing over Time}
+  \end{subfigure}
+  \label{fig:244620_rl_i}
+
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/244620/rl_gdp.eps}
+    \caption{Normalized GDP changing over Time}
+  \end{subfigure}
+  \label{fig:244620_rl_gdp}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/244620/rl_stringency.pdf}
+    \caption{Stringency changing over Time}
+  \end{subfigure}
+  \label{fig:244620_rl_stringency}
+
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/244620/rl_r_eff.eps}
+    \caption{$R_e$ changing over Time}
+  \end{subfigure}
+  \label{fig:244620_rl_r_eff}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/244620/rl_reward.eps}
+    \caption{Reward changing over Time}
+  \end{subfigure}
+  \label{fig:244620_rl_reward}
+\end{figure}
+
+An alternative outcome is presented by the agent, demonstrating a gradual escalation in stringency over time.
+This strategy results in the majority of infections occurring prior to the vaccine's release, with a subsequent cessation of new infections.
+This approach, however, has implications for the normalized Gross Domestic Product (GDP), as seen by the observed decline in figure \cref{fig:244620_rl_gdp}.
+Furthermore, the effective reproduction number ($R_e$) does not achieve the target value of 1.2 under this strategy.
+These findings underscore the complexity of managing public health crises and the need for careful strategic planning to balance health outcomes with economic considerations.
+
+\section{Discussion}
 The paper seeks to inspire epidemiologists by highlighting the advancements achieved through the application of reinforcement learning in policy making during the pandemic.
@@
 
-\subsection*{Experiment Settings}
+\section{Experiment Settings}
 \subsection{Dataset}
 The population-level epidemiological data can be obtained from the 'Our World In Data COVID-19' dataset: \url{https://ourworldindata.org/coronavirus} or more specifically: \url{https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-data.csv}~\cite{owidcoronavirus}.
+Data for the total cases, and recovered was acquired by scraping the Worldometers website \cite{WorldometerCorona} used Internet Archive \cite{InternetArchive}.
 Quaterly GDP data can be obtained from the 'Organisation for Economic Co-operation and Development': \url{https://www.oecd-ilibrary.org/economics/data/main-economic-indicators/main-economic-indicators-complete-database_data-00052-en}~\cite{economic_indicators_data}.
@@
 \subsection{Code}
-We used stable-baseline3~\cite{JMLR:v22:20-1364}, Pytorch~\cite{paszke2019pytorch}, Scipy~\cite{Virtanen_2020}, Pandas, Matplotlib, Python~\cite{ScientificPython}.
+We used stable-baseline3~\cite{JMLR:v22:20-1364}, Pytorch~\cite{paszke2019pytorch}, Scipy~\cite{Virtanen_2020}, Pandas, Matplotlib, in Python~\cite{ScientificPython}.
 Code: \url{https://github.com/psymbio/sir_rl}

### 2024-02-06 04_reinforcement_learning.py

@@ -18,8 +18,8 @@ import time
 env = SIREnvironment()
 check_env(env)
 
-models_dir = f"models/{int(time.time())}/"
-logdir = f"logs/{int(time.time())}/"
+models_dir = f"models/{int(time.time())}_herd/"
+logdir = f"logs/{int(time.time())}_herd/"
 
 if not os.path.exists(models_dir):
 	os.makedirs(models_dir)

### 2024-02-06 05_reinforcement_run.py

@@ -24,8 +24,13 @@ env.reset()
 # models/1705085840/16452.zip
 # models/1705154668/548400.zi
 # models/1705305871/578562.zip
-# model_path = os.path.join(MODELS_DIR, "1705305871", "578562.zip")
-model_path = os.path.join(MODELS_DIR, "1705318802", "586788.zip")
+# model_path = os.path.join(MODELS_DIR, "1705305871", "578562.zip")\
+# models/1705330479/839052.zip
+# model_path = os.path.join(MODELS_DIR, "1705330479", "839052.zip")
+# model_path = os.path.join(MODELS_DIR, "1705049761", "79518.zip")
+# models/1705427702_herd/274200.zip
+# model_path = os.path.join(MODELS_DIR, "1705427702_herd", "274200.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run", "1746654.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 10
@@ -40,11 +45,17 @@ for episode in range(1, episodes+1):
         actions_taken.append(action)
         obs, reward, terminated, truncated, info = env.step(action)
         info_save.append(info)
-        score += reward
+        # score += reward
+        # score += info['reward']
+        # with open(os.path.join(OUTPUT_DIR, "score_check.txt"), "a") as f:
+        #     for action in actions_taken:
+        #         f.write(str(reward) +"\n")
+    info_save_df = pd.DataFrame.from_dict(info_save)
+    score = sum(info_save_df['reward'])
+    info_save_df.to_csv(os.path.join(OUTPUT_DIR, "info_save", f"{score:.2f}.csv"), index=False)
+    env.render(score=score, stringency=np.array(info_save_df['stringency_index']))
     print(f'Episode: {episode}, Score: {score}')
     with open(os.path.join(OUTPUT_DIR, "actions_taken", f"{score:.2f}.txt"), "w") as f:
         for action in actions_taken:
             f.write(str(action) +"\n")
-    info_save_df = pd.DataFrame.from_dict(info_save)
-    info_save_df.to_csv(os.path.join(OUTPUT_DIR, "info_save", f"{score:.2f}.csv"), index=False)
-    env.render(score=score, stringency=np.array(info_save_df['stringency_index']))
+    

### 2024-02-06 06_train_on_previous_model.py

@@ -9,7 +9,7 @@ env = SIREnvironment()
 # model_path = os.path.join(MODELS_DIR, "1705154668", "548400.zip")
 # models/1705318802/586788.zip
 model_path = os.path.join(MODELS_DIR, "1705318802", "586788.zip")
-iters = 586788 / TIMESTEPS
+iters = int(586788 / TIMESTEPS)
 model = PPO.load(model_path)
 env.reset()
 model.set_env(env)

### 2024-02-06 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -106,7 +106,7 @@ class SIREnvironment(gym.Env):
         # each action is on ith day
         t_ith_day = np.linspace(0, self.ith_day, self.ith_day + 1)
         self.stringency_index_list.append(self.stringency_index)
-        predictions = odeint(deriv, self.y0, t_ith_day, args=(self.N, self.optimal_beta, self.optimal_gamma, np.array(self.nu_varying), np.array(self.stringency_index_list) / 100))
+        predictions = odeint(deriv, self.y0, t_ith_day, args=(self.N, self.optimal_beta, self.optimal_gamma, np.array(self.nu_varying), np.array(self.stringency_index_list[:TOTAL_DAYS]) / 100))
         S, I, R = predictions.T
         self.store_S[self.ith_day] = S[-1]
         self.store_I[self.ith_day] = I[-1]
@@ -127,9 +127,12 @@ class SIREnvironment(gym.Env):
         # REMEMBER: to change this definition of the reward in the render as well!!!
         # self.reward = self.normalized_GDP - (2 * self.r_eff)
 
-        reward_inertia = abs(diff_action)*-1*5
+        reward_inertia= abs(diff_action)*-1*5
+        # reward_inertia = abs(self.stringency_index_list[-2] - self.stringency_index_list[-1]) * -1 * 5
+        # print("reward_inertia: ", reward_inertia, "reward_inertia_old: ", reward_inertia_old)
         reward_r_eff = 10 if self.r_eff <= 1.9 else -10
-        reward_I_percentage = -5000 if self.I_proportion >= 0.003 else 20
+        reward_I_percentage = -5000 if self.I_proportion >= 0.003 else 0
+        reward_herd_immunity = 2000 if self.I_proportion <= 0.003 and self.I_proportion >= 0.0005 else 0
 
         gdp_reward_weight_1 = 100
         gdp_reward_weight_2 = 200
@@ -140,7 +143,8 @@ class SIREnvironment(gym.Env):
         else:
             reward_weighted = gdp_reward_weight_2 * self.normalized_GDP
 
-        self.reward = reward_weighted + reward_inertia + reward_r_eff + reward_I_percentage
+        self.reward = reward_weighted + reward_inertia + reward_r_eff + reward_I_percentage + reward_herd_immunity
+        # print(self.ith_day, self.reward)
         self.store_reward[self.ith_day] = self.reward
[…]

### 2024-02-07 paper_draft/main.tex
paper changes - need to add figure captions

@@
 These compartments create the SIR model which can be represented as follows:
-
 \subsection{Simple SIR Model}
-\begin{figure}[H]
+\begin{figure}[htbp!]
 \centering    
@@
 \end{equation}
-
 Here, $\lambda$ is the force of infection, it is the rate at which susceptible individuals acquire an infectious disease~\cite{hens_aerts_faes_shkedy_lejeune_vandamme_beutels_2010}.
 It depends on other factors:
-
 \begin{equation}
@@
 \end{equation}
-
 Here, $c$ is the average number of contacts a susceptible person makes per day.
@@
 And, $\beta$ the effective transmission rate is defined as:
-
 \begin{equation}
@@
 \end{equation}
-
 During an epidemic, the fundamental drivers of an epidemic growth is the rate of infection $\beta$ i.e.
@@
 $R_e$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic, therefore, it takes into account the existing immunity of the system.
-
 \begin{equation}
@@
 \end{equation}
-
 $S$ is the number of susceptible people, $N$ is the total population.
@@
 However, sometimes we must strike a balance between modeling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}).
-
 \begin{equation}
@@
 \end{equation}
-
 In the above equation, $y$ is the actual data and $f(x)$ is the prediction.
-
 \begin{equation}
@@
 \end{equation}
-
 Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered.
@@
 The following parameters and loss is obtained:
-
 \begin{equation}
@@
 \end{equation}
-
 See \cref{fig:SIR_model_IND} to see how the model compares with the actual data.
@@
 a higher score does not necessarily mean that a country's response is "better" than others lower on the index.
-
 To define the new time-varying beta that is dependent on the current stringency index, the following equations have been formulated:
-
 \begin{equation}
@@
 \end{equation}
-
 Where, $s(t)$ is the stringency index at time $t$ and is scaled down by a factor of 100 to normalize it and bring it in the range $s(t)/100 \in [0, 1]$.
@@
 Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method, we get the following parameters and loss:
-
 \begin{equation}
@@
 \end{equation}
-
 \subsection{SIRV Model with Lockdown}
 Lastly, an additional flow from the susceptible population to the recovered population can be shown by adding a vaccination rate $\nu$ in the model. 
-
 \begin{equation}
@@
 \end{equation}
-
 Optimizing these equations with \cref{eq:cost_SIR} using the Nelder-Mead method:
-
 \begin{equation}
@@
     R_0 \in [0.165, 3.047]
-  \begin{split}
+  \end{split}
   \label{eq:r0_with_lockdown_and_nu}
@@
 
-\subsection{Optimizing Window Length for Time-varying $\nu$}\label{optimizing_window_length}
+\subsection{Optimizing Window Length for Time-varying Vaccination Rate}\label{optimizing_window_length}
 However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time.
@@
 It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
-
 \begin{equation}
@@
 \end{equation}
-
 The main difference here is instead of supplying these equations with the entire time series, only data lying within the sub-interval of $[{start}, {start} + {window\_length}]$ is given, then in the next step data of the next sub-inteval is provided to optimize for, i.e., $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$.
@@
 
-\begin{figure}
+\begin{figure}[htbp!]
   \caption{Loss for Different Window Lengths}
@@
     \caption{Loss for Different Window Lengths for Susceptible, Infected and Recovered Population}
+    \label{fig:window_length_loss_SIR_IND}
   \end{subfigure}
-  \label{fig:window_length_loss_SIR_IND}
   \hfill
@@
     \caption{Loss for Different Window Lengths for Infected Population}
+    \label{fig:window_length_loss_I_IND}
   \end{subfigure}
-\label{fig:window_length_loss_I_IND}
 \end{figure}
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 
-\subsection{SIRV Model with Lockdown and time-varying $nu$}
+\subsection{SIRV Model with Lockdown (Time-varying Vaccination Rate)}
 Lastly, we present our model using the following set of equations:
-\begin{figure}[H]
+\begin{figure}[htbp!]
 \begin{tikzpicture}[node distance=1cm, auto,
@@
 \end{figure}
-
 \begin{equation}
@@
 \end{equation}
-
-See \Crefrange{fig:SIR_model_IND,fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \cref{fig:comparing_costs_IND} to see how the different models compare against each other.
+See \Crefrange{fig:SIR_model_IND}{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \cref{fig:comparing_costs_IND} to see how the different models compare against each other.
 
@@
 For India after fitting a 3 degree polynomial, the following equation is obtained:
-
 \begin{equation}
@@
 
+\subsection{Reinforcement Learning}
 Given that the government is an agent that takes decisions in a deterministic environment defined above, we use reinforcement learning to model the competing costs of the environment.
@@
 \end{itemize}
-
 In general, the policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$.
@@
 We train the model for $2742$ time steps for $200$ iterations and the best results are presented below.
-
 \newcommand{\DrawNeuronalNetwork}[2][]{
@@
 \begin{tikzpicture}[x=1.5cm, y=1.5cm,
->=stealth,font=\sffamily,nodes={align=center}]
+>=stealth,font=\sffamily,nodes={align=center}, scale=0.8, every node/.style={scale=0.8}]
  \begin{scope}[local bounding box=T]
@@
 Using the simple SIR model from \Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown},  to model the disease dynamics we get:
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 Combining the lockdown dynamics in the SIR model using \Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}, we get the following:
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 \end{figure}
-\begin{figure}[H]\ContinuedFloat
+\begin{figure}[htbp!]\ContinuedFloat
   \begin{subfigure}[t]{\textwidth}
@@
 Combining vaccination dynamics with SIR with lockdown model using \Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}, we get the following:
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
   \end{subfigure}
-
 \end{figure}
@@
 For time-varying $\nu$ and SIRV model with lockdown from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following:
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp!]
+	\caption{Loss for Different Models}
+	\begin{subfigure}[t]{0.48\textwidth}
 		\centering
-  \caption{Loss for Different Models}
-  \includegraphics[width=\linewidth]{images/comparing_costs_IND.pdf}
-  \label{fig:comparing_costs_IND}
+		\includegraphics[width=\linewidth]{images/comparing_costs_SIR_IND.pdf}
+		\caption{Loss for Different Models for Susceptible, Infected and Recovered Population}
+		\label{fig:comparing_costs_SIR_IND}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.48\textwidth}
+		\centering
+		\includegraphics[width=\linewidth]{images/comparing_costs_I_IND.pdf}
+		\caption{Loss for Different Models for Infected Population}
+		\label{fig:comparing_costs_I_IND}
+	\end{subfigure}
 \end{figure}
@@
 Below are scatter plots with pearson correlation, for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May, 2020 to October, 2022.
-
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 \end{figure}
-
 It can be observed that stringency has a negative impact on the normalized Gross domestic product (GDP).
@@
 This is not the case for Advanced Economies like (USA, Japan, Canada).
-
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 \end{figure}
-
 From this we conclude, that maybe there are other factors that explain the dip in the normalized GDP than just an increase in the stringency.
@@
 After median filtering to smooth the output (to reinforce the negative reward from changing the stringencies) from the trained reinforcement learning agent, here are some of the results obtained:
-
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 \end{figure}
-
 In the presented result, we can see the reinforcement learning agent outperform the modelled outcome.
@@
 This would have potentially led to a more optimal balance between controlling the spread of the disease and minimizing economic impact.
-
-\begin{figure}[H]
+\begin{figure}[htbp!]
   \centering
@@
 \end{figure}
-
 An alternative outcome is presented by the agent, demonstrating a gradual escalation in stringency over time.

### 2024-02-14 paper_draft/main.tex
paper edits, graphs changed

@@
 
-In this paper we optimize the government policies regarding stringency.
-Therefore, alongside epidemiological data, we use the measures of globally comparable government responses~\cite{Hale2021}.
-We use the simple SIR model without vital dynamics~\cite{Hethcote1989, Hethcote2008, ALLEN2017128} as it is assumed that the timescale is small enough that it can be neglected~\cite{Cooper2020}.
-By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we can effectively address the real-world conditions and propose a solution that is both effective and extendable.
-The current model (SIRV with lockdown) accounts for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{SIRLockdown,NBERw26981,NBERw26867}.
-Traditional SIR model neglects the time-varying property of parameters like $\beta$ and $\gamma$, but in this paper since, lockdown is a parameter of the proposed model, and it changes with time therefore, we propose a time-dependent SIR model\cite{TimeDependentSIR}.
+In this paper we optimize the government policies regarding stringency as it controls both the spread of the disease and the economy.
+To model the epidemiological data~\cite{WorldometerCorona} we use the simple SIR model without vital dynamics~\cite{Hethcote1989, Hethcote2008, ALLEN2017128} as it is assumed that the timescale is small enough that it can be neglected~\cite{Cooper2020}.
+By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we effectively model the disease progression.
+Our model (SIRV with lockdown and time-varying vaccination rate) in addition to the SIR model, accounts for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{Hale2021,SIRLockdown,NBERw26981,NBERw26867}.
+Eventhough the traditional SIR model is a valuable tool for understanding the spread of infectious diseases, it assumes that parameters like the transmission rate ($\beta$) and recovery rate ($\gamma$) are constant over time, which may not always be the case.
+In this paper, we propose a more sophisticated approach by introducing a time-dependent SIR model~\cite{TimeDependentSIR}, enabling us to account for the changing dynamics of the pandemic due to factors such as lockdowns and vaccination rates.
+This proposed model effectively addresses the real-world conditions acts as a solution that is both effective and extendable.
 However, the study has limitations; First, the deterministic SIR model (predecessor to our proposed model) fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}.
@@
 
-After modelling the disease with lockdown (via stringency index) and vaccination, we try to understand the effects of lockdown on the GDP.
+After modelling the disease with lockdown (via stringency index) and vaccination, we try to understand the effects of lockdown on the GDP~\cite{StringencyEconomicDecline,Cilloni2020,Arinaminpathy2010}.
 Therefore, decisions made by the government regarding the level of lockdown to be enforced plays a role on both the public health outcomes and economic stability during a pandemic.
@@
 
-In summary, by using reinforcement learning augmented with deep learning techniques for the SIRV with lockdown environment, we can better understand the effects of lockdown measures on both public health outcomes and economic stability during a pandemic.
+In summary, by using reinforcement learning augmented with deep learning techniques for the SIRV with lockdown and time-varying vaccination rate environment, we can better understand the effects of lockdown measures on both public health outcomes and economic stability during a pandemic.
 However, it is crucial to consider the limitations of this approach and take into account a comprehensive set of decision factors in order to make informed policy decisions that are tailored to specific socioeconomic contexts.
@@
 
-At the peak of an epidemic $R_0$ declines as there are no more susceptible people left in the pool, therefore, $R_e$ (effective reproductive number) comes into play.
-$R_e$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic, therefore, it takes into account the existing immunity of the system.
+At the peak of an epidemic there is a decline as there are no more susceptible people left in the pool, therefore, $R_e$ (effective reproductive number) comes into play.
+$R_e$ is defined as the average number of secondary cases arising from an infected case, at a given point in an epidemic, therefore, it takes into account the existing immunity of the system~\cite{MASSAD2017232}.
 \begin{equation}
-  R_e = R_0 \frac{S}{N}
+  R_e = R_0 \frac{S(t)}{N}
   \label{eq:R_effective}
@@
 We generally calibrate the model using \cref{eq:cost_SIR}.
-However, sometimes we must strike a balance between modeling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}).
+However, sometimes we must strike a balance between modeling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}, where a hyperparameter is selected keeping both these losses in mind).
 \begin{equation}
@@
 \subsection{SIRV Model with Lockdown}
-Lastly, an additional flow from the susceptible population to the recovered population can be shown by adding a vaccination rate $\nu$ in the model. 
+Lastly, an additional flow from the susceptible to recovered population can be shown by adding a vaccination rate $\nu$ in the model. 
 \begin{equation}
@@
 \begin{equation}
-  \textrm{loss\_SIR} = 10840360.995
+  \textrm{loss\_I} = 10840360.995
   \label{eq:cost_I_with_lockdown_and_nu}
@@
 This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}.
-Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} i.e., the window length that result in the least loss from \cref{eq:cost_SIR}.
-Using different window lengths to estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$, we observe the following.
+Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} i.e., a window length that results in the least loss from \cref{eq:cost_SIR,eq:cost_I}.
+Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ then for the next time-step $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$ we again estimate $\nu$ -- so on and so forth.
 It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
@@
 \end{equation}
-The main difference here is instead of supplying these equations with the entire time series, only data lying within the sub-interval of $[{start}, {start} + {window\_length}]$ is given, then in the next step data of the next sub-inteval is provided to optimize for, i.e., $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$.
 This model gives shows how $\nu$ varies with time.
@@
 
-From the following graph, although the window length of $10$ is best for all three groups of the population but this window length leads to a bad approximation for the infected population therefore, we choose a window length of $15$.
+The following graph indicates that a window length of $10$ days yields the least overall loss for all three population groups.
+However, this window length results in a poor approximation for the infected group, which is crucial for accurately modeling the spread of the disease.
+Consequently, we have decided to use a window length of $15$ days, which provides a more accurate approximation for the infected population while still maintaining reasonable loss for the other groups.
 
@@
 
-This shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India \cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
+This shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
 Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIRV Model with time varying $\nu$.
@@
 \subsection{SIRV Model with Lockdown (Time-varying Vaccination Rate)}
-Lastly, we present our model using the following set of equations:
+Finally, we integrate the time-varying vaccination rate ($\nu$) into the SIR model that includes lockdown measures, resulting in the following set of equations, which represents our final model:
 \begin{figure}[htbp!]
@@
 \end{equation}
-See \Crefrange{fig:SIR_model_IND}{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \cref{fig:comparing_costs_IND} to see how the different models compare against each other.
+See \Crefrange{fig:SIR_model_IND}{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \Crefrange{fig:comparing_costs_SIR_IND}{fig:comparing_costs_I_IND} to see how the different models compare against each other.
 
-Now, that a relation between $\beta$ and $s(t)$ is set up, it must be investigated how stringency index affects the normalized Gross domestic product (GDP)~\cite{OECDNormalizedGDP,OECDNormalizedGDP2}.
-To do this a polynomial equation of the third degree is fitted to the data points $(x, y)$ where $x$ is the $s(t)$ stringency at time $t$, and $y$ the normalized GDP and minimize the squared error.
+\subsection{Modelling Normalized GDP with Stringency}
+Now, that a relation between $\beta$ and $s(t)$ is set up, it must be investigated how stringency index affects the normalized gross domestic product (GDP)~\cite{OECDNormalizedGDP,OECDNormalizedGDP2}.
+To do this a polynomial equation of the third degree is fitted to the data points $f(x) = a x^3 + b x^2 + c x + d = y$, here, $x$ is the $s(t)$ stringency at time $t$, and $y$ the normalized GDP, and we minimize the squared error to find the values of coefficients $a, b, c, d$.
 For India after fitting a 3 degree polynomial, the following equation is obtained:
@@
     \begin{split}
-      GDP = -5.96640236 \times 10^{-5} s^{3} + 6.65064332 \times 10^{-3} s^{2} - 2.23109924 \times 10^{-1} s^{1} \\
+      \textrm{normalized\_GDP} = -5.96640236 \times 10^{-5} s^{3} + 6.65064332 \times 10^{-3} s^{2} - 2.23109924 \times 10^{-1} s^{1} \\
     + 1.01357226 \times 10^{2}
@@
 Given that the government is an agent that takes decisions in a deterministic environment defined above, we use reinforcement learning to model the competing costs of the environment.
-This environment is formally as a Markov decision process, and can be described as follows:
-
+This environment is known as a Markov decision process (MDP), and can be described as follows:
+% https://arxiv.org/pdf/1708.05866.pdf
 \begin{itemize}
     % \item \url{https://jmlr.org/papers/volume22/20-1364/20-1364.pdf} RL Paper that references \url{https://araffin.github.io/slides/rl-tuto-jnrr19/#/13} RL ppt (for our RL notes).
-    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP ($(GDP_{predicted} - GDP_{min}) / (GDP_{max} - GDP_{min})$), $R_e$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible, infected and recovered.
+    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP ($(GDP_{predicted} - GDP_{min}) / (GDP_{max} - GDP_{min})$), $R_e$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible ($S$), infected ($I$) and recovered ($R$).
 The starting states are simply these values at the starting date and no previous actions.
@@
     \item Immediate reward $\altmathcal{R}\left(\mathbf{s}_t, \mathbf{a}_t, \mathbf{s}_{t+1}\right)$.
-Here we define a reward strategy however it should be noted that this strategy can be easily changed to prioritize different needs.
+The agent observes the state of the environment $\mathbf{s}_t$ at time $t$ and takes an action $\mathbf{a}_t$, after which the state transitions to $\mathbf{s}_{t+1}$ and the agent receives a reward $\mathbf{r}_{t+1}$ as feedback.
+In \cref{defining_reward_function} we define a reward strategy however it should be noted that this work serves as a framework where the strategy can be easily swapped for another to prioritize different needs.
     \item Discount Factor $\gamma \in[0,1]$, where lower values place more emphasis on immediate rewards. Here, we choose the default discount factor of 0.99.
 \end{itemize}
+
 In general, the policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$.
 If the MDP is episodic, i.e., the state is reset after each episode of length $T$, then the sequence of states, actions and rewards in an episode constitutes a trajectory or rollout of the policy.
-Every rollout of a policy accumulates rewards from the environment, resulting in the return $R=\sum_{t=0}^{T-1} \gamma^t r_{t+1}$.
+Every rollout of a policy accumulates rewards from the environment, resulting in the return $R=\sum_{t=0}^{t=T-1} \gamma^t r_{t+1}$.
 The goal of RL is to find an optimal policy, $\pi^*$, which achieves the maximum expected return from all states.
 To achieve this, reinforcement learning start with an initial arbitrary policy, i.e., a $Q$-table with no entries.
-$Q$-table is a mapping from states $s_k \in \mathscr{S}$ to a predefined set of actions to increase or decrease the stringency at time t, which are the actions $a_k \in \mathscr{A}$.
-Each entry of the $Q$-table $\left(Q_k\left(s_k, a_k\right)\right)$ associates an action in the finite sequence $\left(\mathscr{A}_j\right)_{j \in \mathbb{J}^{+}}$ to a state of the finite sequence $\left(\mathscr{S}_i\right)_{i \in \mathbb{I}^{+}}$ \cite{sutton2018reinforcement}.
-In this case of epidemic control by NPI based strategies this policy represents the series of stringencies to be imposed upon the population to shift the initial status of the environment to a targeted status which is equivalent to the desired set of system states.
-This is how the $Q$-table updates saying, if in state $s_k$ the most ideal action is $a_k$.
-After having more and more experience with the environment and understanding which actions lead to a higher reward $r$ an optimal policy is derived be maximizing the expected value of discounted reward $J\left(r_k\right)=\mathbb{E}\left[\sum_{k=1}^{\infty} \gamma^{(k-1)} r_k\right]$, where, discount factor $\gamma \in [0, 1]$ and time steps $k = 1, 2, \dots$.
+$Q$-table is a mapping from states $s_t \in \mathscr{S}$ to a predefined set of actions to increase or decrease the stringency at time t, which are the actions $a_t \in \mathscr{A}$.
+Each entry of the $Q$-table $\left(Q_t\left(s_t, a_t\right)\right)$ associates an action in the finite sequence $\left(\mathscr{A}_j\right)_{j \in \mathbb{J}^{+}}$ to a state of the finite sequence $\left(\mathscr{S}_i\right)_{i \in \mathbb{I}^{+}}$~\cite{sutton2018reinforcement}.
 
+In this case of epidemic control by non-pharmaceutical interventions (NPI) based strategies this policy represents the series of stringencies to be imposed upon the population to shift the initial status of the environment to a targeted status which is equivalent to the desired set of system states.
+This is how the $Q$-table updates saying, if in state $s_k$ the most ideal action is $a_t$.
+After having more and more experience with the environment and understanding which actions lead to a higher reward $r$ an optimal policy is derived be maximizing the expected value of discounted reward $J\left(r_t\right)=\mathbb{E}\left[\sum_{t=1}^{\infty} \gamma^{(t-1)} r_t\right]$, where, discount factor $\gamma \in [0, 1]$ and time steps $k = 1, 2, \dots$.
+
+\subsubsection{Defining the Reward Function}\label{defining_reward_function}
 The stringency index emerges as a critical factor influencing both the Gross Domestic Product (GDP) and the rate of infection spread.
@@
 By designing a proper reward function, it is possible to generate such an agent that may follow the human desired situation.
-
+While designing a reward function it is important to note that the rewards we set up truly indicate what we want accomplished.
+In particular, the reward signal is not the place to impart to the agent prior knowledge about how to achieve what we want it to do~\cite{sutton2018reinforcement}.
 % TODO add: http://www.incompleteideas.net/book/ebook/node29.html
@@
 The reward function is parameterized to account for key factors influencing decision-making.
-To incentivize reduction of $R{eff}$ (effective reproductive number) a negative reward is imposed of $-20 * R_e$, but as the $R_e$ is between $[1.9, 1,5]$ we start to positively reward it with a multiple of the GDP ($100 * \textrm{min\_max\_normalized\_GDP}$) and if the $R_e$ is below $1.5$ then $200 * \textrm{min\_max\_normalized\_GDP}$.
-Furthermore, there's a simple positive reward of $10$ if $R_e$ is below the threshold value ($1.9$) and negative reward of $-10$ otherwise.
-To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| * -1 * 5$).
+To incentivize reduction of $R_e$ (effective reproductive number) the reward is $((R_e^{2}) \times -100) + 100$, this means when $R_e = 1$ there is no reward, but it exponentially decreases when $R_e > 1$ and increases if $R_e < 1$.
+Similarly, for the normalized GDP the reward function is $((\textrm{min\_max\_normalized\_GDP} ** 5)*300) - 80$ which tends to become negative below the average
+
 If the proportion of the infected population were to rise above 0.003 the model is punished ($-5000$) and otherwise rewarded ($+20$).
-As a reward for herd immunity, another reward is defined using the proportion of the infected population.
-If it lies between $[0.0005, 0.003]$ a reward of 2000 is given - else $0$.
+To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| * -20$).
 
@@
 The population-level epidemiological data can be obtained from the 'Our World In Data COVID-19' dataset: \url{https://ourworldindata.org/coronavirus} or more specifically: \url{https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-data.csv}~\cite{owidcoronavirus}.
-Data for the total cases, and recovered was acquired by scraping the Worldometers website \cite{WorldometerCorona} used Internet Archive \cite{InternetArchive}.
+Data for the total cases, and recovered was acquired by scraping the Worldometers website~\cite{WorldometerCorona} used Internet Archive \cite{InternetArchive}.
 Quaterly GDP data can be obtained from the 'Organisation for Economic Co-operation and Development': \url{https://www.oecd-ilibrary.org/economics/data/main-economic-indicators/main-economic-indicators-complete-database_data-00052-en}~\cite{economic_indicators_data}.

### 2024-02-14 05_reinforcement_run.py

@@ -3,34 +3,8 @@ from rl_utils_new_complex_reward_sirvlockdown_nu_varying import *
 env = SIREnvironment()
 env.reset()
 
-# models/1704106458/10000.zip
-# models/1704189286/1033.zip
-# models/1704190312/4132.zip
-# models/1704192521/72310.zip
-# models/1704205452_ppo_100iter/103300.zip
-# models/1704256387_ppo_100iter_hospital-100/103300.zip
-# models/1704275312/1033.zip
-# models/1704351335/542325.zip
-# models/1704382102/346055.zip
-
-# models/1704534388/1229270.zip
-# models/1704534388/356385.zip
-# models/1704534388/139455.zip
-
-# models/1704873042/5165.zip
-# models/1704874281/374740.zip
-# models/1704994098/2742.zip
-# models/1705049761/79518.zip
-# models/1705085840/16452.zip
-# models/1705154668/548400.zi
-# models/1705305871/578562.zip
-# model_path = os.path.join(MODELS_DIR, "1705305871", "578562.zip")\
-# models/1705330479/839052.zip
-# model_path = os.path.join(MODELS_DIR, "1705330479", "839052.zip")
-# model_path = os.path.join(MODELS_DIR, "1705049761", "79518.zip")
-# models/1705427702_herd/274200.zip
-# model_path = os.path.join(MODELS_DIR, "1705427702_herd", "274200.zip")
-model_path = os.path.join(MODELS_DIR, "hpc_run", "1746654.zip")
+# models/hpc_run/2667966.zip
+model_path = os.path.join(MODELS_DIR, "hpc_run", "2667966.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 10

### 2024-02-14 google_colab_upload/rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -162,7 +162,7 @@ class SIREnvironment(gym.Env):
         info = {
             "action": action,
             "reward_inertia": reward_inertia,
-            "reward_r_eff": reward_r_eff,
+            # "reward_r_eff": reward_r_eff,
             "reward_I_percentage": reward_I_percentage,
             "reward_weigthed": reward_weighted,
             "reward": self.reward,

### 2024-02-14 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -51,17 +51,13 @@ def deriv(y, t, N, beta, gamma, nu_varying, lockdown):
     return dSdt, dIdt, dRdt
 
 def calculate_reward_weighted(gdp_min_max_normalized_list, r_eff_list):
-    GDP_WEIGHT_1 = 100 # change this value and see how it affects the reward
-    GDP_WEIGHT_2 = 200 # change this value and see how it affects the reward
-    reward = []
-    for i in range(len(gdp_min_max_normalized_list)):
-        if r_eff_list[i] > 1.9:
-            reward.append(-20 * r_eff_list[i])
-        elif r_eff_list[i] <= 1.9 and r_eff_list[i] >= 1.5:
-            reward.append(GDP_WEIGHT_1 * gdp_min_max_normalized_list[i])
-        else:
-            reward.append(GDP_WEIGHT_2 * gdp_min_max_normalized_list[i])
-    return reward
+    reward_list = []
+    for gdp, r_eff in zip(gdp_min_max_normalized_list, r_eff_list):
+        gdp_reward = ((gdp ** 5)*300) - 80
+        r_eff_reward = ((r_eff**2)*-100) + 100
+        total_reward = gdp_reward + r_eff_reward
+        reward_list.append(total_reward)
+    return reward_list
 
 class SIREnvironment(gym.Env):
     metadata = {"render_modes": ["human"], "render_fps": 30}
@@ -119,42 +115,33 @@ class SIREnvironment(gym.Env):
         self.S_proportion = self.store_S[self.ith_day]/self.N
         self.I_proportion = self.store_I[self.ith_day]/self.N
         self.R_proportion = self.store_R[self.ith_day]/self.N
-        self.normalized_GDP = (fit_line_loaded(self.stringency_index) - MIN_GDP) / (MAX_GDP - MIN_GDP)
-        self.r_eff = (self.optimal_beta / self.optimal_gamma) * (self.store_S[self.ith_day] / self.N)
+        self.normalized_GDP_min_max_normalized = (fit_line_loaded(self.stringency_index) - MIN_GDP) / (MAX_GDP - MIN_GDP)
+        self.r_eff = (self.optimal_beta / self.optimal_gamma) *  (1 - (self.stringency_index/100)) * (self.store_S[self.ith_day] / self.N)
         self.store_r_eff[self.ith_day] = self.r_eff
-        self.store_normalized_gdp[self.ith_day] = self.normalized_GDP
[…]

### 2024-02-17 paper_draft/main.tex
some more edits

@@
 
-In general, the policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$.
+Given that at each timestep $t$ the agent has to choose an action $a_t$ so as to maximize the reward $r_{t+1}$ a policy is formulized by the agent.
+The policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$.
+Reinforcement learning methods specify how the agent changes its policy as a result of its experience.
 If the MDP is episodic, i.e., the state is reset after each episode of length $T$, then the sequence of states, actions and rewards in an episode constitutes a trajectory or rollout of the policy.
@@
 
-With this basis we define the reward function~\cite{Ohi2020}.
-In Deep Reinforcement Learning (DRL), positive rewards promote and negative rewards demote actions.
+In reinforcement learning, positive rewards promote and negative rewards demote actions.
 The agent tries to generate such a policy/knowledge to avoid the discouraging situation by following the policy.
@@
 In particular, the reward signal is not the place to impart to the agent prior knowledge about how to achieve what we want it to do~\cite{sutton2018reinforcement}.
+Taking inspiration from similar work~\cite{Ohi2020}, we define the reward function.
 % TODO add: http://www.incompleteideas.net/book/ebook/node29.html
@@
 To incentivize reduction of $R_e$ (effective reproductive number) the reward is $((R_e^{2}) \times -100) + 100$, this means when $R_e = 1$ there is no reward, but it exponentially decreases when $R_e > 1$ and increases if $R_e < 1$.
-Similarly, for the normalized GDP the reward function is $((\textrm{min\_max\_normalized\_GDP} ** 5)*300) - 80$ which tends to become negative below the average
+Similarly, for the normalized GDP the reward function is $((\textrm{min\_max\_normalized\_GDP}^5) \times 300) - 80$ which tends to become negative somewhere below the average of the $\textrm{min\_max\_normalized\_GDP}$ which is $0.75$.
+Since these are competing costs when added together for the real data they either cancel each other out for the duration of the pandemic and towards the end when both objectives are met they add up together.
+% TODO: add image here
+If the proportion of the infected population were to rise above 0.003 (peak in the actual data) the model is punished ($-2000$) and otherwise rewarded ($+20$).
+To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| \times -20$).
 
-If the proportion of the infected population were to rise above 0.003 the model is punished ($-5000$) and otherwise rewarded ($+20$).
-To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| * -20$).
-
-It should be realized there can be an infinite number of ways to design the reward function to be more human and upgrade the way a decision is taken given the situation \cite{AWSDeepRacer}.
+It should be realized there can be an infinite number of ways to design the reward function to be more human and upgrade the way a decision is taken given the situation~\cite{AWSDeepRacer}.
 Therefore, this research act as a  framework for promoting the development of more efficient reward strategies for the same.
 
-The agent observes the percentage of the population that is susceptible, infected, recovered, and the GDP which extrapolated from the stringency index and all the past moves played, i.e., all the past stringency indexes decided by the agent.
-The following values are fed into a simple network.
-Stable Baselines3 supports multiple inputs by using Dict Gym space.
-This can be done using MultiInputPolicy, which by default uses the CombinedExtractor features extractor to turn multiple inputs into a single vector, handled by the network arch network.
-For data that varies with time (stringency, normalized GDP, $R_e$) we use a simple LSTM architecture \cite{LSTM}.
+The agent observes the percentage of the population that is susceptible, infected, recovered, and time varying data like the GDP, and previous actions taken to change the stringency, and $R_e$.
+Since Stable Baselines3 can support multiple inputs(time-series data, single data points and images) by using Dict Gym space.
+We use the CombinedExtractor features extractor to turn multiple inputs into a single vector, handled by the network arch network.
+For data that varies with time (stringency, normalized GDP, $R_e$) we use a simple LSTM architecture~\cite{LSTM}.
 For other data like, the current proportion of the population that is susceptible, infected, and recovered we use a simple fully connected layer.
@@
 With a time-varying $\nu$ (vaccination rate) and the effect of lockdown, our model is able to account for the infected individuals and reduce the cost in comparison to all the previously formalized models for the data.
-This shows how interventions and changes in the way people behave in response of a epidemic \cite{Caldwell2021} play a major role in the way the epidemic unfolds.
+This shows how interventions and changes in the way people behave in response of a epidemic~\cite{Caldwell2021} play a major role in the way the epidemic unfolds.
 
@@
 The population-level epidemiological data can be obtained from the 'Our World In Data COVID-19' dataset: \url{https://ourworldindata.org/coronavirus} or more specifically: \url{https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-data.csv}~\cite{owidcoronavirus}.
-Data for the total cases, and recovered was acquired by scraping the Worldometers website~\cite{WorldometerCorona} used Internet Archive \cite{InternetArchive}.
+Data for the total cases, and recovered was acquired by scraping the Worldometers website~\cite{WorldometerCorona} used Internet Archive~\cite{InternetArchive}.
 Quaterly GDP data can be obtained from the 'Organisation for Economic Co-operation and Development': \url{https://www.oecd-ilibrary.org/economics/data/main-economic-indicators/main-economic-indicators-complete-database_data-00052-en}~\cite{economic_indicators_data}.

### 2024-02-17 05_reinforcement_run.py

@@ -4,7 +4,9 @@ env = SIREnvironment()
 env.reset()
 
 # models/hpc_run/2667966.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run", "2667966.zip")
+# models/hpc_run/4483170.zip
+# models/hpc_run_2/1225674.zip
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "1225674.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 10

### 2024-02-19 paper_draft/main.tex
changing the reward function

@@
     \item Actions $\altmathcal{A}$: The stringency index variable was analyzed with a sample size of 915.
-The mean value was approximately $61.965049$, with a standard deviation of $17.669831$.
-The minimum value was $31.480000$, while the maximum value reached $96.300000$.
-And the differences between two consecutive stringencies had a mean of $-0.070919$, and standard deviation of $1.427145$, with the minimum being $-14.360000$ and maximum $16.670000$.
+The mean value was approximately $61.96505$, with a standard deviation of $17.66983$.
+The minimum value was $31.48$, while the maximum value reached $96.3$.
+And the differences between two consecutive stringencies had a mean of $-0.070919$, and standard deviation of $1.42715$, with the minimum being $-14.36$ and maximum $16.67$.
 Based on this we define the discrete action space.
@@
 
-Given that at each timestep $t$ the agent has to choose an action $a_t$ so as to maximize the reward $r_{t+1}$ a policy is formulized by the agent.
+Given that at each timestep $t$ the agent has to choose an action $a_t$ to maximize the reward $r_{t+1}$ a policy is formulized by the agent.
 The policy $\pi$ is a mapping from states to a probability distribution over actions: $\pi: \altmathcal{S} \rightarrow p(\altmathcal{A}=\mathbf{a} \mid \altmathcal{S})$.
@@
 Since Stable Baselines3 can support multiple inputs(time-series data, single data points and images) by using Dict Gym space.
-We use the CombinedExtractor features extractor to turn multiple inputs into a single vector, handled by the network arch network.
 For data that varies with time (stringency, normalized GDP, $R_e$) we use a simple LSTM architecture~\cite{LSTM}.

### 2024-02-19 05_reinforcement_run.py

@@ -6,7 +6,7 @@ env.reset()
 # models/hpc_run/2667966.zip
 # models/hpc_run/4483170.zip
 # models/hpc_run_2/1225674.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run_2", "1225674.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "1831656.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 10

### 2024-02-19 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -50,13 +50,24 @@ def deriv(y, t, N, beta, gamma, nu_varying, lockdown):
     dRdt = gamma * I + nu_varying[int(t)] * S
     return dSdt, dIdt, dRdt
 
+# def calculate_reward_weighted(gdp_min_max_normalized_list, r_eff_list):
+#     reward_list = []
+#     for gdp, r_eff in zip(gdp_min_max_normalized_list, r_eff_list):
+#         gdp_reward = ((gdp ** 5)*300) - 80
+#         r_eff_reward = ((r_eff**2)*-100) + 100
+#         total_reward = gdp_reward + r_eff_reward
+#         reward_list.append(total_reward)
+#     return reward_list
+
 def calculate_reward_weighted(gdp_min_max_normalized_list, r_eff_list):
     reward_list = []
     for gdp, r_eff in zip(gdp_min_max_normalized_list, r_eff_list):
-        gdp_reward = ((gdp ** 5)*300) - 80
-        r_eff_reward = ((r_eff**2)*-100) + 100
-        total_reward = gdp_reward + r_eff_reward
-        reward_list.append(total_reward)
+        if r_eff > 1.5:
+            reward_list.append(-20 * r_eff)
+        elif r_eff >= 1.25 and r_eff <= 1.5:
+            reward_list.append(100 * gdp)
+        else:
+            reward_list.append(200 * gdp)
     return reward_list
 
 class SIREnvironment(gym.Env):
@@ -123,12 +134,18 @@ class SIREnvironment(gym.Env):
         # REMEMBER: to change this definition of the reward in the render as well!!!
         # self.reward = self.normalized_GDP_min_max_normalized - (2 * self.r_eff)
 
-        reward_inertia= abs(diff_action)*-1*20
+        reward_inertia= abs(diff_action)*-1*8
         reward_I_percentage = -2000 if self.I_proportion >= 0.003 else 50
         
-        gdp_reward = ((self.normalized_GDP_min_max_normalized**5)*300) - 80
-        r_eff_reward = ((self.r_eff**2)*-100) + 100
-        reward_weighted = gdp_reward + r_eff_reward
+        # gdp_reward = ((self.normalized_GDP_min_max_normalized**5)*300) - 80
+        # r_eff_reward = ((self.r_eff**2)*-100) + 100
+        # reward_weighted = gdp_reward + r_eff_reward
+        if self.r_eff > 1.5:
+            reward_weighted = -20 * self.r_eff
[…]

### 2024-02-19 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py
changing the reward function

@@ -209,10 +209,10 @@ class SIREnvironment(gym.Env):
         I_reward_moves = [hospital_capacity_punishment if I_percentage >= hospital_capacity else hospital_capacity_reward for I_percentage in self.df["I_moves"] / N]
 
         
-        inertia_rewards_actual = np.array([0] + [abs(diff)*20*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
+        inertia_rewards_actual = np.array([0] + [abs(diff)*8*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
         # modelled reward for intertia is same as actual
-        inertia_rewards_modelled = np.array([0] + [abs(diff)*20*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
-        inertia_rewards_moves = np.array([0] + [abs(diff)*20*-1 for diff in (stringency[i] - stringency[i - 1] for i in range(1, len(stringency)))])
+        inertia_rewards_modelled = np.array([0] + [abs(diff)*8*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
+        inertia_rewards_moves = np.array([0] + [abs(diff)*8*-1 for diff in (stringency[i] - stringency[i - 1] for i in range(1, len(stringency)))])
 
         reward_actual = np.array(calculate_reward_weighted(self.df["gdp_min_max_normalized"], self.df["r_eff_actual_" + modelling_type])) + I_reward_actual + inertia_rewards_actual
         reward_modelled = np.array(calculate_reward_weighted(self.df["gdp_normalized_modelled_min_max_normalized"], self.df["r_eff_modelled_" + modelling_type])) + I_reward_modelled + inertia_rewards_modelled

### 2024-02-20 05_reinforcement_run.py
inertia factor changed from 8 to 12

@@ -6,7 +6,7 @@ env.reset()
 # models/hpc_run/2667966.zip
 # models/hpc_run/4483170.zip
 # models/hpc_run_2/1225674.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run_2", "1831656.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "2141502.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 10

### 2024-02-20 rl_utils_new_complex_reward_sirvlockdown_nu_varying.py

@@ -134,7 +134,7 @@ class SIREnvironment(gym.Env):
         # REMEMBER: to change this definition of the reward in the render as well!!!
         # self.reward = self.normalized_GDP_min_max_normalized - (2 * self.r_eff)
 
-        reward_inertia= abs(diff_action)*-1*8
+        reward_inertia= abs(diff_action)*-1*12
         reward_I_percentage = -2000 if self.I_proportion >= 0.003 else 50
         
         # gdp_reward = ((self.normalized_GDP_min_max_normalized**5)*300) - 80
@@ -209,10 +209,10 @@ class SIREnvironment(gym.Env):
         I_reward_moves = [hospital_capacity_punishment if I_percentage >= hospital_capacity else hospital_capacity_reward for I_percentage in self.df["I_moves"] / N]
 
         
-        inertia_rewards_actual = np.array([0] + [abs(diff)*8*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
+        inertia_rewards_actual = np.array([0] + [abs(diff)*12*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
         # modelled reward for intertia is same as actual
-        inertia_rewards_modelled = np.array([0] + [abs(diff)*8*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
-        inertia_rewards_moves = np.array([0] + [abs(diff)*8*-1 for diff in (stringency[i] - stringency[i - 1] for i in range(1, len(stringency)))])
+        inertia_rewards_modelled = np.array([0] + [abs(diff)*12*-1 for diff in (self.df['stringency_index'][i] - self.df['stringency_index'][i - 1] for i in range(1, len(self.df)))])
+        inertia_rewards_moves = np.array([0] + [abs(diff)*12*-1 for diff in (stringency[i] - stringency[i - 1] for i in range(1, len(stringency)))])
 
         reward_actual = np.array(calculate_reward_weighted(self.df["gdp_min_max_normalized"], self.df["r_eff_actual_" + modelling_type])) + I_reward_actual + inertia_rewards_actual
         reward_modelled = np.array(calculate_reward_weighted(self.df["gdp_normalized_modelled_min_max_normalized"], self.df["r_eff_modelled_" + modelling_type])) + I_reward_modelled + inertia_rewards_modelled

### 2024-02-24 paper_draft/main.tex
paper with references

@@
 This approach also enhances transparency in governmental decision-making by establishing a well-defined reward function for the reinforcement learning agent.
-In essence, this study introduces an innovative and ethical strategy to navigate the intricate challenge of balancing public health and economic stability amidst infectious disease outbreaks.
+In essence, this study introduces an innovative and ethical strategy to navigate the challenge of balancing public health and economic stability amidst infectious disease outbreaks.
 \end{abstract}
@@
 By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we effectively model the disease progression.
-Our model (SIRV with lockdown and time-varying vaccination rate) in addition to the SIR model, accounts for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{Hale2021,SIRLockdown,NBERw26981,NBERw26867}.
-Eventhough the traditional SIR model is a valuable tool for understanding the spread of infectious diseases, it assumes that parameters like the transmission rate ($\beta$) and recovery rate ($\gamma$) are constant over time, which may not always be the case.
+Our model (SIRV with lockdown and time-varying vaccination rate) builds on the foundational SIR model, by accounting for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{Hale2021,SIRLockdown,NBERw26981,NBERw26867}.
+Although the traditional SIR model is a valuable tool for understanding the spread of infectious diseases, it assumes that parameters like the transmission rate ($\beta$) and recovery rate ($\gamma$) are constant over time, which may not always be the case.
 In this paper, we propose a more sophisticated approach by introducing a time-dependent SIR model~\cite{TimeDependentSIR}, enabling us to account for the changing dynamics of the pandemic due to factors such as lockdowns and vaccination rates.
@@
 Deep reinforcement learning is an advancement to reinforcement learning which helps normalize the input and reduce its dimensionality~\cite{Lavet2018,Arulkumaran2017,Henderson_Islam_Bachman_Pineau_Precup_Meger_2018,Mnih2015}.
-We use a Long Short Term Memory recurrent neural network for time-series data~\cite{NIPS2001_a38b1617,LSTM} and a simple fully connected network for the non-time-series data.
+We use a Long Short Term Memory recurrent neural network for time-series data~\cite{NIPS2001_a38b1617,LSTM} and a simple fully connected network for data points that don't vary with time.
 
@@
 \section{Mathematical Formulation and Numerical Computation}
-In SIRV-RL, we use a compartmental model to model infectious disease environment.
-We iteratively develop this model to fit the actual data better.
+In this paper, we use a compartmental model to model infectious disease environment.
+We iteratively develop this model, starting with the foundational SIR model, to fit the actual data better.
 In an SIR model people of the population are divided based on whether they are yet to come into contact with an infected person (Susceptible), are infectious themselves (Infectious), or have recovered from the infection (Recovered).
@@
 Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered.
-Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown} , we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data.
+Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown}, we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data.
 The following parameters and loss is obtained:

### 2024-02-24 05_reinforcement_run.py

@@ -6,10 +6,10 @@ env.reset()
 # models/hpc_run/2667966.zip
 # models/hpc_run/4483170.zip
 # models/hpc_run_2/1225674.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run_2", "2141502.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "786954.zip")
 model = PPO.load(model_path, env=env)
 
-episodes = 10
+episodes = 30
 for episode in range(1, episodes+1):
     obs, info = env.reset()
     terminated = False

### 2024-03-03 05_reinforcement_run.py
recompiled paper

@@ -6,7 +6,7 @@ env.reset()
 # models/hpc_run/2667966.zip
 # models/hpc_run/4483170.zip
 # models/hpc_run_2/1225674.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run_2", "786954.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "2210052.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 30

### 2024-03-06 paper_draft/main.tex
recompiled paper
paper edits: fig captions, sirv to sir with v

@@
 
-\title{SIRV-RL: Reinforcement Learning for Optimized Policy Control during Epidemiological Outbreaks in Emerging Market and Developing Economies}
+\title{SIR-RL: Reinforcement Learning for Optimized Policy Control during Epidemiological Outbreaks in Emerging Market and Developing Economies}
 \author[1,*]{Maeghal Jain}
@@
 By lesioning the model, as opposed to proposing a new mathematical model with more specialized compartments to more accurately represent the actual environment~\cite{Bjrnstad2020, Mwalili2020}, we effectively model the disease progression.
-Our model (SIRV with lockdown and time-varying vaccination rate) builds on the foundational SIR model, by accounting for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{Hale2021,SIRLockdown,NBERw26981,NBERw26867}.
+Our model (SIR with lockdown and time-varying vaccination rate) builds on the foundational SIR model, by accounting for the recovery reached through vaccination~\cite{Marinov2022,MaurciodeCarvalho2023,Thater,Turkyilmazoglu2022,YALADANDA2022101052} and the effects of lockdown~\cite{Hale2021,SIRLockdown,NBERw26981,NBERw26867}.
 Although the traditional SIR model is a valuable tool for understanding the spread of infectious diseases, it assumes that parameters like the transmission rate ($\beta$) and recovery rate ($\gamma$) are constant over time, which may not always be the case.
@@
 However, the study has limitations; First, the deterministic SIR model (predecessor to our proposed model) fails to account for chance in disease spread and lacks confidence intervals on results and while stochastic models incorporate chance, they are typically more challenging to analyze than their deterministic counterparts~\cite{Hethcote2008}.
-Second, the underreporting of cases during the period selected by our study.
-Lastly, the reinforcement learning agent should be resistant to how the vaccination rate changes -- or to when it is first introduced because then it scopes the environment for succumbing to wishful thinking which can be  potentially dangerous.
+Secondly, the underreporting of cases during the period selected by our study.
+Lastly, the reinforcement learning agent should be resistant to how the vaccination rate changes, and different values for $\beta$ and $\gamma$ -- keeping them the same scopes the environment for succumbing to wishful thinking which can be potentially dangerous.
+Therefore, before an actual deployment of the model, it would be a good measure to introduce stochasticity to these parameters($\beta$ and $\gamma$) and the vaccination rate.
 
@@
 \end{equation}
-During an epidemic, the fundamental drivers of an epidemic growth is the rate of infection $\beta$ i.e.
-the average number of infections per infected case and the infectious period $1/\gamma$ i.e.
-the average period for which the infected case is infected for.
+During an epidemic, the fundamental drivers of an epidemic growth is the rate of infection $\beta$ i.e., the average number of infections per infected case and the infectious period $1/\gamma$ i.e., the average period for which the infected case is infected for.
 Epidemics can only happen if the case is infectious enough for long enough and this defined by $R_0 = \beta / \gamma$.
@@
 We generally calibrate the model using \cref{eq:cost_SIR}.
-However, sometimes we must strike a balance between modeling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}, where a hyperparameter is selected keeping both these losses in mind).
+However, sometimes we must strike a balance between modelling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}, where a hyperparameter is selected keeping both these losses in mind).
 \begin{equation}
@@
 Where, $S$ is the number of susceptible people and $\hat{S}$ is the predicted number of susceptible people, similarly, $I$ for infected and $\hat{I}$ for the predicted number of infected people, and $R$ for recovered.
-Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown}, we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data.
+Using the equations \Crefrange{eq:S_without_lockdown}{eq:R_without_lockdown}, we minimize the cost function \cref{eq:cost_SIR_without_lockdown} using the Nelder-Mead method~\cite{Gao2010} to estimate the parameters like $\beta$, $\gamma$, and fit the model to actual data.
 The following parameters and loss is obtained:
@@
 The stringency index is a composite measure based on nine response indicators including school closures, workplace closures, and travel bans, rescaled to a value from 0 to 100 (100 = strictest)~\cite{owidcoronavirus}.
-This index simply records the strictness of government policies and does not measure or imply the appropriateness or effectiveness of a country's response i.e.
-a higher score does not necessarily mean that a country's response is "better" than others lower on the index.
+This index simply records the strictness of government policies and does not measure or imply the appropriateness or effectiveness of a country's response i.e., a higher score does not necessarily mean that a country's response is "better" than others lower on the index.
+
 To define the new time-varying beta that is dependent on the current stringency index, the following equations have been formulated:
@@
 \end{equation}
-\subsection{SIRV Model with Lockdown}
+\subsection{SIR Model with Lockdown and Vaccination}
 Lastly, an additional flow from the susceptible to recovered population can be shown by adding a vaccination rate $\nu$ in the model. 
@@
 \subsection{Optimizing Window Length for Time-varying Vaccination Rate}\label{optimizing_window_length}
-However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIRV_model_with_lockdown_infections_IND} suggests that $\nu$ might be varying with time.
+However, as observed by the value of $\nu_{optimal}$ from \cref{eq:nu_optimal_with_lockdown_and_nu} which is almost negligible and the overestimation of infected individuals in \cref{fig:SIR_model_with_lockdown_with_vaccination_infections_IND} suggests that $\nu$ might be varying with time.
 This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}.
@@
 \begin{figure}[htbp!]
-  \caption{Loss for Different Window Lengths}
   \begin{subfigure}[t]{0.48\textwidth}
@@
   \end{subfigure}
+  \caption{\textbf{Loss for Different Window Lengths}.
+Here we change the window length to see which window length yields the best loss in both cases, while predicting the susceptible, infected and recovered populations or just the infected population.}
 \end{figure}
 
-The following graph indicates that a window length of $10$ days yields the least overall loss for all three population groups.
-However, this window length results in a poor approximation for the infected group, which is crucial for accurately modeling the spread of the disease.
+The following results in \Crefrange{fig:window_length_loss_SIR_IND}{fig:window_length_loss_I_IND}, indicate that a window length of $10$ days yields the least overall loss for all three population groups.
+However, this window length results in a poor approximation for just the infected group, which is crucial for accurately modelling the spread of the disease.
 Consequently, we have decided to use a window length of $15$ days, which provides a more accurate approximation for the infected population while still maintaining reasonable loss for the other groups.
@@
   \centering
-  \caption{$\nu$ Varying with Time}
   \includegraphics[width=\linewidth]{images/interpolated_nu_varying_with_time_IND.pdf}
   \label{fig:interpolated_nu_varying_with_time_IND}
+  \caption{\textbf{$\nu$ Varying with Time.} This shows how the vaccination rate $\nu$ varies with time and when the vaccination drive was launched in India.}
 \end{figure}
@@
 This shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
-Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIRV Model with time varying $\nu$.
+Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIR Model with lockdown and time-varying $\nu$.
 
-\subsection{SIRV Model with Lockdown (Time-varying Vaccination Rate)}
+\subsection{SIR Model with Lockdown and Time-varying Vaccination Rate}
 Finally, we integrate the time-varying vaccination rate ($\nu$) into the SIR model that includes lockdown measures, resulting in the following set of equations, which represents our final model:
@@
 \end{equation}
-See \Crefrange{fig:SIR_model_IND}{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND} and \Crefrange{fig:comparing_costs_SIR_IND}{fig:comparing_costs_I_IND} to see how the different models compare against each other.
+See \Crefrange{fig:SIR_model_IND}{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND} and \Crefrange{fig:comparing_costs_SIR_IND}{fig:comparing_costs_I_IND} to see how the different models compare against each other.
 
@@
 Increasing the stringency decreases the spread of the infection.
-Conversely, it must be noted that herd immunity can only be achieved when the epidemic reaches its peak i.e.
-when the effective reproductive number is equal to one ($R_e = 1$).
+Conversely, it must be noted that herd immunity can only be achieved when the epidemic reaches its peak i.e., when the effective reproductive number is equal to one ($R_e = 1$).
 This can only happen by lowering the stringency index which would allow the natural dynamics of the epidemic to transpire such that the population of susceptible individuals has depleted enough such that it is insufficient to propagate the disease further.
@@
 
-The agent observes the percentage of the population that is susceptible, infected, recovered, and time varying data like the GDP, and previous actions taken to change the stringency, and $R_e$.
+The agent observes the percentage of the population that is susceptible, infected, recovered, and time-varying data like the GDP, and previous actions taken to change the stringency, and $R_e$.
 Since Stable Baselines3 can support multiple inputs(time-series data, single data points and images) by using Dict Gym space.
@@
   \centering
-  \caption{SIR Model with lockdown for India}
   \begin{subfigure}[t]{\textwidth}
@@
     \includegraphics[scale=0.5]{images/SIR_model_IND.pdf}
-    \caption{SIR model with lockdown}
+    \caption{SIR Model with Lockdown}
     \label{fig:SIR_model_IND}
@@
     \includegraphics[scale=0.50]{images/SIR_model_infections_IND.pdf}
-    \caption{Infections modelled with SIR model}
+    \caption{Infections Modelled with SIR Model}
     \label{fig:SIR_model_infections_IND}
   \end{subfigure}
+  \caption{\textbf{SIR Model for India.} The following shows the fit for the simple SIR model(\Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown}) compared to the real data.
+Here we can see an over approximation of the infected population.}
 \end{figure}
@@
   \centering
-  \caption{SIR Model with lockdown for India}
-
   \begin{subfigure}[t]{\textwidth}
@@
     \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_IND.pdf}
-    \caption{SIR model with lockdown}
+    \caption{SIR Model with Lockdown}
     \label{fig:SIR_model_with_lockdown_IND}
@@
     \includegraphics[scale=0.50]{images/stringency_varying_with_time_IND.pdf}
-    \caption{Stringency varying with Time}
+    \caption{Stringency Varying with Time}
     \label{fig:stringency_varying_with_time_IND}
@@
     \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_infections_IND.pdf}
-    \caption{Infections modelled with SIR model with Lockdown}
+    \caption{Infections Modelled with SIR Model with Lockdown}
     \label{fig:SIR_model_with_lockdown_infections_IND}
   \end{subfigure}
+  \caption{\textbf{SIR Model with lockdown for India.} The following shows the fit for the SIR model with lockdown(\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}) compared to the real data.
+Here we observe how introducing stringency shows an effect on the dynamics of the disease progression.
+There's still an over approximation, but the peaks of the model align with the data.}
 \end{figure}
@@
   \centering
-  \caption{SIRV Model with lockdown for India}
-
   \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_IND.pdf}
-    \caption{SIRV model with lockdown}
-    \label{fig:SIRV_model_with_lockdown_IND}
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_IND.pdf}
+    \caption{SIR Model with Lockdown and Vaccination}
+    \label{fig:SIR_model_with_lockdown_with_vaccination_IND}
   \end{subfigure}
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_IND.pdf}
-    \caption{Infections modelled with SIRV model with Lockdown}
-    \label{fig:SIRV_model_with_lockdown_infections_IND}
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_infections_IND.pdf}
+    \caption{Infections Modelled with SIR Model with Lockdown and Vaccination}
+    \label{fig:SIR_model_with_lockdown_with_vaccination_infections_IND}
   \end{subfigure}
+  \caption{\textbf{SIR Model with Lockdown and Vaccination for India.} The following shows the fit for the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}) compared to the real data.
+This is similar to the infections shown by the SIR model with lockdown as seen in \cref{fig:SIR_model_with_lockdown_infections_IND} and this is because the rate of vaccination is negligible \cref{eq:nu_optimal_with_lockdown_and_nu}.
+This is suggestive of a rate of vaccination that varies with time.}
 \end{figure}
@@
 
-For time-varying $\nu$ and SIRV model with lockdown from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following:
+For SIR model with lockdown and time-varying vaccination rate from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following:
 \begin{figure}[htbp!]
   \centering
-  \caption{SIRV Model with lockdown and time-varying $\nu$ for India}
-
   \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_time_varying_nu_IND.pdf}
-    \caption{SIRV model with lockdown and time-varying $\nu$}
-    \label{fig:SIRV_model_with_lockdown_time_varying_nu_IND}
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_time_varying_nu_IND.pdf}
+    \caption{SIR Model with Lockdown and Time-varying Vaccination Rate}
+    \label{fig:SIR_model_with_lockdown_with_vaccination_time_varying_nu_IND}
   \end{subfigure}
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIRV_model_with_lockdown_infections_time_varying_nu_IND.pdf}
-    \caption{Infections modelled with SIRV model with Lockdown and time-varying $\nu$}
-    \label{fig:SIRV_model_with_lockdown_infections_time_varying_nu_IND}
+    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_time_varying_nu_infections_IND.pdf}
+    \caption{Infections Modelled with SIR Model with Lockdown and Time-varying Vaccination Rate}
+    \label{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND}
   \end{subfigure}
@@
 	\end{subfigure}
+  \caption{\textbf{SIR Model with Lockdown and Time-varying Vaccination Rate for India.} The following shows the fit for the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}) compared to the real data.
+This model captures the trends in infected population accurately.}
 \end{figure}
@@
   \centering
-  \caption{Stringency and GDP for Developing Economies}
   \begin{subfigure}[t]{0.48\textwidth}
@@
   \label{fig:gdp_modelled_with_stringency_BRA}
+  \caption{\textbf{Stringency and GDP for Developing Economies.} For countries with developing economies, when we model the normalized GDP with stringency we see significant p-values and high r2 scores.}
 \end{figure}
@@
   \centering
-  \caption{Stringency and GDP for Advanced Economies}
   \begin{subfigure}[t]{0.48\textwidth}
@@
   \label{fig:gdp_modelled_with_stringency_CAN}
+  \caption{\textbf{Stringency and GDP for Advanced Economies.} In economically advanced nations, when modeling the normalized GDP against stringency measures, we consistently observe substantial p-values, providing robust evidence against the null hypothesis.
+However, the significance levels are not as high as those found for developing economies(\Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA}).
+This is reflected in the lower R-squared scores, which indicates that the relationship between these variables may be less pronounced when compared to developing economies.}
 \end{figure}
-From this we conclude, that maybe there are other factors that explain the dip in the normalized GDP than just an increase in the stringency.
+From \Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_CAN} we conclude, that there are other factors that explain the dip in the normalized GDP than just an increase in the stringency in advanced economies.
 

### 2024-03-06 05_reinforcement_run.py

@@ -6,7 +6,7 @@ env.reset()
 # models/hpc_run/2667966.zip
 # models/hpc_run/4483170.zip
 # models/hpc_run_2/1225674.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run_2", "2210052.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "2643288.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 30

### 2024-03-06 paper_draft/main.tex
paper edits

@@
   \end{subfigure}
-  \caption{\textbf{Loss for Different Window Lengths}.
-Here we change the window length to see which window length yields the best loss in both cases, while predicting the susceptible, infected and recovered populations or just the infected population.}
+  \caption{\textbf{Loss for Different Window Lengths.} Here we change the window length to see which window length yields the best loss in both cases, while predicting the susceptible, infected and recovered populations or just the infected population.}
 \end{figure}
@@
   \centering
-  \caption{Strategy from Reinforcement Learning Agent 1}
   \begin{subfigure}[t]{0.48\textwidth}
     \centering
-    \includegraphics[width=\linewidth]{images/914857/rl_sir.eps}
-    \caption{SIR Dynamics}
+    \includegraphics[width=\linewidth]{images/176647/rl_stringency.pdf}
+    \caption{Stringency changing over Time}
   \end{subfigure}
-  \label{fig:914857_rl_sir}
+  \label{fig:176647_rl_stringency}
   \hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{images/914857/rl_i.eps}
-    \caption{Infected Population changing over Time}
+    \includegraphics[width=\linewidth]{images/176647/rl_sir.pdf}
+    \caption{SIR Dynamics}
   \end{subfigure}
-  \label{fig:914857_rl_i}
-
-  \begin{subfigure}[t]{0.48\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{images/914857/rl_gdp.eps}
-    \caption{Normalized GDP changing over Time}
-  \end{subfigure}
-  \label{fig:914857_rl_gdp}
+  \label{fig:176647_rl_sir}
   \hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{images/914857/rl_stringency.pdf}
-    \caption{Stringency changing over Time}
+    \includegraphics[width=\linewidth]{images/176647/rl_i.pdf}
+    \caption{Infected Population changing over Time}
   \end{subfigure}
-  \label{fig:914857_rl_stringency}
-
-  \begin{subfigure}[t]{0.48\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{images/914857/rl_r_eff.eps}
-    \caption{$R_e$ changing over Time}
-  \end{subfigure}
-  \label{fig:914857_rl_r_eff}
+  \label{fig:176647_rl_i}
   \hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{images/914857/rl_reward.eps}
+    \includegraphics[width=\linewidth]{images/176647/rl_gdp.pdf}
+    \caption{Normalized GDP changing over Time}
+  \end{subfigure}
+  \label{fig:176647_rl_gdp}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/176647/rl_r_eff.pdf}
+    \caption{$R_e$ changing over Time}
+  \end{subfigure}
+  \label{fig:176647_rl_r_eff}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/176647/rl_reward.pdf}
     \caption{Reward changing over Time}
   \end{subfigure}
-  \label{fig:914857_rl_reward}
+  \label{fig:176647_rl_reward}
+  \caption{\textbf{Strategy from Reinforcement Learning Agent.} Here, "(actual)" is the real data, and "(rl)" is the new strategy we propose.
+\textbf{(a)} The new strategy proposed highlights a decrease in stringency from July, 2020 till October, 2020 and after which there's an increase and then a steady decline towards the end.
+\textbf{(b)} SIR Dynamics, no observable differences.
+\textbf{(c)} }
 \end{figure}
@@
 A strategic decision is made by the agent to maintain the stringency index below 80 throughout the course of the study.
-This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to $1.2$, significantly ahead of the projected timeline (refer to figure \cref{fig:914857_rl_r_eff}).
+This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to $1.2$, significantly ahead of the projected timeline (refer to figure \cref{fig:176647_rl_r_eff}).
 A subsequent decrease in stringency in the later stages leads to an increase in the normalized Gross Domestic Product (GDP), indicating an economic upturn.
@@
     \centering
-    \includegraphics[width=\linewidth]{images/244620/rl_sir.eps}
+    \includegraphics[width=\linewidth]{images/175975/rl_sir.pdf}
     \caption{SIR Dynamics}
   \end{subfigure}
-  \label{fig:244620_rl_sir}
+  \label{fig:175975_rl_sir}
   \hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{images/244620/rl_i.eps}
+    \includegraphics[width=\linewidth]{images/175975/rl_i.pdf}
     \caption{Infected Population changing over Time}
   \end{subfigure}
-  \label{fig:244620_rl_i}
+  \label{fig:175975_rl_i}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{images/244620/rl_gdp.eps}
+    \includegraphics[width=\linewidth]{images/175975/rl_gdp.pdf}
     \caption{Normalized GDP changing over Time}
   \end{subfigure}
-  \label{fig:244620_rl_gdp}
+  \label{fig:175975_rl_gdp}
   \hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{images/244620/rl_stringency.pdf}
+    \includegraphics[width=\linewidth]{images/175975/rl_stringency.pdf}
     \caption{Stringency changing over Time}
   \end{subfigure}
-  \label{fig:244620_rl_stringency}
+  \label{fig:175975_rl_stringency}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{images/244620/rl_r_eff.eps}
+    \includegraphics[width=\linewidth]{images/175975/rl_r_eff.pdf}
     \caption{$R_e$ changing over Time}
   \end{subfigure}
-  \label{fig:244620_rl_r_eff}
+  \label{fig:175975_rl_r_eff}
   \hfill
@@
     \centering
-    \includegraphics[width=\linewidth]{images/244620/rl_reward.eps}
+    \includegraphics[width=\linewidth]{images/175975/rl_reward.pdf}
     \caption{Reward changing over Time}
   \end{subfigure}
-  \label{fig:244620_rl_reward}
+  \label{fig:175975_rl_reward}
 \end{figure}
@@
 This strategy results in the majority of infections occurring prior to the vaccine's release, with a subsequent cessation of new infections.
-This approach, however, has implications for the normalized Gross Domestic Product (GDP), as seen by the observed decline in figure \cref{fig:244620_rl_gdp}.
+This approach, however, has implications for the normalized Gross Domestic Product (GDP), as seen by the observed decline in figure \cref{fig:175975_rl_gdp}.
 Furthermore, the effective reproduction number ($R_e$) does not achieve the target value of 1.2 under this strategy.

### 2024-03-06 02_data_calibration_final.py

@@ -1,0 +1,1510 @@
+import numpy as np
+
+from scipy.integrate import odeint
+from scipy.optimize import minimize
+from scipy.stats import poisson
+from scipy import spatial
+
+from scipy import stats
+
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.ticker import LogFormatterSciNotation
+from matplotlib.ticker import FuncFormatter
+
+import pandas as pd
+import os
+import seaborn as sns
+import json
+from datetime import date, timedelta
+import math
+np.random.seed(seed=42)
+
+# {'dejavusans':'ok', 'dejavuserif':'not_tried', 'cm':'no', 'stix':'no', 'stixsans':'no', 'custom':'best'}
+matplotlib.rcParams['mathtext.fontset'] = 'custom'
+matplotlib.rcParams['font.family'] = 'STIXGeneral'
+
+# plt.rcParams['text.usetex'] = True
+# plt.rcParams["font.family"] = "Times New Roman"
+# sns.set_style("whitegrid", rc={'font.family': 'Times New Roman', 'font.size': 16})
+plt.rc('legend',fontsize=16)
+plt.rcParams.update({'font.size': 16, 'axes.labelsize': 14, 'axes.titlesize': 14, 'xtick.labelsize': 14, 'ytick.labelsize': 14})
+sns.set_style("whitegrid", rc={'font.size': 16})
+
+from constants import LOCATION_CHOOSEN, LOCATION_CHOOSEN_2, OUTPUT_DIR, DATA_CACHE_DIR, OPTIMAL_VALUES_FILE, STRINGENCY_BASED_GDP
+
+
+
+data_path = os.path.join(DATA_CACHE_DIR, LOCATION_CHOOSEN + "_merged_data.csv")
+if os.path.exists(data_path):
+    df = pd.read_csv(data_path)
+else:
+    df = pd.read_excel("owid-covid-data.xlsx")
+    df = df.loc[df['iso_code'] == LOCATION_CHOOSEN]
+    df.to_csv(data_path, index=False)
+
+
+
+start_date = date(2020, 5, 1) # -> works, but not for GDP
+# start_date = date(2020, 2, 10) -> doesn't work
+# start_date = date(2020, 3, 1) -> doesn't work for SIR with lockdown
+# start_date = date(2020, 3, 15) -> doesn't work for SIR with lockdown
+# start_date = date(2020, 4, 1) -> doesn't work for SIR with lockdown
[…]

### 2024-03-09 paper_draft/main.tex
paper edits

@@
 To estimate the parameters $\beta$ and $\gamma$ for India, based on the data from May, 2020, to October, 2022, we simply define two cost functions \cref{eq:cost_SIR,eq:cost_I} to calibrate the model with the use of huber loss~\cite{huberloss}.
-We generally calibrate the model using \cref{eq:cost_SIR}.
-However, sometimes we must strike a balance between modelling all three population groups (susceptible, infected, and recovered) and focusing just on the infected group, which is most important because it drives the spread of the disease, therefore, in that case we consider the loss obtained from \cref{eq:cost_SIR,eq:cost_I} (see \cref{optimizing_window_length}, where a hyperparameter is selected keeping both these losses in mind).
+Our model typically uses \cref{eq:cost_SIR}, considering all three compartments: susceptible, infected, and recovered.
+However, there are instances where we need to balance modeling all groups and focusing on the infected group, which largely drives disease spread.
+In such cases, we consider losses from both \cref{eq:cost_SIR,eq:cost_I}.
+A weighted sum of the loss functions allows for trade-offs between comprehensive modeling and focusing on infected group dynamics(see \cref{optimizing_window_length}, where a hyperparameter(window length) is selected keeping both these losses in mind).
 \begin{equation}
@@
   \end{subfigure}
-  \caption{\textbf{Loss for Different Window Lengths.} Here we change the window length to see which window length yields the best loss in both cases, while predicting the susceptible, infected and recovered populations or just the infected population.}
+  \caption{\textbf{Loss for Different Window Lengths.} We try different window lengths to find the optimal loss for both cases, either when predicting all three populations(susceptible, infected, recovered) or just the infected population.}
 \end{figure}
@@
   \label{fig:interpolated_nu_varying_with_time_IND}
-  \caption{\textbf{$\nu$ Varying with Time.} This shows how the vaccination rate $\nu$ varies with time and when the vaccination drive was launched in India.}
+  \caption{\textbf{$\nu$ Varying with Time.} This depicts how the vaccination rate($\nu$) changes over time and highlights the introduction of the vaccination campaign in India.}
 \end{figure}
 
-This shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
+This \cref{fig:interpolated_nu_varying_with_time_IND} shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
 Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIR Model with lockdown and time-varying $\nu$.
@@
 Given that the government is an agent that takes decisions in a deterministic environment defined above, we use reinforcement learning to model the competing costs of the environment.
-This environment is known as a Markov decision process (MDP), and can be described as follows:
+This environment is known as a Markov decision process (MDP) and is characterized by the Markov property.
+To possess the Markov property is to create a compact state signal that retains all relevant information from past sensations without requiring the complete history.
+The Markov property ensures that the probability of transitioning to the next state and receiving a reward depends only on the current state and action, without requiring the entire history~\cite{sutton2018reinforcement}.
+Our MDP is defined as follows:
 % https://arxiv.org/pdf/1708.05866.pdf
@@
 The reward function is parameterized to account for key factors influencing decision-making.
-To incentivize reduction of $R_e$ (effective reproductive number) the reward is $((R_e^{2}) \times -100) + 100$, this means when $R_e = 1$ there is no reward, but it exponentially decreases when $R_e > 1$ and increases if $R_e < 1$.
-Similarly, for the normalized GDP the reward function is $((\textrm{min\_max\_normalized\_GDP}^5) \times 300) - 80$ which tends to become negative somewhere below the average of the $\textrm{min\_max\_normalized\_GDP}$ which is $0.75$.
-Since these are competing costs when added together for the real data they either cancel each other out for the duration of the pandemic and towards the end when both objectives are met they add up together.
-% TODO: add image here
-If the proportion of the infected population were to rise above 0.003 (peak in the actual data) the model is punished ($-2000$) and otherwise rewarded ($+20$).
-To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| \times -20$).
+To incentivize reduction of $R_e$(effective reproductive number) and the increase in the normalized GDP the reward is:
+$$\[
+\textrm{Reward} = \begin{cases}
+-20 \times R_e & \textrm{if } R_e > 1.5 \\
+100 \times \textrm{min\_max\_normalized\_GDP} & \textrm{if } 1.25 \leq R_e \leq 1.5 \\
+200 \times \textrm{min\_max\_normalized\_GDP} & \textrm{if } R_e < 1.25
+\end{cases}
+\]$$
+
+This reward function is parameterized to account for key factors influencing decision-making.
+When the effective reproductive number ($R_e$) exceeds 1.5, indicating a high transmission rate of the disease, the reward is negatively impacted to incentivize a reduction in $R_e$.
+As $R_e$ decreases within the range $1.25 \leq R_e \leq 1.5$, indicating a moderate transmission rate, the reward is directly proportional to the normalized GDP, reflecting the importance of both controlling the spread of the disease and maintaining economic stability.
+Notably, when $R_e$ drops below 1, signaling a declining transmission rate and potential containment of the disease, the reward function shifts focus towards economic recovery.
+In this scenario, the reward incentivizes an increase in the normalized GDP, emphasizing the need to stimulate economic activity and promote recovery efforts following successful control measures.
+
+Additionally, if the proportion of the infected population were to rise above 0.003(peak in the actual data) the model is punished ($-2000$) and otherwise rewarded ($50$).
+To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| \times -12$).
 
@@
 
-We train the model for $2742$ time steps for $200$ iterations and the best results are presented below.
 \newcommand{\DrawNeuronalNetwork}[2][]{
@@
 \begin{tikzpicture}[x=1.5cm, y=1.5cm,
->=stealth,font=\sffamily,nodes={align=center}, scale=0.8, every node/.style={scale=0.8}]
+>=stealth,font=\sffamily,nodes={align=center}]
  \begin{scope}[local bounding box=T]
@@
 \end{tikzpicture}
+
+We train the model for $2742$ time steps and some of the best results are presented below.
 
@@
   \end{subfigure}
-  \caption{\textbf{SIR Model for India.} The following shows the fit for the simple SIR model(\Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown}) compared to the real data.
-Here we can see an over approximation of the infected population.}
+  \caption{\textbf{SIR Model Comparison for India.} The figure presents a comparison between the fitted simple SIR model(\Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown}) and real data.
+Here an evident overestimation of the infected population is observed.}
 \end{figure}
@@
   \end{subfigure}
-  \caption{\textbf{SIR Model with lockdown for India.} The following shows the fit for the SIR model with lockdown(\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}) compared to the real data.
-Here we observe how introducing stringency shows an effect on the dynamics of the disease progression.
-There's still an over approximation, but the peaks of the model align with the data.}
+  \caption{\textbf{SIR Model with Lockdown Analysis for India.} This figure illustrates the fitting of the SIR model with lockdown(\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}) in comparison to real data.
+The introduction of lockdown measures showcases discernible effects on the dynamics of disease progression.
+While an overestimation persists, the model's peaks now closely align with the observed data and are able to capture key trends.}
 \end{figure}
+
 Here, it can be observed there's an overestimation of infected individuals, but, the two stages of the epidemic are being accounted for.
@@
   \end{subfigure}
-  \caption{\textbf{SIR Model with Lockdown and Vaccination for India.} The following shows the fit for the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}) compared to the real data.
-This is similar to the infections shown by the SIR model with lockdown as seen in \cref{fig:SIR_model_with_lockdown_infections_IND} and this is because the rate of vaccination is negligible \cref{eq:nu_optimal_with_lockdown_and_nu}.
+  \caption{\textbf{SIR Model with Lockdown and Vaccination for India.} This figure displays the fitting of the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}) compared to the real data.
+The infection trends closely resemble those depicted by the SIR model with lockdown, as illustrated in \cref{fig:SIR_model_with_lockdown_infections_IND} and this is because the rate of vaccination is negligible \cref{eq:nu_optimal_with_lockdown_and_nu}.
 This is suggestive of a rate of vaccination that varies with time.}
@@
 	\end{subfigure}
-  \caption{\textbf{SIR Model with Lockdown and Time-varying Vaccination Rate for India.} The following shows the fit for the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}) compared to the real data.
+  \caption{\textbf{SIR Model with Lockdown and Time-varying Vaccination Rate for India.}  This figure presents the fitting of the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}) compared to the actual data.
 This model captures the trends in infected population accurately.}
@@
   \label{fig:gdp_modelled_with_stringency_CAN}
-  \caption{\textbf{Stringency and GDP for Advanced Economies.} In economically advanced nations, when modeling the normalized GDP against stringency measures, we consistently observe substantial p-values, providing robust evidence against the null hypothesis.
+  \caption{\textbf{Stringency and GDP for Advanced Economies.} In economically advanced nations, when modeling the normalized GDP against stringency measures, we observe substantial p-values, providing evidence against the null hypothesis.
 However, the significance levels are not as high as those found for developing economies(\Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA}).
@@
 
-After median filtering to smooth the output (to reinforce the negative reward from changing the stringencies) from the trained reinforcement learning agent, here are some of the results obtained:
+After median filtering to smooth the output(to reinforce the negative reward from changing the stringencies) from the trained reinforcement learning agent, here are some of the results obtained:
 \begin{figure}[htbp!]
@@
   \caption{\textbf{Strategy from Reinforcement Learning Agent.} Here, "(actual)" is the real data, and "(rl)" is the new strategy we propose.
-\textbf{(a)} The new strategy proposed highlights a decrease in stringency from July, 2020 till October, 2020 and after which there's an increase and then a steady decline towards the end.
-\textbf{(b)} SIR Dynamics, no observable differences.
-\textbf{(c)} }
+\textbf{(a)} The new strategy proposed highlights a decrease in stringency from July, 2020 till October, 2020 compared to the actual data.
+After October, 2020 there's an increase and then a steady decline towards the end.
+\textbf{(c)} There's a peak in the number of infected people around October, 2020 and then a second peak after October, 2022 \textbf{(d)} The normalized GDP is also maintained and doesn't show a dip during April, 2022.
+\textbf{(e)} The $R_e$ is maintained below $1.5$ throughout, and below $1.2$ after October, 2020.
+\textbf{(f)} A higher reward is achieved by the reinforcement learning agent.}
 \end{figure}
 In the presented result, we can see the reinforcement learning agent outperform the modelled outcome.
-A strategic decision is made by the agent to maintain the stringency index below 80 throughout the course of the study.
-This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to $1.2$, significantly ahead of the projected timeline (refer to figure \cref{fig:176647_rl_r_eff}).
-A subsequent decrease in stringency in the later stages leads to an increase in the normalized Gross Domestic Product (GDP), indicating an economic upturn.
+A strategic decision is made by the agent to maintain the stringency index below 80 after the April, 2020 \cref{fig:176647_rl_stringency}.
+This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to below $1.2$ after October, 2020(refer to figure \cref{fig:176647_rl_r_eff}).
+
+A subsequent decrease in stringency in the later stages leads to an increase in the normalized GDP, indicating an economic upturn.
 While this strategy poses a higher risk in terms of infection rates during the initial phase of the epidemic (prior to vaccine rollout), it proves to be more beneficial for the nation's economy in the long run.
@@
   \centering
-  \caption{Strategy from Reinforcement Learning Agent 2}
+  \begin{subfigure}[t]{0.48\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{images/175975/rl_stringency.pdf}
+    \caption{Stringency changing over Time}
+  \end{subfigure}
+  \label{fig:175975_rl_stringency}
+  \hfill
   \begin{subfigure}[t]{0.48\textwidth}
@@
   \label{fig:175975_rl_i}
-
+  \hfill
   \begin{subfigure}[t]{0.48\textwidth}
@@
   \hfill
-  \begin{subfigure}[t]{0.48\textwidth}
-    \centering
-    \includegraphics[width=\linewidth]{images/175975/rl_stringency.pdf}
-    \caption{Stringency changing over Time}
-  \end{subfigure}
-  \label{fig:175975_rl_stringency}
-
   \begin{subfigure}[t]{0.48\textwidth}
@@
   \label{fig:175975_rl_reward}
+  \caption{\textbf{Strategy from Reinforcement Learning Agent.} Here, "(actual)" is the real data, and "(rl)" is the new strategy we propose.
+\textbf{(a)} The new strategy proposed highlights an increase in stringency from October, 2020 till April, 2021 compared to the actual data and then a steady decline towards the end.
+\textbf{(c)} No sharp peaks in the infected population is observed.
+\textbf{(d)} The normalized GDP is affected by the increase in the stringency from October, 2020 till April, 2021.
+\textbf{(e)} The $R_e$ is maintained below $1.2$ throughout.
+\textbf{(f)} A higher reward is achieved by the reinforcement learning agent.}
 \end{figure}
+
 An alternative outcome is presented by the agent, demonstrating a gradual escalation in stringency over time.
@@
 This approach, however, has implications for the normalized Gross Domestic Product (GDP), as seen by the observed decline in figure \cref{fig:175975_rl_gdp}.
-Furthermore, the effective reproduction number ($R_e$) does not achieve the target value of 1.2 under this strategy.
 These findings underscore the complexity of managing public health crises and the need for careful strategic planning to balance health outcomes with economic considerations.

### 2024-03-09 05_reinforcement_run.py

@@ -6,7 +6,7 @@ env.reset()
 # models/hpc_run/2667966.zip
 # models/hpc_run/4483170.zip
 # models/hpc_run_2/1225674.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run_2", "2643288.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "3320562.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 30

### 2024-03-09 paper_draft/main.tex
paper edits

@@
 The need for a nuanced understanding of how interventions impact both health outcomes and economic indicators became increasingly evident, prompting a comprehensive examination by epidemiologists to assist policy makers~\cite{Anderson2020}.
-The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022,Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, stringency measures/non-pharmaceutical interventions (NPIs)~\cite{Jalloh2022,Caldwell2021,Ferguson2020}, and financial considerations~\cite{DeFoo2023}.
+The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022,Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, non-pharmaceutical interventions(NPIs) (like restricting population movements and gatherings, closing schools and businesses, and requiring masks indoors)~\cite{Jalloh2022,Caldwell2021,Ferguson2020}, and financial considerations~\cite{DeFoo2023}.
 Despite the numerous studies conducted, very few explore how common interventions meet multiple policy objectives or how a precise articulation of the main policy goals directs the selection of the most effective interventions in terms of health and economic results~\cite{Hollingsworth2011, Song2022,Pangallo2023,Ash2022,Ohi2020, PADMANABHAN2021102676,NBERw26981,analytical_covid_lockdown_model}.
@@
 The reward function is parameterized to account for key factors influencing decision-making.
-To incentivize reduction of $R_e$(effective reproductive number) and the increase in the normalized GDP the reward is:
+To incentivize reduction of $R_e$(effective reproductive number) and the increase of the normalized GDP after $R_e$ is below $1.5$.
+The reward is defined as follows:
+
 $$\[
@@
 This reward function is parameterized to account for key factors influencing decision-making.
-When the effective reproductive number ($R_e$) exceeds 1.5, indicating a high transmission rate of the disease, the reward is negatively impacted to incentivize a reduction in $R_e$.
+When the effective reproductive number ($R_e$) exceeds $1.5$, indicating a high transmission rate of the disease, the reward is negatively impacted to incentivize a reduction in $R_e$.
 As $R_e$ decreases within the range $1.25 \leq R_e \leq 1.5$, indicating a moderate transmission rate, the reward is directly proportional to the normalized GDP, reflecting the importance of both controlling the spread of the disease and maintaining economic stability.
-Notably, when $R_e$ drops below 1, signaling a declining transmission rate and potential containment of the disease, the reward function shifts focus towards economic recovery.
+Notably, when $R_e$ drops below $1$, signaling a declining transmission rate and potential containment of the disease, the reward function shifts focus towards economic recovery.
 In this scenario, the reward incentivizes an increase in the normalized GDP, emphasizing the need to stimulate economic activity and promote recovery efforts following successful control measures.
@@
 Since Stable Baselines3 can support multiple inputs(time-series data, single data points and images) by using Dict Gym space.
-For data that varies with time (stringency, normalized GDP, $R_e$) we use a simple LSTM architecture~\cite{LSTM}.
+For data that varies with time(stringency, normalized GDP, $R_e$) we use a simple Long Short Term Memory(LSTM) architecture~\cite{LSTM}.
 For other data like, the current proportion of the population that is susceptible, infected, and recovered we use a simple fully connected layer.
@@
 % https://www.nature.com/articles/s41598-020-79147-8 (see this to write results)
-Using the simple SIR model from \Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown},  to model the disease dynamics we get:
+Using the simple SIR model from \Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown},  to model the disease dynamics we get \Crefrange{fig:SIR_model_IND}{fig:SIR_model_infections_IND}.
+Here, it can be observed that the SIR model accurately fits the susceptible population and recovered population but overestimates the infected population by a significant margin which can create complications.
+This is because disease dynamics are controlled by this population and our work involves rewarding the agent when the proportion of infected individuals falls below a predetermined threshold.
+Therefore, an overestimation of the infected population could lead to incorrect decision-making and undesirable outcomes.
+
 \begin{figure}[htbp!]
@@
 \end{figure}
-Here, it can be observed that the SIR model accurately fits the susceptible population and recovered population but overestimates the infected population by a significant margin.
-Although overestimating the infected population may not always be problematic, in our specific case, it can create complications.
-This is because our research involves rewarding the agent when the proportion of infected individuals falls below a predetermined threshold.
-Consequently, an overestimation of the infected population could lead to incorrect decision-making and undesirable outcomes.
 
-Combining the lockdown dynamics in the SIR model using \Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}, we get the following:
+Combining the lockdown dynamics in the SIR model using \Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}, we get the following \Crefrange{fig:SIR_model_with_lockdown_IND}{fig:SIR_model_with_lockdown_infections_IND}.
+Here, it can be observed there's an overestimation of infected individuals, but, the two stages of the epidemic are being accounted for.
+This is what suggests that there might be depletion of infected individuals through vaccination.
+
 \begin{figure}[htbp!]
@@
 
-Here, it can be observed there's an overestimation of infected individuals, but, the two stages of the epidemic are being accounted for.
-This is what suggests that there might be depletion of infected individuals through vaccination.
+Incorporating vaccination dynamics into the SIR model with lockdown measures, as described by \Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}, we get the following \Crefrange{fig:SIR_model_with_lockdown_with_vaccination_IND}{fig:SIR_model_with_lockdown_with_vaccination_infections_IND}.
+Here, because the value of $\nu$ \cref{eq:nu_optimal_with_lockdown_and_nu} is negligible, it doesn't change the results significantly compared to the previous model(\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown} and \Crefrange{fig:SIR_model_with_lockdown_IND}{fig:SIR_model_with_lockdown_infections_IND}).
+Therefore, a time-varying $\nu$ shall be able to better account for the these dynamics.
 
-Combining vaccination dynamics with SIR with lockdown model using \Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}, we get the following:
 \begin{figure}[htbp!]
@@
 \end{figure}
-Here, because the value of $\nu$ \cref{eq:nu_optimal_with_lockdown_and_nu} is negligible, it makes no change. However, time-varying $\nu$ shall be able to better account for the these dynamics.
 
-For SIR model with lockdown and time-varying vaccination rate from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following:
+For SIR model with lockdown and time-varying vaccination rate from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following \Crefrange{fig:SIR_model_with_lockdown_with_vaccination_time_varying_nu_IND}{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND}.
+With a time-varying $\nu$ (vaccination rate) and the effect of lockdown, our model is able to account for the infected individuals and reduce the cost in comparison to all the previously formalized models for the data.
+This shows how interventions and changes in the way people behave in response of a epidemic~\cite{Caldwell2021} play a major role in the way the epidemic unfolds.
+
 \begin{figure}[htbp!]
@@
 
-With a time-varying $\nu$ (vaccination rate) and the effect of lockdown, our model is able to account for the infected individuals and reduce the cost in comparison to all the previously formalized models for the data.
-This shows how interventions and changes in the way people behave in response of a epidemic~\cite{Caldwell2021} play a major role in the way the epidemic unfolds.
+While non-pharmaceutical interventions(NPIs) can effectively manage the epidemic, they impose economic burdens on developing nations.
+Below are scatter plots with pearson correlation, for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May, 2020 to October, 2022.
 
-However, Non-pharmaceutical Interventions (NPIs) come with costs for developing nations.
-Below are scatter plots with pearson correlation, for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May, 2020 to October, 2022.
 \begin{figure}[htbp!]
@@
 \end{figure}
-It can be observed that stringency has a negative impact on the normalized Gross domestic product (GDP).
-Therefore, in some countries policies made during an epidemic have competing costs.
-This is not the case for Advanced Economies like (USA, Japan, Canada).
+
+It can be observed from \Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA} indicates that strict policies have a negative effect on the normalized GDP in developing economies such as India, Mexico, and Brazil.
+However, this trend is not uniformly seen in advanced economies like the USA, Japan, and Canada as shown in \Crefrange{fig:stringency_vs_gdp_USA}{fig:gdp_modelled_with_stringency_CAN}.
+In these countries, other factors could be contributing to the decrease in GDP besides the implementation of stricter policies.
+
 \begin{figure}[htbp!]
@@
 \end{figure}
-From \Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_CAN} we conclude, that there are other factors that explain the dip in the normalized GDP than just an increase in the stringency in advanced economies.
 
@@
 \end{figure}
-In the presented result, we can see the reinforcement learning agent outperform the modelled outcome.
+
+In the presented result \Crefrange{fig:176647_rl_stringency}{fig:176647_rl_reward}, we can see the reinforcement learning agent outperform the modelled outcome.
 A strategic decision is made by the agent to maintain the stringency index below 80 after the April, 2020 \cref{fig:176647_rl_stringency}.
 This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to below $1.2$ after October, 2020(refer to figure \cref{fig:176647_rl_r_eff}).
+After October, 2021 there's a decrease in the stringency which leads to an increase in the normalized GDP, indicating an economic upturn.
+While this strategy poses a higher risk in terms of infection rates during the initial phase of the epidemic(prior to vaccine rollout) as well as the later phase(second peak of infected individuals \cref{fig:176647_rl_i}), but it proves to be more beneficial for the nation's economy in the long run.
+Despite the economic benefits in the long run, this strategy is not the most effective for the government to adopt due to the high number of infected individuals.
+Therefore, we propose an alternative strategy that involves some loss in terms of the economic impact.
 
-A subsequent decrease in stringency in the later stages leads to an increase in the normalized GDP, indicating an economic upturn.
-While this strategy poses a higher risk in terms of infection rates during the initial phase of the epidemic (prior to vaccine rollout), it proves to be more beneficial for the nation's economy in the long run.
-It is suggested that a more balanced policy, falling between the actual and the proposed strategy in this study, could have been a more effective approach for the government to adopt.
-This would have potentially led to a more optimal balance between controlling the spread of the disease and minimizing economic impact.
 \begin{figure}[htbp!]
@@
 
-An alternative outcome is presented by the agent, demonstrating a gradual escalation in stringency over time.
-This strategy results in the majority of infections occurring prior to the vaccine's release, with a subsequent cessation of new infections.
-This approach, however, has implications for the normalized Gross Domestic Product (GDP), as seen by the observed decline in figure \cref{fig:175975_rl_gdp}.
+In contrast, an alternative output is presented in \Crefrange{fig:175975_rl_stringency}{fig:175975_rl_r_eff}, which demonstrates a gradual increase in stringency from October, 2020 till April, 2021.
+This strategy results in a decline of infections occurring prior to the vaccine's release, and a subsequent cessation of new infections.
+This approach, however, has implications for the normalized Gross Domestic Product (GDP), as seen by the observed decline in \cref{fig:175975_rl_gdp}.
 These findings underscore the complexity of managing public health crises and the need for careful strategic planning to balance health outcomes with economic considerations.

### 2024-03-12 paper_draft/main.tex
fixes for the results section

@@
 \MakeOuterQuote{"}
-
-\tikzset{%
-   neuron missing/.style={
-    draw=none, 
-    scale=2,
-    text height=0.333cm,
-    execute at begin node=\color{black}$\vdots$
-  },
-}
 \urlstyle{same}
@@
 The need for a nuanced understanding of how interventions impact both health outcomes and economic indicators became increasingly evident, prompting a comprehensive examination by epidemiologists to assist policy makers~\cite{Anderson2020}.
-The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022,Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, non-pharmaceutical interventions(NPIs) (like restricting population movements and gatherings, closing schools and businesses, and requiring masks indoors)~\cite{Jalloh2022,Caldwell2021,Ferguson2020}, and financial considerations~\cite{DeFoo2023}.
+The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022,Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, non-pharmaceutical interventions (NPIs) like restricting population movements and gatherings, closing schools and businesses, and requiring masks indoors~\cite{Jalloh2022,Caldwell2021,Ferguson2020}, and financial considerations~\cite{DeFoo2023}.
 Despite the numerous studies conducted, very few explore how common interventions meet multiple policy objectives or how a precise articulation of the main policy goals directs the selection of the most effective interventions in terms of health and economic results~\cite{Hollingsworth2011, Song2022,Pangallo2023,Ash2022,Ohi2020, PADMANABHAN2021102676,NBERw26981,analytical_covid_lockdown_model}.
@@
 
-In summary, by using reinforcement learning augmented with deep learning techniques for the SIRV with lockdown and time-varying vaccination rate environment, we can better understand the effects of lockdown measures on both public health outcomes and economic stability during a pandemic.
+In summary, by using reinforcement learning augmented with deep learning techniques for the SIR with lockdown and time-varying vaccination rate environment, we can better understand the effects of lockdown measures on both public health outcomes and economic stability during a pandemic.
 However, it is crucial to consider the limitations of this approach and take into account a comprehensive set of decision factors in order to make informed policy decisions that are tailored to specific socioeconomic contexts.
@@
 In such cases, we consider losses from both \cref{eq:cost_SIR,eq:cost_I}.
-A weighted sum of the loss functions allows for trade-offs between comprehensive modeling and focusing on infected group dynamics(see \cref{optimizing_window_length}, where a hyperparameter(window length) is selected keeping both these losses in mind).
+A weighted sum of the loss functions allows for trade-offs between comprehensive modeling and focusing on infected group dynamics (see \cref{optimizing_window_length}, where a hyperparameter(window length) is selected keeping both these losses in mind).
 \begin{equation}
@@
   \caption{\textbf{Loss for Different Window Lengths.} We try different window lengths to find the optimal loss for both cases, either when predicting all three populations(susceptible, infected, recovered) or just the infected population.}
+  \label{fig:window_length_loss_IND}
 \end{figure}
 
-The following results in \Crefrange{fig:window_length_loss_SIR_IND}{fig:window_length_loss_I_IND}, indicate that a window length of $10$ days yields the least overall loss for all three population groups.
+The results in \cref{fig:window_length_loss_IND}, indicate that a window length of $10$ days yields the least overall loss for all three population groups.
 However, this window length results in a poor approximation for just the infected group, which is crucial for accurately modelling the spread of the disease.
@@
   \centering
-  \includegraphics[width=\linewidth]{images/interpolated_nu_varying_with_time_IND.pdf}
+  \includegraphics[scale=0.35]{images/interpolated_nu_varying_with_time_IND.pdf}
+  \caption{\textbf{$\nu$ Varying with Time.} This depicts how the vaccination rate ($\nu$) changes over time and highlights the introduction of the vaccination campaign in India.}
   \label{fig:interpolated_nu_varying_with_time_IND}
-  \caption{\textbf{$\nu$ Varying with Time.} This depicts how the vaccination rate($\nu$) changes over time and highlights the introduction of the vaccination campaign in India.}
 \end{figure}
@@
 \end{equation}
-See \Crefrange{fig:SIR_model_IND}{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND} and \Crefrange{fig:comparing_costs_SIR_IND}{fig:comparing_costs_I_IND} to see how the different models compare against each other.
+See \Crefrange{fig:SIR_model_with_lockdown_IND_parent}{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND_parent} and \cref{fig:comparing_costs_IND_parent} to see how the different models compare against each other.
 
@@
 \subsubsection{Defining the Reward Function}\label{defining_reward_function}
-The stringency index emerges as a critical factor influencing both the Gross Domestic Product (GDP) and the rate of infection spread.
+The stringency index emerges as a critical factor influencing both the normalized GDP and the rate of infection spread.
 The decision to escalate or de-escalate the stringency index is a strategic one, with significant implications.
@@
 The reward function is parameterized to account for key factors influencing decision-making.
-To incentivize reduction of $R_e$(effective reproductive number) and the increase of the normalized GDP after $R_e$ is below $1.5$.
+To incentivize reduction of $R_e$ (effective reproductive number) and the increase of the normalized GDP after $R_e$ is below $1.5$.
 The reward is defined as follows:
 
-$$\[
+\[
 \textrm{Reward} = \begin{cases}
@@
 \end{cases}
-\]$$
+\]
 
@@
 The agent observes the percentage of the population that is susceptible, infected, recovered, and time-varying data like the GDP, and previous actions taken to change the stringency, and $R_e$.
-Since Stable Baselines3 can support multiple inputs(time-series data, single data points and images) by using Dict Gym space.
-For data that varies with time(stringency, normalized GDP, $R_e$) we use a simple Long Short Term Memory(LSTM) architecture~\cite{LSTM}.
+Since Stable Baselines3 can support multiple inputs (time-series data, single data points and images) by using Dict Gym space.
+For data that varies with time (stringency, normalized GDP, $R_e$) we use a simple Long Short Term Memory (LSTM) architecture~\cite{LSTM}.
 For other data like, the current proportion of the population that is susceptible, infected, and recovered we use a simple fully connected layer.
@@
 
-\newcommand{\DrawNeuronalNetwork}[2][]{
-\xdef\Xmax{0}
-\foreach \Layer/\X/\Col/\Miss/\Lab/\Count/\Content [count=\Y] in {#2}
-{\pgfmathsetmacro{\Xmax}{max(\X,\Xmax)}
- \xdef\Xmax{\Xmax}
- \xdef\Ymax{\Y}
-}
-\foreach \Layer/\X/\Col/\Miss/\Lab/\Count/\Content [count=\Y] in {#2}
-{\node[anchor=south] at ({2*\Y},{\Xmax/2+0.1}) {\Layer};
- \foreach \m in {1,...,\X}
- {
-  \ifnum\m=\Miss
-   \node [neuron missing] (neuron-\Y-\m) at ({2*\Y},{\X/2-\m}) {};
-  \else
-   \node [circle,fill=\Col!50,minimum size=1cm] (neuron-\Y-\m) at 
-  ({2*\Y},{\X/2-\m}) {\Content};
- \ifnum\Y=1
-  \else
-   \pgfmathtruncatemacro{\LastY}{\Y-1}
-   \foreach \Z in {1,...,\LastX}
-   {
-    \ifnum\Z=\LastMiss
-    \else
-     \draw[->] (neuron-\LastY-\Z) -- (neuron-\Y-\m);
-    \fi
-    }
-  \fi
- \fi
- \ifnum\Y=1
-  \ifnum\m=\X
-   \draw [overlay] (neuron-\Y-\m) -- (state);
-  \else
-   \ifnum\m=\Miss
-   \else
-    \draw [overlay] (neuron-\Y-\m) -- (state);
-   \fi
-  \fi
- \else
- \fi     
- }
- \xdef\LastMiss{\Miss}
- \xdef\LastX{\X}
-}
-}
-\begin{tikzpicture}[x=1.5cm, y=1.5cm,
->=stealth,font=\sffamily,nodes={align=center}]
- \begin{scope}[local bounding box=T]
-  \path  node[draw,minimum width=6em,minimum height=4em] (state) {State};
-  \begin{scope}[local bounding box=NN]
-   \DrawNeuronalNetwork{Input Layer/5/green/4///,
-     Hidden Layer/5/blue/4//11/,
-     Output Layer/4/red/3//11/}
-  \end{scope}
-  \path (NN.south) node[below]{Estimated parameter\\ $\theta$};
-  \path(NN.east) -- node[above]{Policy\\ $\Pi(\theta,a)$}++ (4em,0);
- \end{scope} 
- \node[fit=(T),label={[anchor=north west]north west:Agent},inner sep=1em,draw]
-  (TF){};
- \node[below=3em of TF,draw,inner sep=1em] (Env) {Environment};
- \draw[<-] (TF.200) -- ++ (-1em,0) |- (Env.160) node[pos=0.45,right]{$r_t$};
- \draw[<-] (TF.180) -- ++ (-2em,0) |- (Env.180) node[pos=0.45,left]{$s_t$};
- \draw[->] (NN.east) -- ++ (7em,0)node[right]{$a_t$} |- (Env);
-\end{tikzpicture}
+\begin{figure}[htbp!]
+  \centering
+  \includegraphics[width=0.7\linewidth]{images/sir_arch.pdf}
+  \caption{\textbf{Deep Reinforcement Learning.} An integration of deep learning and reinforcement learning to enable more complex decision-making.}
+  \label{fig:sir_arch}
+\end{figure}
 
@@
 % https://www.nature.com/articles/s41598-020-79147-8 (see this to write results)
-Using the simple SIR model from \Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown},  to model the disease dynamics we get \Crefrange{fig:SIR_model_IND}{fig:SIR_model_infections_IND}.
+Using the simple SIR model from \Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown},  to model the disease dynamics we get \cref{fig:SIR_model_IND_parent}.
 Here, it can be observed that the SIR model accurately fits the susceptible population and recovered population but overestimates the infected population by a significant margin which can create complications.
@@
     \centering
-    \includegraphics[scale=0.5]{images/SIR_model_IND.pdf}
-    \caption{SIR Model with Lockdown}
+    \includegraphics[scale=0.35]{images/SIR_model_IND.pdf}
+    \caption{SIR Model}
     \label{fig:SIR_model_IND}
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_infections_IND.pdf}
+    \includegraphics[scale=0.35]{images/SIR_model_infections_IND.pdf}
     \caption{Infections Modelled with SIR Model}
@@
   \end{subfigure}
-  \caption{\textbf{SIR Model Comparison for India.} The figure presents a comparison between the fitted simple SIR model(\Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown}) and real data.
+  \caption{\textbf{SIR Model Comparison for India.} The figure presents a comparison between the fitted simple SIR model (\Crefrange{eq:S_without_lockdown}{eq:cost_I_without_lockdown}) and real data.
 Here an evident overestimation of the infected population is observed.}
+  \label{fig:SIR_model_IND_parent}
 \end{figure}
 
-Combining the lockdown dynamics in the SIR model using \Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}, we get the following \Crefrange{fig:SIR_model_with_lockdown_IND}{fig:SIR_model_with_lockdown_infections_IND}.
+Combining the lockdown dynamics in the SIR model using \Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}, we get the following \cref{fig:SIR_model_with_lockdown_IND_parent}.
 Here, it can be observed there's an overestimation of infected individuals, but, the two stages of the epidemic are being accounted for.
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_IND.pdf}
+    \includegraphics[scale=0.35]{images/SIR_model_with_lockdown_IND.pdf}
     \caption{SIR Model with Lockdown}
@@
   \end{subfigure}
-
   \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.50]{images/stringency_varying_with_time_IND.pdf}
+    \includegraphics[scale=0.35]{images/stringency_varying_with_time_IND.pdf}
     \caption{Stringency Varying with Time}
@@
   \end{subfigure}
-\end{figure}
-\begin{figure}[htbp!]\ContinuedFloat
   \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_infections_IND.pdf}
+    \includegraphics[scale=0.35]{images/SIR_model_with_lockdown_infections_IND.pdf}
     \caption{Infections Modelled with SIR Model with Lockdown}
@@
   \end{subfigure}
-  \caption{\textbf{SIR Model with Lockdown Analysis for India.} This figure illustrates the fitting of the SIR model with lockdown(\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}) in comparison to real data.
+  \caption{\textbf{SIR Model with Lockdown Analysis for India.} This figure illustrates the fitting of the SIR model with lockdown (\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown}) in comparison to real data.
 The introduction of lockdown measures showcases discernible effects on the dynamics of disease progression.
 While an overestimation persists, the model's peaks now closely align with the observed data and are able to capture key trends.}
+  \label{fig:SIR_model_with_lockdown_IND_parent}
 \end{figure}
 
-Incorporating vaccination dynamics into the SIR model with lockdown measures, as described by \Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}, we get the following \Crefrange{fig:SIR_model_with_lockdown_with_vaccination_IND}{fig:SIR_model_with_lockdown_with_vaccination_infections_IND}.
-Here, because the value of $\nu$ \cref{eq:nu_optimal_with_lockdown_and_nu} is negligible, it doesn't change the results significantly compared to the previous model(\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown} and \Crefrange{fig:SIR_model_with_lockdown_IND}{fig:SIR_model_with_lockdown_infections_IND}).
+Incorporating vaccination dynamics into the SIR model with lockdown measures, as described by \Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}, we get the following \cref{fig:SIR_model_with_lockdown_with_vaccination_IND_parent}.
+Here, because the value of $\nu$ \cref{eq:nu_optimal_with_lockdown_and_nu} is negligible, it doesn't change the results significantly compared to the previous model (\Crefrange{eq:S_with_lockdown}{eq:cost_I_with_lockdown} and \cref{fig:SIR_model_with_lockdown_IND_parent}).
 Therefore, a time-varying $\nu$ shall be able to better account for the these dynamics.
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_IND.pdf}
+    \includegraphics[scale=0.35]{images/SIR_model_with_lockdown_with_vaccination_IND.pdf}
     \caption{SIR Model with Lockdown and Vaccination}
@@
   \end{subfigure}
-
   \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_infections_IND.pdf}
+    \includegraphics[scale=0.35]{images/SIR_model_with_lockdown_with_vaccination_infections_IND.pdf}
     \caption{Infections Modelled with SIR Model with Lockdown and Vaccination}
@@
   \end{subfigure}
-  \caption{\textbf{SIR Model with Lockdown and Vaccination for India.} This figure displays the fitting of the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}) compared to the real data.
+  \caption{\textbf{SIR Model with Lockdown and Vaccination for India.} This figure displays the fitting of the SIR model with lockdown (\Crefrange{eq:S_with_lockdown_and_nu}{eq:cost_I_with_lockdown_and_nu}) compared to the real data.
 The infection trends closely resemble those depicted by the SIR model with lockdown, as illustrated in \cref{fig:SIR_model_with_lockdown_infections_IND} and this is because the rate of vaccination is negligible \cref{eq:nu_optimal_with_lockdown_and_nu}.
 This is suggestive of a rate of vaccination that varies with time.}
+  \label{fig:SIR_model_with_lockdown_with_vaccination_IND_parent}
 \end{figure}
 
-For SIR model with lockdown and time-varying vaccination rate from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following \Crefrange{fig:SIR_model_with_lockdown_with_vaccination_time_varying_nu_IND}{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND}.
+For SIR model with lockdown and time-varying vaccination rate from \Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}, we get the following \cref{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND_parent}.
 With a time-varying $\nu$ (vaccination rate) and the effect of lockdown, our model is able to account for the infected individuals and reduce the cost in comparison to all the previously formalized models for the data.
-This shows how interventions and changes in the way people behave in response of a epidemic~\cite{Caldwell2021} play a major role in the way the epidemic unfolds.
+This shows how interventions and changes in the way people behave in response of an epidemic~\cite{Caldwell2021} play a major role in the way the epidemic unfolds.
 
@@
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_time_varying_nu_IND.pdf}
+    \includegraphics[scale=0.35]{images/SIR_model_with_lockdown_with_vaccination_time_varying_nu_IND.pdf}
     \caption{SIR Model with Lockdown and Time-varying Vaccination Rate}
@@
   \end{subfigure}
-
   \begin{subfigure}[t]{\textwidth}
     \centering
-    \includegraphics[scale=0.50]{images/SIR_model_with_lockdown_with_vaccination_time_varying_nu_infections_IND.pdf}
+    \includegraphics[scale=0.35]{images/SIR_model_with_lockdown_with_vaccination_time_varying_nu_infections_IND.pdf}
     \caption{Infections Modelled with SIR Model with Lockdown and Time-varying Vaccination Rate}
@@
   \end{subfigure}
+  \caption{\textbf{SIR Model with Lockdown and Time-varying Vaccination Rate.} This figure displays the fitting of the SIR model with lockdown and time-varying vaccination rate (\Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}) compared to the real data.
+Incorporating a time-varying vaccination rate enhances the model's ability to capture variations in the infected population over time.}
+  \label{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND_parent}
 \end{figure}
@@
 \begin{figure}[htbp!]
-	\caption{Loss for Different Models}
 	\begin{subfigure}[t]{0.48\textwidth}
@@
 	\end{subfigure}
-  \caption{\textbf{SIR Model with Lockdown and Time-varying Vaccination Rate for India.}  This figure presents the fitting of the SIR model with lockdown(\Crefrange{eq:S_with_lockdown_and_time_varying_nu}{eq:cost_I_with_lockdown_and_time_varying_nu}) compared to the actual data.
-This model captures the trends in infected population accurately.}
+  \caption{\textbf{Loss for Different Models.} Here, we can observe that the loss is the least for the SIR model with lockdown and time-varying vaccination rate.}
+  \label{fig:comparing_costs_IND_parent}
 \end{figure}
 
-While non-pharmaceutical interventions(NPIs) can effectively manage the epidemic, they impose economic burdens on developing nations.
-Below are scatter plots with pearson correlation, for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May, 2020 to October, 2022.
+While non-pharmaceutical interventions (NPIs) can effectively manage the epidemic, they impose economic burdens on developing nations.
+In \cref{fig:stringency_gdp_developing}, we plot the normalized GDP against the stringency and calculate various metrics like the Pearson correlation coefficient, coefficient of determination ($r^2$), and p-value for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May 2020 to October 2022.
+It can be observed from \cref{fig:stringency_gdp_developing} that strict policies have a negative effect on the normalized GDP in developing economies such as India, Mexico, and Brazil.
+However, this trend is not uniformly seen in advanced economies like the USA, Japan, and Canada as shown in \cref{fig:stringency_gdp_advanced}.
+In these countries, other factors could be contributing to the decrease in normalized GDP besides the implementation of stricter policies.
 
@@
     \caption{Stringency and Normalized GDP for India}
+    \label{fig:stringency_vs_gdp_IND}
   \end{subfigure}
-  \label{fig:stringency_vs_gdp_IND}
   \hfill
@@
     \caption{Normalized GDP modelled with Stringency for India}
+    \label{fig:gdp_modelled_with_stringency_IND}
   \end{subfigure}
-  \label{fig:gdp_modelled_with_stringency_IND}
-
   \begin{subfigure}[t]{0.48\textwidth}
@@
     \caption{Stringency and Normalized GDP for Mexico}
+    \label{fig:stringency_vs_gdp_MEX}
   \end{subfigure}
-  \label{fig:stringency_vs_gdp_MEX}
   \hfill
@@
     \caption{Normalized GDP modelled with Stringency for Mexico}
+    \label{fig:gdp_modelled_with_stringency_MEX}
   \end{subfigure}
-  \label{fig:gdp_modelled_with_stringency_MEX}
-
   \begin{subfigure}[t]{0.48\textwidth}
@@
     \caption{Stringency and Normalized GDP for Brazil}
+    \label{fig:stringency_vs_gdp_BRA}
   \end{subfigure}
-  \label{fig:stringency_vs_gdp_BRA}
   \hfill
@@
     \caption{Normalized GDP modelled with Stringency for Brazil}
+    \label{fig:gdp_modelled_with_stringency_BRA}
   \end{subfigure}
-  \label{fig:gdp_modelled_with_stringency_BRA}
   \caption{\textbf{Stringency and GDP for Developing Economies.} For countries with developing economies, when we model the normalized GDP with stringency we see significant p-values and high r2 scores.}
+  \label{fig:stringency_gdp_developing}
 \end{figure}
-
-It can be observed from \Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA} indicates that strict policies have a negative effect on the normalized GDP in developing economies such as India, Mexico, and Brazil.
-However, this trend is not uniformly seen in advanced economies like the USA, Japan, and Canada as shown in \Crefrange{fig:stringency_vs_gdp_USA}{fig:gdp_modelled_with_stringency_CAN}.
-In these countries, other factors could be contributing to the decrease in GDP besides the implementation of stricter policies.
 
@@
     \caption{Stringency and Normalized GDP for United States}
+    \label{fig:stringency_vs_gdp_USA}
   \end{subfigure}
-  \label{fig:stringency_vs_gdp_USA}
   \hfill
@@
     \caption{Normalized GDP modelled with Stringency for United States}
+    \label{fig:gdp_modelled_with_stringency_USA}
   \end{subfigure}
-  \label{fig:gdp_modelled_with_stringency_USA}
-
   \begin{subfigure}[t]{0.48\textwidth}
@@
     \caption{Stringency and Normalized GDP for Japan}
+    \label{fig:stringency_vs_gdp_JPN}
   \end{subfigure}
-  \label{fig:stringency_vs_gdp_JPN}
   \hfill
@@
     \caption{Normalized GDP modelled with Stringency for Japan}
+    \label{fig:gdp_modelled_with_stringency_JPN}
   \end{subfigure}
-  \label{fig:gdp_modelled_with_stringency_JPN}
-
   \begin{subfigure}[t]{0.48\textwidth}
@@
     \caption{Stringency and Normalized GDP for Canada}
+    \label{fig:stringency_vs_gdp_CAN}
   \end{subfigure}
-  \label{fig:stringency_vs_gdp_CAN}
   \hfill
@@
     \caption{Normalized GDP modelled with Stringency for Canada}
+    \label{fig:gdp_modelled_with_stringency_CAN}
   \end{subfigure}
-  \label{fig:gdp_modelled_with_stringency_CAN}
   \caption{\textbf{Stringency and GDP for Advanced Economies.} In economically advanced nations, when modeling the normalized GDP against stringency measures, we observe substantial p-values, providing evidence against the null hypothesis.
-However, the significance levels are not as high as those found for developing economies(\Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA}).
+However, the significance levels are not as high as those found for developing economies (\Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA}).
 This is reflected in the lower R-squared scores, which indicates that the relationship between these variables may be less pronounced when compared to developing economies.}
+  \label{fig:stringency_gdp_advanced}
 \end{figure}
 
-After median filtering to smooth the output(to reinforce the negative reward from changing the stringencies) from the trained reinforcement learning agent, here are some of the results obtained:
 \begin{figure}[htbp!]
@@
     \caption{Stringency changing over Time}
+    \label{fig:176647_rl_stringency}
   \end{subfigure}
-  \label{fig:176647_rl_stringency}
   \hfill
@@
     \caption{SIR Dynamics}
+    \label{fig:176647_rl_sir}
   \end{subfigure}
-  \label{fig:176647_rl_sir}
   \hfill
@@
     \caption{Infected Population changing over Time}
+    \label{fig:176647_rl_i}
   \end{subfigure}
-  \label{fig:176647_rl_i}
   \hfill
@@
     \caption{Normalized GDP changing over Time}
+    \label{fig:176647_rl_gdp}
   \end{subfigure}
-  \label{fig:176647_rl_gdp}
   \hfill
@@
     \caption{$R_e$ changing over Time}
+    \label{fig:176647_rl_r_eff}
   \end{subfigure}
-  \label{fig:176647_rl_r_eff}
   \hfill
@@
     \caption{Reward changing over Time}
+    \label{fig:176647_rl_reward}
   \end{subfigure}
-  \label{fig:176647_rl_reward}
-  \caption{\textbf{Strategy from Reinforcement Learning Agent.} Here, "(actual)" is the real data, and "(rl)" is the new strategy we propose.
+  \caption{\textbf{Strategy from Reinforcement Learning Agent.} Here, "(actual)" is the real data, "(modelled)" is the result from real world stringency imposed with the use of the SIR model with lockdown and time-varying vaccination rate, and "(rl)" is the new stringency strategy we propose.
 \textbf{(a)} The new strategy proposed highlights a decrease in stringency from July, 2020 till October, 2020 compared to the actual data.
@@
 \textbf{(f)} A higher reward is achieved by the reinforcement learning agent.}
+  \label{fig:176647_parent}
 \end{figure}
 
-In the presented result \Crefrange{fig:176647_rl_stringency}{fig:176647_rl_reward}, we can see the reinforcement learning agent outperform the modelled outcome.
+After median filtering to smooth the output (to reinforce the negative reward from changing the stringencies) from the trained reinforcement learning agent, here are some of the results obtained.
+In the presented result \cref{fig:176647_parent}, we can see the reinforcement learning agent outperform the modelled outcome.
 A strategic decision is made by the agent to maintain the stringency index below 80 after the April, 2020 \cref{fig:176647_rl_stringency}.
-This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to below $1.2$ after October, 2020(refer to figure \cref{fig:176647_rl_r_eff}).
+This approach allows for the natural progression of disease dynamics, resulting in a rapid reduction of the effective reproduction number $R_e$ to below $1.2$ after October, 2020 (refer to figure \cref{fig:176647_rl_r_eff}).
 After October, 2021 there's a decrease in the stringency which leads to an increase in the normalized GDP, indicating an economic upturn.
-While this strategy poses a higher risk in terms of infection rates during the initial phase of the epidemic(prior to vaccine rollout) as well as the later phase(second peak of infected individuals \cref{fig:176647_rl_i}), but it proves to be more beneficial for the nation's economy in the long run.
+While this strategy poses a higher risk in terms of infection rates during the initial phase of the epidemic (prior to vaccine rollout) as well as the later phase (second peak of infected individuals \cref{fig:176647_rl_i}), but it proves to be more beneficial for the nation's economy in the long run.
 Despite the economic benefits in the long run, this strategy is not the most effective for the government to adopt due to the high number of infected individuals.
@@
     \caption{Stringency changing over Time}
+    \label{fig:175975_rl_stringency}
   \end{subfigure}
-  \label{fig:175975_rl_stringency}
   \hfill
@@
     \caption{SIR Dynamics}
+    \label{fig:175975_rl_sir}
   \end{subfigure}
-  \label{fig:175975_rl_sir}
   \hfill
@@
     \caption{Infected Population changing over Time}
+    \label{fig:175975_rl_i}
   \end{subfigure}
-  \label{fig:175975_rl_i}
   \hfill
@@
     \caption{Normalized GDP changing over Time}
+    \label{fig:175975_rl_gdp}
   \end{subfigure}
-  \label{fig:175975_rl_gdp}
   \hfill
@@
     \caption{$R_e$ changing over Time}
+    \label{fig:175975_rl_r_eff}
   \end{subfigure}
-  \label{fig:175975_rl_r_eff}
   \hfill
@@
     \caption{Reward changing over Time}
+    \label{fig:175975_rl_reward}
   \end{subfigure}
-  \label{fig:175975_rl_reward}
-  \caption{\textbf{Strategy from Reinforcement Learning Agent.} Here, "(actual)" is the real data, and "(rl)" is the new strategy we propose.
+  \caption{\textbf{Strategy from Reinforcement Learning Agent.} Here, "(actual)" is the real data, "(modelled)" is the result from real world stringency imposed with the use of the SIR model with lockdown and time-varying vaccination rate, and "(rl)" is the new stringency strategy we propose.
 \textbf{(a)} The new strategy proposed highlights an increase in stringency from October, 2020 till April, 2021 compared to the actual data and then a steady decline towards the end.
@@
 \textbf{(f)} A higher reward is achieved by the reinforcement learning agent.}
+  \label{fig:175975_parent}
 \end{figure}
 
-In contrast, an alternative output is presented in \Crefrange{fig:175975_rl_stringency}{fig:175975_rl_r_eff}, which demonstrates a gradual increase in stringency from October, 2020 till April, 2021.
+In contrast, an alternative output is presented in \cref{fig:175975_parent}, which demonstrates a gradual increase in stringency from October, 2020 till April, 2021.
 This strategy results in a decline of infections occurring prior to the vaccine's release, and a subsequent cessation of new infections.

### 2024-03-13 paper_draft/main.tex
small fixes

@@
 
-The need for a nuanced understanding of how interventions impact both health outcomes and economic indicators became increasingly evident, prompting a comprehensive examination by epidemiologists to assist policy makers~\cite{Anderson2020}.
+The need for a nuanced understanding of how interventions impact both health outcomes and economic indicators became increasingly evident, prompting a comprehensive examination by epidemiologists to assist policymakers~\cite{Anderson2020}.
 The outbreak of COVID-19 has prompted epidemiologists to research on various aspects, including mobility control~\cite{Song2022,Chinazzi2020}, vaccination strategies~\cite{Nguyen2021, Kim2022}, non-pharmaceutical interventions (NPIs) like restricting population movements and gatherings, closing schools and businesses, and requiring masks indoors~\cite{Jalloh2022,Caldwell2021,Ferguson2020}, and financial considerations~\cite{DeFoo2023}.
@@
 Many economists have studied the effect of COVID-19 on the economy of nations~\cite{Gagnon2023,GagnonKorea,Deb2020,Eichenbaum2021}.
-In advanced economies like Korea, where the stringency index was below the median the recession was milder than other advanced economies like the United Kingdom where the stringency was much higher~\cite{GagnonKorea}, they achieved it mostly with very aggressive testing, contact tracing, and enforced quarantines and isolations~\cite{Lim2023,KoreaMinister}.
+In advanced economies like Korea, where the stringency index was below the median the recession was milder than other advanced economies like the United Kingdom where the stringency was much higher~\cite{GagnonKorea}, they achieved it mostly with very aggressive testing, contact tracing, and enforced quarantines~\cite{Lim2023,KoreaMinister}.
 In India, social distancing and containment measures have been effective in reducing the number of COVID-19 cases but have come with economic costs.
@@
 Lastly, the reinforcement learning agent should be resistant to how the vaccination rate changes, and different values for $\beta$ and $\gamma$ -- keeping them the same scopes the environment for succumbing to wishful thinking which can be potentially dangerous.
-Therefore, before an actual deployment of the model, it would be a good measure to introduce stochasticity to these parameters($\beta$ and $\gamma$) and the vaccination rate.
+Therefore, before an actual deployment of the model, it would be a good measure to introduce stochasticity to these parameters ($\beta$ and $\gamma$) and the vaccination rate ($\nu$).
 
@@
     % \item \url{https://jmlr.org/papers/volume22/20-1364/20-1364.pdf} RL Paper that references \url{https://araffin.github.io/slides/rl-tuto-jnrr19/#/13} RL ppt (for our RL notes).
-    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP ($(GDP_{predicted} - GDP_{min}) / (GDP_{max} - GDP_{min})$), $R_e$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible ($S$), infected ($I$) and recovered ($R$).
+    \item Set of States $\altmathcal{S}$: The state of the environment are described through the descriptors like the normalized GDP, $R_e$, a list of all the previous actions (in changing the stringency) and the proportion of the population that was susceptible ($S$), infected ($I$) and recovered ($R$).
 The starting states are simply these values at the starting date and no previous actions.
@@
 
+\subsubsection{Deep Reinforcement Learning and Training}\label{drl_and_training}
 The agent observes the percentage of the population that is susceptible, infected, recovered, and time-varying data like the GDP, and previous actions taken to change the stringency, and $R_e$.
@@
 The output from both these networks are concatenated and used by the reinforcement learning agent to train on.
+We train the model for $2742$ time steps and some of the best results are presented.
 
@@
 \end{figure}
-
-We train the model for $2742$ time steps and some of the best results are presented below.
 
@@
 In \cref{fig:stringency_gdp_developing}, we plot the normalized GDP against the stringency and calculate various metrics like the Pearson correlation coefficient, coefficient of determination ($r^2$), and p-value for three countries (India, Mexico, Brazil) which are Emerging Market and Developing Economies~\cite{IMFCovid} from May 2020 to October 2022.
-It can be observed from \cref{fig:stringency_gdp_developing} that strict policies have a negative effect on the normalized GDP in developing economies such as India, Mexico, and Brazil.
+It can be observed from \cref{fig:stringency_gdp_developing} that strict policies have a negative effect on the normalized GDP in these economies.
 However, this trend is not uniformly seen in advanced economies like the USA, Japan, and Canada as shown in \cref{fig:stringency_gdp_advanced}.
@@
 This strategy results in a decline of infections occurring prior to the vaccine's release, and a subsequent cessation of new infections.
-This approach, however, has implications for the normalized Gross Domestic Product (GDP), as seen by the observed decline in \cref{fig:175975_rl_gdp}.
+This approach, however, has implications for the normalized GDP, as seen by the observed decline in \cref{fig:175975_rl_gdp}.
 These findings underscore the complexity of managing public health crises and the need for careful strategic planning to balance health outcomes with economic considerations.

### 2024-03-17 paper_draft/main.tex
small fixes

@@
 Our proposed approach demonstrates compelling efficacy in achieving optimal decision-making, effectively balancing the formidable challenges posed by the pandemic and economic considerations.
-We are confident that this research contribution will forge a connection between epidemic studies and reinforcement learning, offering valuable insights to help humanity better defend against the ongoing pandemic crisis.
+We are confident that this research contribution will forge a connection between epidemic studies and reinforcement learning, offering valuable insights that will help humanity better defend against potential pandemic crises in the future.
 

### 2024-03-17 05_reinforcement_run.py

@@ -6,7 +6,7 @@ env.reset()
 # models/hpc_run/2667966.zip
 # models/hpc_run/4483170.zip
 # models/hpc_run_2/1225674.zip
-model_path = os.path.join(MODELS_DIR, "hpc_run_2", "3320562.zip")
+model_path = os.path.join(MODELS_DIR, "hpc_run_2", "5069958.zip")
 model = PPO.load(model_path, env=env)
 
 episodes = 30

### 2024-03-20 paper_draft/main.tex
small fixes

@@
 Our model typically uses \cref{eq:cost_SIR}, considering all three compartments: susceptible, infected, and recovered.
-However, there are instances where we need to balance modeling all groups and focusing on the infected group, which largely drives disease spread.
+However, there are instances where we need to balance modeling all groups and focusing on the infected group (population that drives disease spread).
 In such cases, we consider losses from both \cref{eq:cost_SIR,eq:cost_I}.
-A weighted sum of the loss functions allows for trade-offs between comprehensive modeling and focusing on infected group dynamics (see \cref{optimizing_window_length}, where a hyperparameter(window length) is selected keeping both these losses in mind).
+A weighted sum of the loss functions allows for trade-offs between comprehensive modeling and focusing on infected group dynamics (see \cref{optimizing_window_length}, where a hyperparameter (window length) is selected keeping both these losses in mind).
 \begin{equation}
@@
 See \cref{fig:SIR_model_IND} to see how the model compares with the actual data.
+[TODO: should we add this to all the models or remove it for the first mention and let the reader read in the results' section which equations were used??]
 
@@
 This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}.
-Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} i.e., a window length that results in the least loss from \cref{eq:cost_SIR,eq:cost_I}.
+Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} for which the value of $\nu$ is constant and results in the least loss from \cref{eq:cost_SIR,eq:cost_I}.
 Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ then for the next time-step $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$ we again estimate $\nu$ -- so on and so forth.
@@
 \end{equation}
-This model gives shows how $\nu$ varies with time.
+% This model gives shows how $\nu$ varies with time.
 Since, the window length acts like a hyperparameter, we can try different window lengths in order to evaluate which would give us the best results.
@@
 This is how the $Q$-table updates saying, if in state $s_k$ the most ideal action is $a_t$.
-After having more and more experience with the environment and understanding which actions lead to a higher reward $r$ an optimal policy is derived be maximizing the expected value of discounted reward $J\left(r_t\right)=\mathbb{E}\left[\sum_{t=1}^{\infty} \gamma^{(t-1)} r_t\right]$, where, discount factor $\gamma \in [0, 1]$ and time steps $k = 1, 2, \dots$.
+After having more and more experience with the environment and understanding which actions lead to a higher reward $r$ an optimal policy is derived be maximizing the expected value of discounted reward $J\left(r_t\right)=\mathbb{E}\left[\sum_{t=1}^{\infty} \gamma^{(t-1)} r_t\right]$, where, discount factor $\gamma \in [0, 1]$ (in our case $\gamma = 0.99$) and time steps $k = 1, 2, \dots$.
 
@@
 
-Additionally, if the proportion of the infected population were to rise above 0.003(peak in the actual data) the model is punished ($-2000$) and otherwise rewarded ($50$).
+Additionally, if the proportion of the infected population were to rise above 0.003 (peak in the actual data) the model is punished ($-2000$) and otherwise rewarded ($50$).
 To reward not changing the stringencies frequently, we reward the absolute different between the previous stringency and the current stringency negatively ($|s(t) - s(t-1)| \times -12$).
@@
   \includegraphics[width=0.7\linewidth]{images/sir_arch.pdf}
-  \caption{\textbf{Deep Reinforcement Learning.} An integration of deep learning and reinforcement learning to enable more complex decision-making.}
+  \caption{\textbf{Deep Reinforcement Learning.} Deep learning algorithms used in reinforcement learning enables more complex decision-making.}
   \label{fig:sir_arch}
@@
 This approach, however, has implications for the normalized GDP, as seen by the observed decline in \cref{fig:175975_rl_gdp}.
-These findings underscore the complexity of managing public health crises and the need for careful strategic planning to balance health outcomes with economic considerations.
+While both these approaches (\cref{fig:175975_parent,fig:176647_parent}) outperform the actual strategy, they underscore the complexity of managing public health crises and the need for careful strategic planning to balance health outcomes with economic considerations.
 

### 2024-03-24 paper_draft/main.tex
fixes in 2.6

@@
 Now, that a relation between $\beta$ and $s(t)$ is set up, it must be investigated how stringency index affects the normalized gross domestic product (GDP)~\cite{OECDNormalizedGDP,OECDNormalizedGDP2}.
-To do this a polynomial equation of the third degree is fitted to the data points $f(x) = a x^3 + b x^2 + c x + d = y$, here, $x$ is the $s(t)$ stringency at time $t$, and $y$ the normalized GDP, and we minimize the squared error to find the values of coefficients $a, b, c, d$.
+To do this a polynomial equation of the third degree is fitted to the data points $f(x) = a x^3 + b x^2 + c x + d = y$, here, $x$ is the stringency ($s$), and $y$ the normalized GDP, and we minimize the squared error to find the values of coefficients $a, b, c, d$.
 For India after fitting a 3 degree polynomial, the following equation is obtained:

### 2024-03-31 paper_draft/main.tex
fixes in 2.6
bib without doi
APA style references

@@
 
-This \cref{fig:interpolated_nu_varying_with_time_IND} shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
+% This \cref{fig:interpolated_nu_varying_with_time_IND} shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine,PIBCovidVaccine,WikipediaCovidVaccine}.
+Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIR Model with lockdown and time-varying $\nu$.
+This \cref{fig:interpolated_nu_varying_with_time_IND} shows us that the $\nu$ coincides with the actual of data of when the vaccination drive was first launched in India~\cite{UnicefCovidVaccine}.
 Therefore, using these values we finally, recompute $\beta_{optimal}$ and $\gamma_{optimal}$ by supplying them into the equations for the SIR Model with lockdown and time-varying $\nu$.
@@
     \begin{split}
-      \textrm{normalized\_GDP} = -5.96640236 \times 10^{-5} s^{3} + 6.65064332 \times 10^{-3} s^{2} - 2.23109924 \times 10^{-1} s^{1} \\
+      \textrm{normalized\_GDP} = -5.96640236 \times 10^{-5} s^{3} + 6.65064332 \times 10^{-3} s^{2} - 2.23109924 \times 10^{-1} s \\
     + 1.01357226 \times 10^{2}
@@
 % \bibliographystyle{plain}
-\bibliography{sample}
+% \bibliographystyle{apalike}
+\bibliography{newsample}
 \end{document}

### 2024-04-01 paper_draft/main.tex
recompiling

@@
 \usepackage[english]{babel}
-\usepackage[autostyle, english = american]{csquotes}
+\usepackage[autostyle]{csquotes}
 \usepackage[nameinlink]{cleveref}
@@
 \MakeOuterQuote{"}
+\MakeAutoQuote{«}{»}
 \urlstyle{same}
@@
 \begin{equation}
-  R_{0_{optimal}} = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.762
+  R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} = 1.762
   \label{eq:r0_without_lockdown}
@@
 \end{equation}
-See \cref{fig:SIR_model_IND} to see how the model compares with the actual data.
-[TODO: should we add this to all the models or remove it for the first mention and let the reader read in the results' section which equations were used??]
+See \cref{fig:SIR_model_IND_parent} to see how the model compares with the actual data. 
+% [TODO: should we add this to all the models or remove it for the first mention and let the reader read in the results' section which equations were used??]
 
@@
   \begin{split}
+    R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} (1 - s(t)) \\
     \overbar{R_0} = 1.693 \quad \text{(Mean)} \\
@@
 \end{equation}
+See \cref{
+  fig:SIR_model_with_lockdown_IND_parent} to see how the model compares with the actual data.
+
 \subsection{SIR Model with Lockdown and Vaccination}
@@
   \begin{split}
+    R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} (1 - s(t)) \\
     \overbar{R_0} = 1.691 \\
@@
 Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} for which the value of $\nu$ is constant and results in the least loss from \cref{eq:cost_SIR,eq:cost_I}.
-Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ then for the next time-step $[{start} + {window\_length}, {start} + {window\_length} + {window\_length}]$ we again estimate $\nu$ -- so on and so forth.
+Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50 \textrm{days}$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ going through the entire data, i.e., $\nu_{optimal}$ is constant only for a specific time window.
+For example, if the entire data period is 100 days, and the window length is 10 days, then $\nu$ is estimated for days $1-10$, then for days $11-20$, and so on.
 It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
@@
 
+See \cref{
+
+fig:SIR_model_with_lockdown_with_vaccination_IND_parent} to see how the model compares with the actual data.
+
 \subsection{SIR Model with Lockdown and Time-varying Vaccination Rate}
@@
   \begin{split}
+    \overbar{\nu_{optimal}} = 0.001 \\
+    \widetilde{\nu_{optimal}} = 0.001 \\
+    \text{Mode}(\nu_{optimal}) = 0.000 \\
+    \sigma_{\nu_{optimal}} = 0.002 \\
+    \nu_{optimal} \in [0.000, 0.006]
+  \end{split}
+  \label{eq:nu_optimal_with_lockdown_and_time_varying_nu}
+\end{equation}
+\begin{equation}
+  \begin{split}
+    R_0 = \frac{\beta_{optimal}}{\gamma_{optimal}} (1 - s(t)) \\
     \overbar{R_0} = 1.546 \\
@@
 \end{equation}
-See \Crefrange{fig:SIR_model_with_lockdown_IND_parent}{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND_parent} and \cref{fig:comparing_costs_IND_parent} to see how the different models compare against each other.
+See \cref{fig:SIR_model_with_lockdown_with_vaccination_infections_time_varying_nu_IND_parent} to see how the model compares with the actual data and \cref{fig:comparing_costs_IND_parent} to see how the different models compare against each other.
 
@@
   \end{subfigure}
-  \caption{\textbf{Stringency and GDP for Developing Economies.} For countries with developing economies, when we model the normalized GDP with stringency we see significant p-values and high r2 scores.}
+  \caption{\textbf{Stringency and GDP for Developing Economies.} Here, "(actual)" is the real data, "(modelled)" is the model for normalized GDP given the stringency.
+For countries with developing economies, when we model the normalized GDP with stringency we see significant p-values and high r2 scores.}
   \label{fig:stringency_gdp_developing}
@@
   \end{subfigure}
-  \caption{\textbf{Stringency and GDP for Advanced Economies.} In economically advanced nations, when modeling the normalized GDP against stringency measures, we observe substantial p-values, providing evidence against the null hypothesis.
+  \caption{\textbf{Stringency and GDP for Advanced Economies.} Here, "(actual)" is the real data, "(modelled)" is the model for normalized GDP given the stringency.
+In economically advanced nations, when modeling the normalized GDP against stringency measures, we observe substantial p-values, providing evidence against the null hypothesis.
 However, the significance levels are not as high as those found for developing economies (\Crefrange{fig:stringency_vs_gdp_IND}{fig:gdp_modelled_with_stringency_BRA}).
@@
 \section{Discussion}
-The paper seeks to inspire epidemiologists by highlighting the advancements achieved through the application of reinforcement learning in policy making during the pandemic.
+The paper seeks to inspire epidemiologists by highlighting the advancements achieved through the application of reinforcement learning in policymaking during the pandemic.
 We introduce a virtual environment that closely simulates a pandemic scenario and thoroughly explore innovative strategies for disease mitigation using reinforcement learning.
@@
 \subsection{Dataset}
-The population-level epidemiological data can be obtained from the 'Our World In Data COVID-19' dataset: \url{https://ourworldindata.org/coronavirus} or more specifically: \url{https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-data.csv}~\cite{owidcoronavirus}.
+The population-level epidemiological data can be obtained from the "Our World In Data COVID-19" dataset: \url{https://ourworldindata.org/coronavirus} or more specifically: \url{https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-data.csv}~\cite{owidcoronavirus}.
 Data for the total cases, and recovered was acquired by scraping the Worldometers website~\cite{WorldometerCorona} used Internet Archive~\cite{InternetArchive}.
-Quaterly GDP data can be obtained from the 'Organisation for Economic Co-operation and Development': \url{https://www.oecd-ilibrary.org/economics/data/main-economic-indicators/main-economic-indicators-complete-database_data-00052-en}~\cite{economic_indicators_data}.
+Quaterly GDP data can be obtained from the "Organisation for Economic Co-operation and Development": \url{https://www.oecd-ilibrary.org/economics/data/main-economic-indicators/main-economic-indicators-complete-database_data-00052-en}~\cite{economic_indicators_data}.
 
 \subsection{Code}
-We used stable-baseline3~\cite{JMLR:v22:20-1364}, Pytorch~\cite{paszke2019pytorch}, Scipy~\cite{Virtanen_2020}, Pandas, Matplotlib, in Python~\cite{ScientificPython}.
-Code: \url{https://github.com/psymbio/sir_rl}
+% We used stable-baseline3~\cite{JMLR:v22:20-1364}, Pytorch~\cite{paszke2019pytorch}, Scipy~\cite{Virtanen_2020}, Pandas, Matplotlib, in Python~\cite{ScientificPython}.
+All code and data will be made open source upon acceptance of the paper.
+% Code: \url{https://github.com/psymbio/sir_rl}
 % \bibliographystyle{plain}

### 2024-04-04 paper_draft/main.tex
authors
changes in 2.4 optimizing window length

@@
 \author[1]{Ziya Uddin}
-
-\affil[1]{BML Munjal University}
-\affil[*]{e-mail: maeghaljain@gmail.com}
+\author[2]{Wubshet Ibrahim}
+\affil[1]{SoET, BML Munjal University, Gurugram, Haryana, 122413, India}
+\affil[2]{Department of Mathematics, Ambo University, Ambo, Ethiopia}
+\affil[*]{e-mail: wubshet.ibrahim@ambou.edu.et}
 \begin{abstract}
@@
   \begin{split}
-  \textrm{loss\_SIR} = \textrm{cost\_function\_SIR}(S, \hat{S}, I, \hat{I}, R, \hat{R}) \\
+  loss\_SIR = \textrm{cost\_function\_SIR}(S, \hat{S}, I, \hat{I}, R, \hat{R}) \\
   = L_{\delta = 1}(S, \hat{S}) + L_{\delta = 1}(I, \hat{I}) + L_{\delta = 1}(R, \hat{R})
@@
 \begin{equation}
-  \textrm{loss\_I} = \textrm{cost\_function\_I}(I, \hat{I}) = L_{\delta = 1}(I, \hat{I})
+  loss\_I = \textrm{cost\_function\_I}(I, \hat{I}) = L_{\delta = 1}(I, \hat{I})
   \label{eq:cost_I}
@@
 \begin{equation}
-  \textrm{loss\_SIR} = 85051490.533
+  loss\_SIR = 85051490.533
   \label{eq:cost_SIR_without_lockdown}
@@
 \begin{equation}
-  \textrm{loss\_I} = 45187665.281
+  loss\_I = 45187665.281
   \label{eq:cost_I_without_lockdown}
@@
 \begin{equation}
-  \textrm{loss\_SIR} = 98438821.456
+  loss\_SIR = 98438821.456
   \label{eq:cost_SIR_with_lockdown}
@@
 \begin{equation}
-  \textrm{loss\_I} = 11345389.686
+  loss\_I = 11345389.686
   \label{eq:cost_I_with_lockdown}
 \end{equation}
-See \cref{
-  fig:SIR_model_with_lockdown_IND_parent} to see how the model compares with the actual data.
+See \cref{fig:SIR_model_with_lockdown_IND_parent} to see how the model compares with the actual data.
 
@@
 \begin{equation}
-  \textrm{loss\_SIR} = 94636860.384
+  loss\_SIR = 94636860.384
   \label{eq:cost_SIR_with_lockdown_and_nu}
@@
 \begin{equation}
-  \textrm{loss\_I} = 10840360.995
+  loss\_I = 10840360.995
   \label{eq:cost_I_with_lockdown_and_nu}
@@
 This suggests to accurately estimate the infected population a time-varying vaccination rate should be used as the transition from susceptibility to direct recovery fluctuates with time~\cite{Liang2021,Marinov2022}.
-Therefore, using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}, we first find the optimal window length~\cite{Liao2020} for which the value of $\nu$ is constant and results in the least loss from \cref{eq:cost_SIR,eq:cost_I}.
-Using different window lengths ($\textrm{window\_length} = 5, 10, 15 \dots 40, 45, 50 \textrm{days}$), we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ going through the entire data, i.e., $\nu_{optimal}$ is constant only for a specific time window.
-For example, if the entire data period is 100 days, and the window length is 10 days, then $\nu$ is estimated for days $1-10$, then for days $11-20$, and so on.
-It is crucial to note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
+We first find the optimal window length~\cite{Liao2020} for which the value of $\nu$ is constant for a time sub-interval and results in the least loss from \cref{eq:cost_SIR,eq:cost_I}.
+For this we use different window lengths ($window\_lengths = 5, 10, 15 \dots 40, 45, 50 \textrm{ days}$).
+
 \begin{equation}
@@
 \end{equation}
+
+Using the model described by \Crefrange{eq:S_with_lockdown_and_nu_calc_nu}{eq:R_with_lockdown_and_nu_calc_nu} for each $window\_length_{i}$ in $window\_lengths$, where $i = 1, 2, 3 \dots 10$, we calculate the $time\_varying\_\nu$.
+For each $time\_varying\_\nu$ we then calculate the $loss\_SIR$ and $loss\_I$.
+This is done as follows:
+
+\begin{enumerate}
+  \item Let $start = 1$ (initial start day), $time\_varying\_\nu = [ ]$ (empty list).
+  \item Repeat the following steps until $start + window\_length_{i}$ exceeds the total number of days in the data:
+  \begin{enumerate}
+    \item Estimate the value of $\nu$ for the sub-interval $[start, start + window\_length_{i}]$.
+    \item Update $start = start + window\_length_{i}$.
+    \item $time\_varying\_\nu.append(\nu)$
+  \end{enumerate}
+  \item Calculate $loss\_SIR$ and $loss\_I$ using the $time\_varying\_\nu$ using the $\beta_{optimal}$ and $\gamma_{optimal}$ from \cref{eq:beta_optimal_with_lockdown_and_nu} and \cref{eq:gamma_optimal_with_lockdown_and_nu}.
+The loss from the different $window\_lengths$ is compared in \cref{fig:window_length_loss_IND}.
+\end{enumerate}
+
+Note that the variable $\nu$ is constrained to be a positive integer, reflecting the inherent one-way nature of vaccination: individuals can only receive vaccinations, not return them.
+
+% , we estimate value of $\nu$ given a sub-interval of $[{start}, {start} + {window\_length}]$ going through the entire data, i.e., $\nu_{optimal}$ is constant only for a specific time window.
+For example, if the entire data period is 100 days, and the window length is 10 days, then $\nu$ is estimated for days $1-10$, then for days $11-20$, and so on.
 % This model gives shows how $\nu$ varies with time.
@@
 Therefore, the below graphs show us the loss using \cref{eq:cost_SIR,eq:cost_I}.
-
 \begin{figure}[htbp!]
@@
 
-See \cref{
-
-fig:SIR_model_with_lockdown_with_vaccination_IND_parent} to see how the model compares with the actual data.
+See \cref{fig:SIR_model_with_lockdown_with_vaccination_IND_parent} to see how the model compares with the actual data.
 
@@
 \begin{equation}
-  \textrm{loss\_SIR} = 29116762.926
+  loss\_SIR = 29116762.926
   \label{eq:cost_SIR_with_lockdown_and_time_varying_nu}
@@
 \begin{equation}
-  \textrm{loss\_I} = 658537.443
+  loss\_I = 658537.443
   \label{eq:cost_I_with_lockdown_and_time_varying_nu}
